---
title: "Exposome AIRWAYS Analysis"
subtitle: "Quality Control Analysis"
author: "Jason Laird"
date: today 
date-format: short
engine: knitr
execute: 
  warning: false
  message: false
  cache: true
format: 
  html:
    code-fold: true
    theme: 
      - simplex
      - custom.scss
    fontcolor: '#0c051a'
---


## Setup 

### Load Libraries

```{r}
#| warning: false
#| message: false

library(tidyverse)
library(janitor)
library(ggpubr)
library(tidyexposomics)
library(MultiAssayExperiment)
```

### Load Data

```{r}
# Meta Data
aw_meta_cv2 <- readRDS("../results/input_data/aw_meta_cv2.rds")

# Codebook
aw_cb <- readRDS("../results/input_data/aw_cb.rds")

# Abundance/Expression Data
cd4_gene_counts   <- readRDS("../results/input_data/cd4_gene_counts.rds")
cd16_gene_counts  <- readRDS("../results/input_data/cd16_gene_counts.rds")

cd4_isoform_counts  <- readRDS("../results/input_data/cd4_isoform_counts.rds")
cd16_isoform_counts <- readRDS("../results/input_data/cd16_isoform_counts.rds")

prot_abd <- readRDS("../results/input_data/prot_abd.rds")
adduct   <- readRDS("../results/input_data/adduct.rds")
adduct_load   <- readRDS("../results/input_data/adduct_load.rds")

# Feature Data
gene_fdata    <- readRDS("../results/input_data/gene_fdata.rds")
isoform_fdata <- readRDS("../results/input_data/isoform_fdata.rds")
prot_fdata    <- readRDS("../results/input_data/prot_fdata.rds")
adduct_fdata  <- readRDS("../results/input_data/adduct_fdata.rds")
adduct_load_fdata <- readRDS("../results/input_data/adduct_load_fdata.rds")

```


### Prepare Data Format

```{r}
omics_list <- list(
  "CD4 T cell RNA"       = cd4_gene_counts,
  "CD16 Monocyte RNA"      = cd16_gene_counts,
  "CD4 T cell Isoforms"  = cd4_isoform_counts,
  "CD16 Monocyte Isoforms" = cd16_isoform_counts,
  "Proteomics"            = prot_abd,
  "Adductomics"           = adduct,
  "Adduct Burden"      = adduct_load
)

fdata <- list(
  "CD4 T cell RNA"       = gene_fdata,
  "CD16 Monocyte RNA"      = gene_fdata,     
  "CD4 T cell Isoforms"  = isoform_fdata,
  "CD16 Monocyte Isoforms" = isoform_fdata,   
  "Proteomics"            = prot_fdata,
  "Adductomics"           = adduct_fdata,
  "Adduct Burden"      = adduct_load_fdata
)

rm(list = ls()[grepl("_fdata|counts|_abd|adduct", ls())])


expom <- create_exposomicset(
  codebook = aw_cb,
  exposure = aw_meta_cv2,
  omics = omics_list,
  row_data = fdata
)

# Keep samples with sex and age
expom <- expom[,!is.na(expom$age) & !is.na(expom$gli_sex)]

exp_cat <- c(
  "Indoor Air",
  "Parabens",
  "Phenols",
  "Antimicrobials",
  
  "Dust Allergens",
  "IgE Respiratory",
  "IgE Superantigen",
  "IgE Total",
  "IgE Food",
  
  "Serum Essential Metals",
  "Serum Non-Essential Metals",
  "Urine Essential Metals",
  "Urine Non-Essential Metals"
  
  )

exp_vars <- extract_results(exposomicset = expom,result = "codebook") |>
  filter(category %in% exp_cat) |> 
  pull(variable)
```

## Missingness/Imputation

```{r}
#| fig-width: 6
#| fig-height: 6
expom |>
    plot_missing(
        plot_type = "summary",
        threshold = 20
    )
```

### LOD Impute Before Removal

Keeping variables where missingness may be due to non-detects

#### Fix

```{r}
trace(run_impute_missing,edit=T)
# remove if condition and replace with:
# (!is.null(exposure_impute_method))
```

```{r}
# Grab codebook variables that are actually in colData
codebook_vars <- extract_results(expom, result = "codebook") |>
  filter(variable %in% (
    expom |> tidyexposomics::pivot_sample() |> colnames()
  ))

# LOD-style imputation
lod_cats <- c(
  
  "Indoor Air",
  "Parabens",
  "Phenols",
  "Antimicrobials",
  
  "Dust Allergens",
  "IgE Respiratory",
  "IgE Superantigen",
  "IgE Total",
  "IgE Food",
  
  "Serum Essential Metals",
  "Serum Non-Essential Metals",
  "Urine Essential Metals",
  "Urine Non-Essential Metals",
  
  "Serum Cytokines",
  "Nasal Cytokines"
  
)

# Split into two lists of variables
lod_vars <- codebook_vars |> 
  filter(category %in% lod_cats) |> 
  pull(variable)

# remove variables with no variance
lod_vars <- expom |>
  tidyexposomics::pivot_sample() |>
  dplyr::select(all_of(lod_vars)) |>
  apply(2,function(col){
    var <- col[!is.na(col)] |> var();
    return(var>0)
  }) |> 
  (\(bool) bool[bool %in% c(TRUE)])() |> 
  names()

# now confirm the missingness of each:
lod_vars <- expom |> 
  tidyexposomics::pivot_sample() |> 
  dplyr::select(all_of(lod_vars)) |> 
  naniar::miss_var_summary() |> 
  filter(pct_miss!=100) |> 
  pull(variable)

# looks like half the samples were just not profiled
# for some chemicals, removing these since 
# they are not non-detects
lod_vars <- lod_vars[!grepl("_sg",lod_vars)]
  
# now let's impute
expom <- expom |>
  run_impute_missing(
    exposure_impute_method = "lod_sqrt2",
    exposure_cols = lod_vars
  ) 
```

### Filter Missing Variables

```{r}
#| message: false
#| warning: false


expom <- expom |> 
  filter_missing(na_thresh = 20)

# update exp vars after filtering
exp_vars <- exp_vars[
  exp_vars %in% names(colData(expom))
]

```

### Impute using MissForest for the Other Variables

```{r}
# Define missForest variables to impute
mf_vars <- expom |>
    tidyexposomics::pivot_sample() |> 
  naniar::miss_var_summary() |> 
  filter(pct_miss>0) |> 
  pull(variable)

# Now run imputation by bucket
expom <- expom |>
  run_impute_missing(
    exposure_impute_method = "missforest",
    exposure_cols = mf_vars,
    omics_impute_method  = "missforest",
    omics_to_impute = c(
      "Proteomics","Adductomics")
  ) 
  
```


## Filtering Omics

```{r}
expom <- expom |>
    filter_omics(
        method = "expression",
        assays = "CD4 T cell RNA",
        assay_name = 1,
        min_value = 1,
        min_prop = 0.3
    ) |>
  filter_omics(
        method = "expression",
        assays = "CD16 Monocyte RNA",
        assay_name = 1,
        min_value = 1,
        min_prop = 0.3
    ) |>
  filter_omics(
        method = "expression",
        assays = "CD4 T cell Isoforms",
        assay_name = 1,
        min_value = 1,
        min_prop = 0.3
    ) |>
  filter_omics(
        method = "expression",
        assays = "CD16 Monocyte Isoforms",
        assay_name = 1,
        min_value = 1,
        min_prop = 0.3
    ) |>
    filter_omics(
        method = "variance",
        assays = "Proteomics",
        assay_name = 1,
        min_var = 0.01
    ) 
```

## Normality Check

```{r}

num_exp_vars <- expom |>
  tidyexposomics::pivot_sample() |> 
  dplyr::select(all_of(exp_vars)) |> 
  dplyr::select_if(is.numeric) |> 
  dplyr::select_if(~length(unique(.))>3) |> 
  # keeping temperature as is
  dplyr::select(-c(iv_tempavg)) |> 
  colnames()

# Check variable normality & transform variables
expom <- expom |>
    # Check variable normality
    run_normality_check(action = "add")  |> 
    # Transform variables
    transform_exposure(
        transform_method = "log2",
        exposure_cols = num_exp_vars
    )
```

## PCA 

```{r}
#| fig-width: 10
#| fig-height: 8


# Perform principal component analysis
expom <- expom |>
    run_pca(
        log_trans_exp = TRUE,
        log_trans_omics = TRUE,
        action = "add"
    )
# Plot principal component analysis results
expom |>
    plot_pca()
```

```{r}
# Filter out sample outliers
expom <- expom |>
    filter_sample_outliers(
        outliers = c("s937", "s3331","s3692")
    )
```

## Principal Component Correlation

```{r}
#| fig-width: 8
#| fig-height: 10
expom <- expom |>
    run_correlation(
        feature_type = "pcs",
        exposure_cols = num_exp_vars,
        n_pcs = 30,
        action = "add",
        correlation_cutoff = 0,
        pval_cutoff = 1
    )

expom |>
    plot_correlation_tile(
        feature_type = "pcs",
        pval_cutoff = 0.1
    )
```
## Summary

* Adductomics data have the most missing data, followed by exposures, then proteomics.

* Most variables are not normally distributed.

* Variation in the omics data appears to be primarily driven by transcriptomics data.

* Two outliers were detected and removed from the data.

* Urinary metals and several IgE subtypes are associated with the most principal components, and likely are associated with features that vary in our data.


## Save Data

```{r}
saveRDS(expom,file="../results/qc/expom.rds")
saveRDS(exp_vars,file="../results/qc/exp_vars.rds")
saveRDS(num_exp_vars,file="../results/qc/num_exp_vars.rds")
```

## Session Info

```{r}
sessionInfo()
```

