---
title: "Exposome AIRWAYS Analysis"
subtitle: "Differential Abundance Analysis"
author: "Jason Laird"
date: today 
date-format: short
engine: knitr
execute: 
  warning: false
  message: false
  cache: true
format: 
  html:
    code-fold: true
    theme: 
      - simplex
      - custom.scss
    fontcolor: '#0c051a'
---


## Setup 

### Load Libraries

```{r}
#| warning: false
#| message: false

library(tidyverse)
library(janitor)
library(ggpubr)
library(ggsci)
library(tidyexposomics)
library(MultiAssayExperiment)
library(GSVA)

# Source QC helper functions
invisible(lapply(
  list.files(
    path = "./",
    pattern = "*.R", full.names = TRUE
  ),
  source
))
```

### Load Data

```{r}
# Codebook
aw_cb <- readRDS("../results/input_data/aw_cb.rds")

# MultiAssayExperiment Object
fev1_fvc_expom <- readRDS("../results/exwas/fev1_fvc_expom.rds")

# Exposure variables
exp_vars <- readRDS("../results/qc/exp_vars.rds")
num_exp_vars <- readRDS("../results/qc/num_exp_vars.rds")

# All variables
vars <- aw_cb$variable |> 
  (\(chr)chr[chr %in% colnames(colData(fev1_fvc_expom))])()

# Setting colors for exposure categories
cat_colors <- scales::alpha(get_palette("aaas",13), 0.5) |> 
  (\(colors){
    names(colors) <- aw_cb |> 
      filter(variable %in% exp_vars) |> 
      pull(category) |> 
      unique()
    colors
  })()
```


## Differential Expression


### FEV1/FVC

```{r}
#| fig-width: 10
#| fig-height: 6


# Run differential abundance analysis
fev1_fvc_expom <- fev1_fvc_expom |>
    run_differential_abundance(
        formula = ~ pftfev1fvc_actual + gli_age + gli_sex + gli_height + fis,
        method = "limma_trend",
        scaling_method = "none",
        action = "add"
    )
```
### Number of DEGs

```{r}
#| fig-width: 5
#| fig-height: 3.5


fev1_fvc_expom |> 
    extract_results("differential_analysis") |> 
    pluck("differential_abundance") |> 
    filter(P.Value<0.05) |>
    mutate(direction=ifelse(logFC>0,"Upregulated","Downregulated")) |> 
    group_by(exp_name,direction) |> 
    reframe(value=n()) |> 
    group_by(exp_name) |> 
    mutate(total=sum(value)) |> 
    ungroup() |> 
  ggplot(aes(
        x=value,
        y=reorder(exp_name,total),
        fill=direction
    ))+
    geom_col(alpha = 0.7,colour = "black",linewidth = 0.005) +
    scale_fill_manual(values = c(
        "blue4","red4"
    ))+
    theme_classic()+
    theme(
        plot.title = element_text(face = "bold.italic"),
        plot.subtitle = element_text(face = "italic"),
        plot.caption = element_text(face = "italic")
    ) +
  scale_y_discrete(labels=c(
  "CD4 T cell RNA"       = expression(CD4^"+" ~ "T cell RNA"),
  "CD4 T cell Isoforms"  = expression(CD4^"+" ~ "T cell Isoforms"),
  "CD4 T cell miRNA"     = expression(CD4^"+" ~ "T cell miRNA"),
  
  "CD16 Monocyte RNA"       = expression(CD16^"+" ~ "Monocyte RNA"),
  "CD16 Monocyte Isoforms"  = expression(CD16^"+" ~ "Monocyte Isoforms"),
  
  "Adductomics"       = "Adductomics",
  "Adductomics Load"  = "Adductomics Load",
  "Proteomics"        = "Proteomics"
))+
    labs(
        x = "No. of Differential Features",
        y = "",
        fill = "Direction",
        title = "FEV1/FVC",
        subtitle = "",
        caption = "Nominal P < 0.05"
    )
```

### Number of Overlapping DEGs

```{r}
#| fig-width: 10
#| fig-height: 10
source("./get_pairwise_overlaps.R")
fev1_fvc_expom |> 
  extract_results("differential_analysis") |> 
  pluck("differential_abundance") |> 
  filter(P.Value<0.05) |> 
  (\(df)split(df,df$exp_name))() |> 
  map(~.x |> 
        pull(feature_map) |> 
        unique()) |> 
  get_pairwise_overlaps() |> 
  arrange(jaccard) |> 
  tidygraph::as_tbl_graph() |> 
  ggraph::ggraph(
        layout = "linear",
        circular = TRUE
    ) +
        ggraph::geom_edge_arc(aes(
            color = jaccard,
            width = log10(num_shared+1)*2
        )) +
        ggraph::geom_node_point(shape = 21,
            size = 4,
            alpha = 0.8
        ) +
        ggraph::geom_node_text(
            aes(
                label = name,
                x = 1.1 * x,
                y = 1.1 * y,
                angle = ggraph::node_angle(x, y)
            ),
            hjust = "outward",
            fontface = "bold.italic",
            check_overlap = TRUE
        ) +
        ggraph::theme_graph() +
        ggraph::scale_edge_color_gradient2(
        high = "#0f0c29",
        mid = "white",
        low = "#43C6AC",
        midpoint = 0.1,
        guide = ggraph::guide_edge_colorbar()
    )+
        ggplot2::coord_fixed(xlim = c(-2, 2), ylim = c(-2, 2)) +
        ggplot2::guides(edge_width = "none") +
        ggplot2::labs(
            edge_color = "Jaccard Coef."
        )
```

### Median LogFC of DEGs

```{r}
#| fig-width: 4.5
#| fig-height: 4
fev1_fvc_expom |> 
  extract_results("differential_analysis") |> 
  pluck("differential_abundance") |> 
  filter(P.Value<0.05) |> 
  group_by(exp_name) |> 
  mutate(mean_abs_logfc=mean(abs(logFC))) |> 
  ungroup() |> 
  ggplot(aes(
    x=reorder(exp_name,mean_abs_logfc),
    y=abs(logFC),
    fill=exp_name,
    color=exp_name
  ))+
  geom_boxplot(alpha=.5) +
  theme_pubr(legend = "none")+
  rotate_x_text(angle=45)+
  ggsci::scale_fill_npg()+
  ggsci::scale_color_npg()+
  labs(
    x="",
    y=expression("Abs(Log"[2]*"FC)")
  )
```

## Sensitivity Analysis

### FEV1/FVC

```{r}
# Perform Sensitivity Analysis
fev1_fvc_expom <- fev1_fvc_expom |>
    run_sensitivity_analysis(
        base_formula = ~ pftfev1fvc_actual + gli_age + gli_sex + gli_height + fis,
        methods = c("limma_trend"),
        scaling_methods = c("none"),
        covariates_to_remove = c(
            "gli_age",
            "gli_sex",
            "gli_height",
            "fis"
        ),
        pval_col = "P.Value",
        logfc_col = "logFC",
        logFC_threshold = log2(1),
        pval_threshold = 0.05,
        stability_metric = "stability_score",
        bootstrap_n = 100,
        action = "add"
    )

fev1_fvc_expom@metadata$differential_analysis$sensitivity_analysis$sensitivity_df <- NULL
```

## Isoform Switching Analysis

### SatuRn

```{r}
cd4_se <- tidyexposomics:::.update_assay_colData(fev1_fvc_expom,"CD4 T cell Isoforms")

cd16_se <- tidyexposomics:::.update_assay_colData(fev1_fvc_expom,"CD16 Monocyte Isoforms")

cd4_iso_res <- run_isoform_switching_satuRn(
    se = cd4_se,
    counts_assay = "counts",
    tx2gene_df = rowData(cd4_se) |> 
      as.data.frame() |> 
      dplyr::select(isoform_id=transcript_id,
                    gene_id=gene_id),
    formula = ~ pftfev1fvc_actual + gli_age + gli_sex + gli_height + fis,
    contrast_var = "pftfev1fvc_actual",
    verbose = TRUE
)

cd16_iso_res <- run_isoform_switching_satuRn(
    se = cd16_se,
    counts_assay = "counts",
    tx2gene_df = rowData(cd16_se) |>
      as.data.frame() |> 
      dplyr::select(isoform_id=transcript_id,
                    gene_id=gene_id),
    formula = ~ pftfev1fvc_actual + gli_age + gli_sex + gli_height + fis,
    contrast_var = "pftfev1fvc_actual",
    verbose = TRUE
)
```


## Enrichment Analysis

```{r}
# deg_df <- extract_results(fev1_fvc_expom,
#                           result = "differential_analysis") |> 
#   pluck("differential_abundance") |> 
#   # temp fix for adduct burden data
#   mutate(feature_map=case_when(
#     exp_name == "Adduct Burden" ~ feature_name,
#     .default = feature_map
#   )) |> 
#   filter(P.Value < 0.05)

robust_deg_df <- fev1_fvc_expom@metadata |> 
  pluck("differential_analysis",
        "sensitivity_analysis",
        "feature_stability") |> 
  filter(effect_consistency>.7) |> 
  inner_join(
    fev1_fvc_expom@metadata |> 
  pluck("differential_analysis",
        "differential_abundance"),
  by=c("exp_name"="exp_name",
       "feature"="feature")) |>
  filter(P.Value<0.05)

robust_deg_lst <- robust_deg_df |> 
  mutate(direction = ifelse(logFC > 0, "Upregulated", "Downregulated")) |> 
  group_by(exp_name, direction) |> 
  group_split(.keep = TRUE) |> 
  set_names(
    robust_deg_df |> 
      mutate(direction = ifelse(logFC > 0, "Upregulated", "Downregulated")) |> 
      group_by(exp_name, direction) |> 
      group_keys() |> 
      transmute(name = paste(exp_name, direction, sep = "_")) |> 
      pull(name)
  ) |> 
  map(
    ~.x$feature_map
  ) 

# Ensure there are enough for enrichment
robust_deg_lst <- robust_deg_lst[lengths(robust_deg_lst)>20]

# # DTU results
# dtu_lst <- cd4_iso_res$results |> 
#   mutate(exp_name="CD4 T cell Isoforms DTU") |> 
#   bind_rows(cd16_iso_res$results |> 
#               mutate(exp_name="CD16 Monocyte Isoforms DTU") ) |> 
#   mutate(direction = ifelse(estimates > 0, "Upregulated", "Downregulated")) |> 
#   filter(pval<0.05) |> 
#   group_by(exp_name, direction) |> 
#   group_split(.keep = TRUE) |> 
#   set_names(
#     cd4_iso_res$results |> 
#       mutate(exp_name="CD4 T cell Isoforms DTU") |> 
#       bind_rows(cd16_iso_res$results |> 
#                   mutate(exp_name="CD16 Monocyte Isoforms DTU") ) |> 
#       mutate(direction = ifelse(estimates > 0, "Upregulated", "Downregulated")) |> 
#       filter(pval<0.05) |> 
#       group_by(exp_name, direction) |> 
#       mutate(direction = ifelse(estimates > 0, "Upregulated", "Downregulated")) |> 
#       group_by(exp_name, direction) |> 
#       group_keys() |> 
#       transmute(name = paste(exp_name, direction, sep = "_")) |> 
#       pull(name)
#   ) |> 
#   map(
#     ~.x$gene_name
#   ) 

rna_features <- pivot_feature(fev1_fvc_expom) |>
  filter(.exp_name =="CD4 T cell RNA") |>
  pull(feature_map)

isoform_features <- pivot_feature(fev1_fvc_expom) |>
  filter(.exp_name =="CD4 T cell Isoforms") |>
  pull(feature_map)

library(furrr)

plan(multisession)

enr_res_rna <- future_map(
  robust_deg_lst |> 
    (\(lst)lst[grepl("RNA",names(lst))])(),
  ~{
    options(GO_ANNOTATION_URL = "http://current.geneontology.org/annotations/goa_human.gaf.gz")
    
    run_fenr(
      selected_genes = .x,
      universe_genes = rna_features,
      db = "GO",
      species = "goa_human",
      feature_col = "gene_symbol"
    )
  },
  .progress = TRUE
) |> 
  bind_rows(.id = "group")


enr_res_iso <- future_map(
  robust_deg_lst |> 
    (\(lst)lst[grepl("Isoform",names(lst))])(),
  ~{
    options(GO_ANNOTATION_URL = "http://current.geneontology.org/annotations/goa_human.gaf.gz")
    
    run_fenr(
      selected_genes = .x,
      universe_genes = isoform_features,
      db = "GO",
      species = "goa_human",
      feature_col = "gene_symbol"
    )
  },
  .progress = TRUE
) |> 
  bind_rows(.id = "group")

enr_res_sig <- enr_res_rna |> 
  bind_rows(enr_res_iso) |> 
  filter(p_adjust<.1,
         n_with_sel>5,
         N_with< 1000)

parent_terms <- find_parent_terms(enr_res_sig)

enr_res_sig <- enr_res_sig |> 
  left_join(parent_terms |> 
              dplyr::select(term,parentTerm),
            by=c("term_name"="term"))

enr_res_sig |> 
  tabyl(group) |> 
  view()
```

### Cell-Type Terms

```{r}
cd4_terms <- enr_res_sig |> 
  mutate(cell=case_when(
    grepl("CD16",group) ~ "CD16+ Monocyte",
    grepl("CD4",group) ~ "CD4+ T cell"
  )) |>
  filter(cell=="CD4+ T cell") |> 
  group_by(term_name) |> 
  reframe(n=length(unique(group)),
          median_p=median(p_adjust),
          median_count=median(n_with_sel))

cd16_terms <- enr_res_sig |> 
  mutate(cell=case_when(
    grepl("CD16",group) ~ "CD16+ Monocyte",
    grepl("CD4",group) ~ "CD4+ T cell"
  )) |>
  filter(cell=="CD16+ Monocyte") |> 
  group_by(term_name) |> 
  reframe(n=length(unique(group)),
          median_p=median(p_adjust),
          median_count=median(n_with_sel))

cd4_unique_terms <- cd4_terms |> 
  filter(term_name %in% 
           setdiff(cd4_terms$term_name,cd16_terms$term_name))

cd16_unique_terms <- cd16_terms |> 
  filter(term_name %in% 
           setdiff(cd16_terms$term_name,cd4_terms$term_name))

shared_terms <- enr_res_sig |> 
  mutate(cell=case_when(
    grepl("CD16",group) ~ "CD16+ Monocyte",
    grepl("CD4",group) ~ "CD4+ T cell"
  )) |>
  group_by(term_name) |> 
  reframe(n=length(unique(group)),
          median_p=median(p_adjust),
          median_count=median(n_with_sel)) |> 
  filter(term_name %in%  
           Reduce(
             intersect,
             list(cd4_terms$term_name,cd16_terms$term_name))
  )

```



### DotPlot

#### Shared

```{r}
enr_res_sig |> 
    filter(term_name %in% (
        shared_terms |> 
            arrange(desc(n)) |> 
            slice_head(n=50) |> 
            pull(term_name)
    )) |> 
    mutate(
        cell=case_when(
            grepl("CD16",group) ~ "CD16+ Monocyte",
            grepl("CD4",group) ~ "CD4+ T cell"
        ),
        direction=case_when(
            grepl("Down",group) ~ "Downregulated",
            grepl("Up",group) ~ "Upregulated"
        )) |> 
    mutate(group=gsub("_.*","",group)) |> 
    ggplot(aes(
        x=group,
        y=fct_reorder2(term_name,paste(group,direction),direction),
        size=n_with_sel,
        color=-log10(p_adjust+1e-300)
    ))+
    geom_point()+
    scale_color_gradient(
        low = "#43C6AC",
        high = "#191654"
    )+
    theme_bw()+
    facet_grid(. ~ cell+direction,space = "free",scales="free")+
    rotate_x_text(angle=45)+
    scale_x_discrete(labels=c(
        "CD4 T cell RNA"             = expression(CD4^"+" ~ "T cell RNA"),
        "CD4 T cell Isoforms"        = expression(CD4^"+" ~ "T cell Isoforms"),
        "CD4 T cell Isoforms DTU"    = expression(CD4^"+" ~ "T cell Isoforms DTU"),
        
        "CD16 Monocyte RNA"          = expression(CD16^"+" ~ "Monocyte RNA"),
        "CD16 Monocyte Isoforms"     = expression(CD16^"+" ~ "Monocyte Isoforms"),
        "CD16 Monocyte Isoforms DTU" = expression(CD16^"+" ~ "Monocyte Isoforms DTU")
    ))+
    labs(
        x="",
        y="",
        color=expression("-Log"[10]*"P"),
        size="Count"
    )+
    theme(strip.text = element_text(face="bold.italic"))
```


#### CD4+ T cell Unique

```{r}
enr_res_sig |> 
    filter(term_name %in% (
        cd4_unique_terms |> 
            arrange(desc(n)) |> 
            slice_head(n=40) |> 
            pull(term_name)
    )) |> 
    mutate(
        cell=case_when(
            grepl("CD16",group) ~ "CD16+ Monocyte",
            grepl("CD4",group) ~ "CD4+ T cell"
        ),
        direction=case_when(
            grepl("Down",group) ~ "Downregulated",
            grepl("Up",group) ~ "Upregulated"
        )) |> 
    mutate(group=gsub("_.*","",group)) |> 
    ggplot(aes(
        x=group,
        y=fct_reorder2(term_name,paste(group,direction),direction),
        size=n_with_sel,
        color=-log10(p_adjust+1e-300)
    ))+
    geom_point()+
    scale_color_gradient(
        low = "#43C6AC",
        high = "#191654"
    )+
    theme_bw()+
    facet_grid(cell ~ direction,space = "free",scales="free")+
    rotate_x_text(angle=45)+
    scale_x_discrete(labels=c(
        "CD4 T cell RNA"             = expression(CD4^"+" ~ "T cell RNA"),
        "CD4 T cell Isoforms"        = expression(CD4^"+" ~ "T cell Isoforms"),
        "CD4 T cell Isoforms DTU"    = expression(CD4^"+" ~ "T cell Isoforms DTU"),
        
        "CD16 Monocyte RNA"          = expression(CD16^"+" ~ "Monocyte RNA"),
        "CD16 Monocyte Isoforms"     = expression(CD16^"+" ~ "Monocyte Isoforms"),
        "CD16 Monocyte Isoforms DTU" = expression(CD16^"+" ~ "Monocyte Isoforms DTU")
    ))+
    labs(
        x="",
        y="",
        color=expression("-Log"[10]*"P"),
        size="Count"
    )+
    theme(strip.text = element_text(face="bold.italic"))
```


#### CD16+ Monocyte Unique

```{r}
enr_res_sig |> 
    filter(term_name %in% (
        cd16_unique_terms |> 
            arrange(desc(n)) |> 
            slice_head(n=40) |> 
            pull(term_name)
    )) |> 
    mutate(
        cell=case_when(
            grepl("CD16",group) ~ "CD16+ Monocyte",
            grepl("CD4",group) ~ "CD4+ T cell"
        ),
        direction=case_when(
            grepl("Down",group) ~ "Downregulated",
            grepl("Up",group) ~ "Upregulated"
        )) |> 
    mutate(group=gsub("_.*","",group)) |> 
    ggplot(aes(
        x=group,
        y=fct_reorder2(term_name,paste(group,direction),direction),
        size=n_with_sel,
        color=-log10(p_adjust+1e-300)
    ))+
    geom_point()+
    scale_color_gradient(
        low = "#43C6AC",
        high = "#191654"
    )+
    theme_bw()+
    facet_grid(cell ~ direction,space = "free",scales="free")+
    rotate_x_text(angle=45)+
    scale_x_discrete(labels=c(
        "CD4 T cell RNA"             = expression(CD4^"+" ~ "T cell RNA"),
        "CD4 T cell Isoforms"        = expression(CD4^"+" ~ "T cell Isoforms"),
        "CD4 T cell Isoforms DTU"    = expression(CD4^"+" ~ "T cell Isoforms DTU"),
        
        "CD16 Monocyte RNA"          = expression(CD16^"+" ~ "Monocyte RNA"),
        "CD16 Monocyte Isoforms"     = expression(CD16^"+" ~ "Monocyte Isoforms"),
        "CD16 Monocyte Isoforms DTU" = expression(CD16^"+" ~ "Monocyte Isoforms DTU")
    ))+
    labs(
        x="",
        y="",
        color=expression("-Log"[10]*"P"),
        size="Count"
    )+
    theme(strip.text = element_text(face="bold.italic"))
```
#### Broad GO Modules

```{r}
mod_df <- enr_res_sig |> 
  filter(!grepl("DTU",group)) |> 
    mutate(
        cell=case_when(
            grepl("CD16",group) ~ "CD16+ Monocyte",
            grepl("CD4",group) ~ "CD4+ T cell"
        ),
        direction=case_when(
            grepl("Down",group) ~ "Downregulated",
            grepl("Up",group) ~ "Upregulated"
        ),
        mode=case_when(
            grepl("RNA",group) ~ "RNA",
            grepl("Isoform",group) ~ "Isoform"
        )) |> group_by(parentTerm,mode,cell,direction) |>
  reframe(n=n(),
          median_p=median(p_adjust+
                            min(enr_res_sig$p_adjust[enr_res_sig$p_adjust!=0])),
          median_count=median(n_with_sel),
          score=-log10(median_p)*median_count) |> 
  mutate(
    dir_sign = ifelse(direction == "Upregulated", 1, -1),
    signed_score = score * dir_sign
  )

mod_df |> 
    group_by(parentTerm, cell, direction,mode) %>%
    summarise(total = sum(n)) %>%
    mutate(direction = ifelse(direction=="Upregulated", "Up", "Down")) |> 
    # filter(parentTerm %in% c(mod_df |> 
    #                            group_by(parentTerm) |>
    #                            mutate(total=sum(n)) |>
    #                            ungroup() |> 
    #                            dplyr::select(parentTerm,total) |>
    #                            distinct() |> 
    #                            arrange(desc(total)) |> 
    #                            slice_head(n=50) |>
    #                            pull(parentTerm))) |> 
    filter(parentTerm %in% c(mod_df |>
                                 group_by(parentTerm) |>
                                 mutate(total=sum(n)) |>
                                 ungroup() |>
                                 group_by(cell) |> 
                                 arrange(desc(n)) |>
                                 slice_head(n=10) |>
                                 pull(parentTerm))) |>
    ggplot( aes(x = cell,
                y = fct_reorder(parentTerm,total), 
                fill = total)) +
    geom_tile(alpha=.9) +
    facet_grid(mode ~ direction) +
    scale_fill_gradientn(
        colors = c(
            #"#03001e",
            #"#7303c0",
            #"#ec38bc",
            #"#fdeff9",
            "#3a1c71",
            "#d76d77",
            "#ffaf7b"
        ),
        na.value = "grey80")+
    # scale_fill_gradient(
    #   low="#0f0c29",
    #   #mid = ,
    #   high="#F0C27B"
    # )+
    # scale_fill_viridis_c(option = "B") +
    theme_bw()+
    rotate_x_text(angle=90)+
  scale_x_discrete(labels=c(
        "CD4+ T cell"       = expression(CD4^"+" ~ "T cell"),
        "CD16+ Monocyte"       = expression(CD16^"+" ~ "Monocyte")))+
    theme(strip.text = element_text(face="bold.italic"))+
    labs(
        x="",
        y="",
        fill=expression(-Log[10]*"P*Count")
    )
```
#### * Parent Term Heatmap

```{r}
mod_df |>
    group_by(parentTerm, direction) |>
    mutate(parent_dir_score = sum(abs(signed_score))) |>
    ungroup() |>
    group_by(parentTerm) |>
    mutate(parent_n = sum(n),has_rna="RNA" %in% mode) |>
    ungroup() |>
    filter(parent_n > 30,has_rna==T) |>
    mutate(
        parentTerm = fct_reorder(parentTerm, parent_dir_score)
    ) |> 
    ggplot( aes(x = direction,
                y = parentTerm, 
                fill = signed_score)) +
    geom_tile(alpha=.9) +
    ggh4x::facet_nested(. ~ cell+mode)+
    scale_fill_gradient2(
        low ="blue4",
        mid="white",
        high = "red4",
        midpoint = 0,
        limits=c(-100,100),
        oob=scales::squish
    )+
    theme_bw()+
    rotate_x_text(angle=90)+
    scale_x_discrete(labels=c(
        "CD4+ T cell"       = expression(CD4^"+" ~ "T cell"),
        "CD16+ Monocyte"       = expression(CD16^"+" ~ "Monocyte")))+
    theme(strip.text = element_text(face="bold.italic"),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank())+
    labs(
        x="",
        y="",
        fill="Enrichment Score"
    )
```

#### Gene Heatmap 

```{r}
top_enr_terms <- enr_res_sig |> group_by(group) |> dplyr::count(parentTerm) |> ungroup() |> group_by(group) |> arrange(desc(n)) |> slice_head(n=5) |> pull(parentTerm) |> unique()

gene_tbl <- map(
  crossing(a=unique(enr_res_sig$group),
           b=top_enr_terms) |>
    mutate(combo=paste(a,b,sep=":")) |>
    pull(combo),
  ~ {
    grp=.x |> strsplit(":") |> (\(chr)chr[[1]][1])()
    
    pt=.x |> strsplit(":") |> (\(chr)chr[[1]][2])()
    
    enr_res_sig |>
      filter(group==grp,
             parentTerm==pt) |> 
      separate_rows(sep=", ",ids) |> 
      tabyl(ids) |> 
      mutate(group=grp,
             parentTerm=pt)
  } ) |> 
  bind_rows() |> 
  group_by(group,parentTerm) |> 
  arrange(desc(n)) |> 
  slice_head(n=5) |> 
  mutate(
    cell=case_when(
      grepl("CD16",group) ~ "CD16+ Monocyte",
      grepl("CD4",group) ~ "CD4+ T cell"
    ),
    direction=case_when(
      grepl("Down",group) ~ "Downregulated",
      grepl("Up",group) ~ "Upregulated"
    ),
    mode=case_when(
      grepl("RNA",group) ~ "RNA",
      grepl("Isoform",group) ~ "Isoform"
    ),
    exp_name = gsub("_.*","",group),
    feature=ids) |> 
  left_join(robust_deg_df |> 
              mutate(direction=case_when(
                logFC<0 ~ "Downregulated",
                logFC>0 ~ "Upregulated"
              )),
            by=c("feature"="feature_map",
                 "exp_name"="exp_name",
                 "direction"="direction"))
  
```

#### Heatmap

```{r}
top_enr_terms <- enr_res_sig |> group_by(group) |> dplyr::count(parentTerm) |> ungroup() |> group_by(group) |> arrange(desc(n)) |> slice_head(n=5) |> pull(parentTerm) |> unique()

gene_tbl <- enr_res_sig |>
  filter(parentTerm %in% top_enr_terms) |>
  separate_rows(ids, sep = ", ") |>
  dplyr::count(group, parentTerm, ids, name = "n") |>
  group_by(group, parentTerm) |>
  slice_max(n, n = 3, with_ties = FALSE) |>
  ungroup() |>
  mutate(
    cell = case_when(
      grepl("CD16", group) ~ "CD16+ Monocyte",
      grepl("CD4",  group) ~ "CD4+ T cell"
    ),
    direction = case_when(
      grepl("Down", group) ~ "Downregulated",
      grepl("Up",   group) ~ "Upregulated"
    ),
    mode = case_when(
      grepl("RNA",     group) ~ "RNA",
      grepl("Isoform", group) ~ "Isoform"
    ),
    exp_name = gsub("_.*", "", group),
    feature  = ids
  ) |>
  filter(mode == "RNA") |> 
  left_join(
    robust_deg_df |>
      mutate(
        direction = case_when(
          logFC < 0 ~ "Downregulated",
          logFC > 0 ~ "Upregulated"
        )
      ),
    by = c(
      "feature"   = "feature_map",
      "exp_name"  = "exp_name",
      "direction" = "direction"
    )
  )

gene_tbl |> 
  filter(mode=="RNA") |> 
  mutate(
        parentTerm = fct_reorder2(parentTerm,ids, cell)
    ) |> 
    ggplot(aes(
        y = ids,
        x = cell,
        color = logFC,
        size=-log10(P.Value)
    )) +
    geom_point() +
    facet_grid(parentTerm ~ cell, 
               scales = "free",
               space = "free") +
    scale_color_gradient2(
        low = "blue4",
        mid = "white",
        high = "red4",
        midpoint = 0, 
        limits = c(-10, 10),
        oob = scales::squish
    ) +
    theme_classic()+
  theme(axis.text.y = element_text(face="italic"),
        strip.text.y = element_text(angle=0))+
  rotate_x_text(angle = 45)

```

#### CD4 Barplot

```{r}
#| fig-width: 12
#| fig-height: 6
library(dplyr)
library(ggplot2)
library(stringr)
library(tidytext)

top_cd4_enr_terms <- enr_res_sig |> 
  filter(grepl("CD4",group),
         grepl("RNA",group)) |> 
  mutate(strength=-log10(p_adjust)*n_with_sel) |> 
  group_by(parentTerm) |> 
  reframe(med_str=median(strength),
          n=n()) |> 
  ungroup() |> 
  filter(n>10) |> 
  arrange(desc(med_str)) |> 
  slice_head(n=15) |> 
  pull(parentTerm) |>
  unique()

enr_res_sig |>
    mutate(
        direction = case_when(
            grepl("Up",   group) ~ "Upregulated",
            grepl("Down", group) ~ "Downregulated"
        ),
        cell = case_when(
            grepl("CD4",  group) ~ "CD4+ T cell",
            grepl("CD16", group) ~ "CD16+ Monocyte"
        ),
        mode = case_when(
            grepl("RNA",     group) ~ "RNA",
            grepl("Isoform", group) ~ "Isoform"
        ),
        strength = -log10(p_adjust + 1e-10)
    ) |>
    filter(
        cell == "CD4+ T cell",
        mode == "RNA",
        !is.na(direction),
        !is.na(cell)
    ) |>
    # group_by(cell) |>
    # arrange(desc(strength)) |>
    # slice_head(n = 8) |>
    # ungroup() |>
    filter(parentTerm %in% top_cd4_enr_terms) |> 
    group_by(parentTerm,direction) |>
    reframe(total = median(strength),cell=cell) |>
    ungroup() |>
  distinct() |> 
    mutate(
        parentTerm = str_wrap(parentTerm, width = 60)
    ) |>
    ggplot(aes(
        x = total,
        y = reorder_within(parentTerm, total, cell),
        fill = direction
    )) +
    geom_col(alpha = 0.8) +
    scale_y_reordered() +
    facet_wrap(~ cell, scales = "free", nrow = 1) +
    scale_fill_manual(
        values = c(
            "Upregulated"   = "red4",
            "Downregulated" = "blue4"
        )
    ) +
    geom_vline(
        xintercept = -log10(0.05 + 1e-10),
        linetype = "dashed"
    ) +
    theme_pubr(legend = "bottom") +
    labs(
        x = expression("-Log"[10] * " Adj. P"),
        y = NULL,
        fill = ""
    ) +
    theme(
        strip.text = element_text(face = "bold.italic"),
        axis.text.y = element_text(size = 10)
    )

```
#### CD16 Barplot

```{r}
top_cd16_enr_terms <- enr_res_sig |> 
  filter(grepl("CD16",group)) |> 
  #group_by(group) |> 
  dplyr::count(parentTerm) |> 
  #ungroup() |> 
  #group_by(group) |> 
  arrange(desc(n)) |> 
  slice_head(n=10) |>
  pull(parentTerm) |>
  unique()

enr_res_sig |>
    mutate(
        direction = case_when(
            grepl("Up",   group) ~ "Upregulated",
            grepl("Down", group) ~ "Downregulated"
        ),
        cell = case_when(
            grepl("CD4",  group) ~ "CD4+ T cell",
            grepl("CD16", group) ~ "CD16+ Monocyte"
        ),
        mode = case_when(
            grepl("RNA",     group) ~ "RNA",
            grepl("Isoform", group) ~ "Isoform"
        ),
        strength = -log10(p_adjust + 1e-10)
    ) |>
    filter(
        cell == "CD16+ Monocyte",
        !is.na(direction),
        !is.na(cell)
    ) |>
    # group_by(cell) |>
    # arrange(desc(strength)) |>
    # slice_head(n = 8) |>
    # ungroup() |>
    filter(parentTerm %in% top_cd4_enr_terms) |> 
    group_by(parentTerm) |>
    mutate(total = sum(strength)) |>
    ungroup() |>
    mutate(
        parentTerm = str_wrap(parentTerm, width = 60)
    ) |>
    ggplot(aes(
        x = strength,
        y = reorder_within(parentTerm, total, cell),
        fill = direction
    )) +
    geom_col(alpha = 0.8) +
    scale_y_reordered() +
    facet_wrap(~ cell, scales = "free", nrow = 1) +
    scale_fill_manual(
        values = c(
            "Upregulated"   = "red4",
            "Downregulated" = "blue4"
        )
    ) +
    geom_vline(
        xintercept = -log10(0.05 + 1e-10),
        linetype = "dashed"
    ) +
    theme_pubr(legend = "bottom") +
    labs(
        x = expression("-Log"[10] * " Adj. P"),
        y = NULL,
        fill = ""
    ) +
    theme(
        strip.text = element_text(face = "bold.italic"),
        axis.text.y = element_text(size = 10)
    )
```


#### Try 3

```{r}
mod_df |>
  group_by(parentTerm) |> 
  mutate(parent_n=sum(n)) |> 
  ungroup() |> 
  filter(parent_n>10) |> 
  ggplot( aes(x = cell,
                y = fct_reorder(parentTerm,n), 
                fill = signed_score)) +
    geom_tile(alpha=.9) +
    facet_grid(mode ~ direction) +
    scale_fill_gradient2(
      low ="blue4",
      mid="white",
      high = "red4",
      midpoint = 0,
      limits=c(-200,200),
      oob=scales::squish
    )+
    theme_bw()+
    rotate_x_text(angle=90)+
  scale_x_discrete(labels=c(
        "CD4+ T cell"       = expression(CD4^"+" ~ "T cell"),
        "CD16+ Monocyte"       = expression(CD16^"+" ~ "Monocyte")))+
    theme(strip.text = element_text(face="bold.italic"))+
    labs(
        x="",
        y="",
        fill=expression(-Log[10]*"P*Count")
    )
```

#### Gene Sets

```{r}
a <- map(
  unique(enr_res_sig$group),
         ~{
           enr_res_sig |>
             filter(group==.x) |> 
             separate_rows(sep=", ",ids) |> 
             filter(grepl("adhesion",term_name)) |> 
             pull(ids) |> 
             unique()
           } ) |> 
  set_names(unique(enr_res_sig$group))

gene_tbl <- enr_res_sig |>
    separate_rows(ids, sep = ", ") |>
    mutate(
        direction = case_when(
            grepl("Upregulated", group) ~ "Up",
            grepl("Downregulated", group) ~ "Down"
        ),
        cell_type = case_when(
            grepl("CD4", group) ~ "CD4",
            grepl("CD16", group) ~ "CD16"
        ),
        mode = case_when(
            grepl("Isoforms", group) ~ "Isoform",
            grepl("RNA", group) ~ "RNA"
        )
    )

a <- robust_deg_df |> 
  mutate(direction=ifelse(logFC>0,"Up","Down")) |> 
  mutate(
        cell_type = case_when(
            grepl("CD4", exp_name) ~ "CD4",
            grepl("CD16", exp_name) ~ "CD16"
        ),
        mode = case_when(
            grepl("Isoforms", exp_name) ~ "Isoform",
            grepl("RNA", exp_name) ~ "RNA"
        )
    ) |> 
  group_by(cell_type,mode,feature_map) |> 
  dplyr::count(direction)
  
a <- gene_tbl |> 
  filter(mode=="Isoform") |> 
  group_by(cell_type,ids) |> 
  dplyr::count(direction) |> 
  mutate(consistency=length(unique(direction)))
```

## Deconvolution

### Monaco Decon

```{r}
monaco <- read_tsv("../data/rna_immune_cell_monaco.tsv")

monaco_cd4_ref <- monaco |> 
  filter(`Immune cell` %in% c(
    "Memory CD4 T-cell TFH",
    "Memory CD4 T-cell Th1" ,
    "Memory CD4 T-cell Th1/Th17",
    "Memory CD4 T-cell Th17",
    "Memory CD4 T-cell Th2",
    "naive CD4 T-cell",
    "T-reg",
    "Terminal effector memory CD4 T-cell"
  )) |> 
  mutate(log=log2(nTPM+1)) |> 
  dplyr::select(Gene,`Immune cell`,log) |> 
  pivot_wider(names_from = "Immune cell",
              values_from = "log") |> 
  column_to_rownames("Gene")

# cd4_se <- tidyexposomics:::.update_assay_colData(
#   fev1_fvc_expom,
#   "CD4 T cell RNA") |> 
#   (\(se){
#     assay(se,"log") <- log2(assay(se,"counts")+1);
#     se
#   })() |> 
#   tidybulk::deconvolve_cellularity(
#     .abundance = "log",
#     method="cibersort",
#     reference = monaco_cd4_ref
#   )

cd4_llsr_se <- tidyexposomics:::.update_assay_colData(
  fev1_fvc_expom,
  "CD4 T cell RNA") |>
  (\(se){
    assay(se,"log") <- log2(assay(se,"counts")+1);
    se
  })() |>
  tidybulk::deconvolve_cellularity(
    .abundance = "log",
    method="llsr",
    reference = monaco_cd4_ref
  )

library(immunedeconv)
cd4_epic_se <- tidyexposomics:::.update_assay_colData(
  fev1_fvc_expom,
  "CD4 T cell RNA") |>
  (\(se){
    assay(se,"log") <- log2(assay(se,"counts")+1);
    se
  })() |>
  tidybulk::deconvolve_cellularity(
    .abundance = "log",
    method="epic",
    reference = monaco_cd4_ref
  )

# How I originally saved the data
# saveRDS( tidyexposomics:::.update_assay_colData(
#   fev1_fvc_expom,
#   "CD4 T cell RNA") |>
#   (\(se){
#     assay(se,"log") <- log2(assay(se,"counts")+1);
#     se
#   })(),
#   file="../results/deconvolution/cd4_se.rds")
# 
# saveRDS(monaco_cd4_ref,
#         file="../results/deconvolution/monaco_cd4_ref.rds")

# run on mac bc of error with tidybulk PC handling
decon <- readRDS("../results/deconvolution/decon.rds")


# decon_res <- map(
#   pivot_sample(decon) |> 
#     colnames() |> 
#     (\(chr)chr[grepl("cell|T.reg",chr)])(),
#   ~ {glm( as.formula(paste0(
#     "pftfev1fvc_actual ~ ",
#     .x,
#     " + gli_age + gli_sex + gli_height + fis")),
#     data=pivot_sample(decon)) |>
#       broom::tidy() |>
#       mutate(cell=.x)}) |>
#   bind_rows() |>
#   filter(grepl("cell|T.reg",term)) |> 
#   mutate(
#     cell_clean = dplyr::case_when(
#       term == "Memory.CD4.T.cell.TFH"        ~ "Tfh (memory CD4)",
#       term == "Memory.CD4.T.cell.Th1"        ~ "Th1 (memory CD4)",
#       term == "Memory.CD4.T.cell.Th2"        ~ "Th2 (memory CD4)",
#       term == "Memory.CD4.T.cell.Th17"       ~ "Th17 (memory CD4)",
#       term == "Memory.CD4.T.cell.Th1.Th17"   ~ "Th1/Th17 (memory CD4)",
#       term == "Terminal.effector.memory.CD4.T.cell" ~ "Terminal effector memory CD4",
#       term == "naive.CD4.T.cell"             ~ "Naive CD4",
#       term == "T.reg"                        ~ "Treg",
#       TRUE                                   ~ term
#     )
#   )
```

### CLR Test

```{r}
library(dplyr)
library(purrr)
library(compositions)
library(broom)

ps <- tidybulk::pivot_sample(decon)

cell_cols <- colnames(ps)[grepl("cell|T.reg", colnames(ps))]

cell_mat <- ps |>
  dplyr::select(all_of(cell_cols)) |>
  as.matrix()

# Zero handling (required for CLR)
cell_mat[cell_mat == 0] <- 1e-6

# CLR transform
clr_mat <- compositions::clr(cell_mat)
colnames(clr_mat) <- paste0("clr_", cell_cols)

meta <- ps |>
  dplyr::select(
    pftfev1fvc_actual,
    gli_age,
    gli_sex,
    gli_height,
    fis
  )

model_df <- bind_cols(meta, as.data.frame(clr_mat))

# Fit one model per CLR component
decon_res <- map_dfr(
  colnames(clr_mat),
  function(clr_col) {
    lm(
      as.formula(paste(
        "pftfev1fvc_actual ~",
        clr_col,
        "+ gli_age + gli_sex + gli_height + fis"
      )),
      data = model_df
    ) |>
      tidy() |>
      filter(term == clr_col) |>
      mutate(cell = sub("^clr_", "", clr_col))
  }
)


decon_res <- decon_res |>
  mutate(
    cell_clean = case_when(
      cell == "Memory.CD4.T.cell.TFH"        ~ "Tfh (memory CD4)",
      cell == "Memory.CD4.T.cell.Th1"        ~ "Th1 (memory CD4)",
      cell == "Memory.CD4.T.cell.Th2"        ~ "Th2 (memory CD4)",
      cell == "Memory.CD4.T.cell.Th17"       ~ "Th17 (memory CD4)",
      cell == "Memory.CD4.T.cell.Th1.Th17"   ~ "Th1/Th17 (memory CD4)",
      cell == "Terminal.effector.memory.CD4.T.cell" ~ "Terminal effector memory CD4",
      cell == "naive.CD4.T.cell"             ~ "Naive CD4",
      cell == "T.reg"                        ~ "Treg",
      TRUE                                   ~ cell
    )
  )

```

### Bubble Plot

```{r}
decon_res |> 
  ggplot(aes(
    x = estimate,
    y = fct_reorder(cell_clean,estimate),
    color = estimate,
    size = -log10(p.value)
  )) +
  geom_point(alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_bw() +
  theme(strip.text.y = element_text(angle = 0,
                                  face="bold.italic"),
        strip.text.x = element_text(face="bold.italic"),
        axis.text.y = element_text(face = "italic")) +
  scale_color_gradient2(
    low = "blue4",
    mid = "grey85",
    high = "red4",
    midpoint = 0,
    # limits=c(-.3,.3),
    # oob=scales::squish
  ) +
  labs(
    x = "Effect size",
    y = NULL,
    color = "Estimate",
    size = expression("-Log"[10] * "P")
  )
```

### Forest Plot

```{r}

decon_res |> 
  ggplot(aes(x = estimate, 
               y = fct_reorder(cell_clean,estimate),
               color = estimate,
               fill = estimate)) +
    geom_point(shape = 18, 
               size = 3, 
               stroke = 0.8) +
    geom_errorbarh(aes(xmin = estimate - 1.96 * std.error,
                       xmax = estimate + 1.96 * std.error),
                   height = 0.15) +
    scale_color_gradient2(
        low="blue4",
        mid="grey85",
        high="red4",
        midpoint = 0
    )+
    scale_fill_gradient2(
        low="blue4",
        mid="grey85",
        high="red4",
        midpoint = 0
    )+
    geom_vline(xintercept = 0, linetype = "dashed") +
    labs(
        title = NULL,
        x = paste("Effect size",
                  "(CLR log-ratio, 95% CI)",
                  sep = "\n"),
        y = "",
        fill ="Effect size",
        color ="Effect size"
    ) +
    theme_bw(base_size = 12) +
    theme(panel.grid.minor = element_blank(),
          axis.text.x = element_text(size=10),
          strip.text.y = element_text(angle=0,
                                      face="bold.italic"),
          axis.text.y = element_text(face = "italic"),
          legend.position = "none")
```

## Module Scores

### Correlation with CD4 DEGs

```{r}

top_cd4_enr_terms <- mod_df |>
  group_by(parentTerm, direction) |>
  mutate(parent_dir_score = sum(abs(signed_score))) |>
  ungroup() |>
  group_by(parentTerm) |>
  mutate(parent_n = sum(n),has_rna="RNA" %in% mode) |>
  ungroup() |>
  #filter(parent_n > 30,has_rna==T) |>
  pull(parentTerm) |>
  unique()

top_cd4_enr_terms <- top_cd4_enr_terms[!is.na(top_cd4_enr_terms)]

cd4_rna_features <- map(
  top_cd4_enr_terms,
  ~{
    enr_res_sig |> 
      filter(parentTerm %in% .x) |> 
      separate_rows(sep=", ",ids) |> 
      pull(ids) |> 
      unique()}) |> 
  set_names(top_cd4_enr_terms)

cd4_se <- tidyexposomics:::.update_assay_colData(
  fev1_fvc_expom,
  "CD4 T cell RNA") |> 
  (\(se){
    assay(se,"log") <- log2(assay(se,"counts")+1);
    se
  })()


cd4_mat <- cd4_se |>
  pivot_se() |>
  dplyr::select(.sample, gene_name, log) |>
  pivot_wider(
    names_from  = gene_name,
    values_from = log,
    values_fn   = mean
  ) |>
  column_to_rownames(".sample") |>
  as.matrix()

cd4_gsva_par <- gsvaParam(
  t(cd4_mat),            # matrix: samples x genes
  cd4_rna_features,         # list of gene sets
  #method = "ssgsea",   # or "gsva", "zscore", "plage"
  kcdf = "Gaussian",   # use "Gaussian" for log-transformed RNA-seq data
  absRanking = TRUE    # use original ssGSEA normalization
)

cd4_gsva_df <- gsva(cd4_gsva_par) |> 
  t() |> 
  as.data.frame() |> 
  rownames_to_column(".sample") |> 
  left_join(
    cd4_se |>
      tidybulk::pivot_sample(),
    by = ".sample"
  ) 

# Assume this gives: .sample, Th1, Th2, Treg, etc.
th_profiles <- decon |> 
    tidybulk::pivot_sample() |>
    dplyr::select(.sample, contains("cell"), contains("T.reg"))

# Join with GSVA scores
cd4_merged <- cd4_gsva_df |>
  dplyr::select(names(cd4_rna_features),.sample) |> 
    inner_join(th_profiles, by = ".sample") |> 
  inner_join(tidybulk::pivot_sample(cd4_se) |> 
               dplyr::select(.sample,
                             gli_age,
                             gli_sex,
                             gli_height,
                             fis),
             by=".sample")

# Get list of GSVA terms and Th features
gsva_cols <- cd4_gsva_df |> 
  dplyr::select(names(cd4_rna_features),.sample) |> 
  colnames() |> 
  setdiff(".sample")
th_cols   <- th_profiles |>
  colnames() |> 
  setdiff(".sample")

# Correlation per GSVA x Th combination
# library(purrr)
# 
# gsva_cor_res <- crossing(gsva = gsva_cols, th = th_cols) |>
#     mutate(
#         cor = map2_dbl(gsva,
#                        th, ~ cor(cd4_merged[[.x]],
#                                  cd4_merged[[.y]],
#                                  method = "spearman")),
#         pval = map2_dbl(gsva,
#                         th, ~ cor.test(cd4_merged[[.x]], 
#                                        cd4_merged[[.y]],
#                                        method = "spearman")$p.value)
#     )

gsva_assoc_res <-
  expand_grid(
    gsva = gsva_cols,
    th   = th_cols
  ) |>
  mutate(
    model = map2(
      gsva,
      th,
      ~ {
        fml <- as.formula(paste(
          "`",
          .x,
          "`~ `",
          .y,
          "`+ gli_age + gli_sex + gli_height + fis"
        ,sep=""))
        lm(fml, data = cd4_merged)
      }
    )
  ) |>
  mutate(
    tidy = map(model, tidy)
  ) |>
  unnest(tidy) |>
  filter(term == th) |>
  select(
    gsva,
    th,
    estimate,
    std.error,
    statistic,
    p.value
  ) |> 
  mutate(
    cell_clean = case_when(
      th == "Memory.CD4.T.cell.TFH"               ~ "Tfh (memory CD4)",
      th == "Memory.CD4.T.cell.Th1"               ~ "Th1 (memory CD4)",
      th == "Memory.CD4.T.cell.Th2"               ~ "Th2 (memory CD4)",
      th == "Memory.CD4.T.cell.Th17"              ~ "Th17 (memory CD4)",
      th == "Memory.CD4.T.cell.Th1.Th17"          ~ "Th1/Th17 (memory CD4)",
      th == "Terminal.effector.memory.CD4.T.cell" ~ "Terminal effector memory CD4",
      th == "naive.CD4.T.cell"                    ~ "Naive CD4",
      th == "T.reg"                               ~ "Treg",
      TRUE                                          ~ th
    )
  )
```

### Stability

```{r}
stability_df <- cd4_merged |>
  dplyr::select(
    all_of(gsva_cols),
    all_of(th_cols),
    gli_age, gli_sex, gli_height, fis
  )

# colnames(stability_df) <- gsub("\\ |\\-","_",colnames(stability_df))
# 
# gsva_cols <- gsub("\\ |\\-","_",gsva_cols)

gsva_th_stability_res <- future_map(
  gsva_cols,
  ~ {
    run_exwas_sensitivity(
      df = stability_df,
      exposures = th_cols,
      outcome = .x,
      covariates = c("gli_age", "gli_sex", "gli_height", "fis"),
      covariates_to_remove = c("gli_age", "gli_sex", "gli_height", "fis"),
      family = gaussian(),
      bootstrap_n = 100
    )$sensitivity_df |>
      mutate(
        pathway = .x
      )
  },.options = furrr_options(seed = 123)
) |> 
  bind_rows()

# gsva_th_stability_res <-
#   # expand_grid(
#   #   gsva = gsva_cols,
#   #   th   = th_cols
#   # ) 
#   # examine the significant associations
# gsva_assoc_res |> 
#   filter(p.value<0.05) |> 
#   dplyr::select(gsva,th) |>
#   mutate(
#     sens = map2(
#       gsva,
#       th,
#       ~ {
#         run_exwas_sensitivity(
#           df = stability_df,
#           exposures = .y,               # Th score
#           outcome   = .x,               # GSVA score
#           covariates = c(
#             "gli_age", "gli_sex",
#             "gli_height", "fis"
#           ),
#           covariates_to_remove = c(
#             "gli_age", "gli_sex",
#             "gli_height", "fis"
#           ),
#           family = gaussian(),
#           bootstrap_n = 100
#         )$sensitivity_df |>
#           mutate(
#             gsva = .x,
#             th   = .y
#           )
#       }
#     )
#   ) |>
#   select(-sens) |>
#   unnest(sens)


gsva_th_stability_df <-
  gsva_th_stability_res |>
  dplyr::rename(th=exposure) |> 
  dplyr::rename(gsva=pathway) |> 
  mutate(
    signif = (p_value <= 0.05),
    sign   = sign(estimate)
  ) |>
  group_by(gsva, th) |>
  summarise(
    n_tests = n(),
    n_signif = sum(signif, na.rm = TRUE),
    prop_signif = n_signif / n_tests,
    median_effect = median(estimate, na.rm = TRUE),
    median_p = median(p_value, na.rm = TRUE),
    sd_effect = sd(estimate, na.rm = TRUE),
    sign_consistency =
      ifelse(
        all(is.na(sign)),
        NA_real_,
        abs(sum(sign, na.rm = TRUE)) / n_tests
      ),
    stability_score = prop_signif * sign_consistency,
    .groups = "drop"
  ) |> 
  inner_join(gsva_assoc_res,
             by=c(
               "th"="th",
               "gsva"="gsva"
             ))

```

### * Barplot

```{r}
gsva_th_stability_df |>
    filter(p.value<0.05,
           sign_consistency>0.7) |>
    group_by(cell_clean) |> 
    reframe(med=median(estimate)) |>
    ggplot(aes(
        x = med,
        y = fct_reorder(cell_clean,med),
        fill = med
    )) +
    geom_col(color = "black", 
             linewidth = 0.005) +
    theme_bw() +
    scale_fill_gradient2(
        low ="blue4",
        mid="white",
        high = "red4",
        midpoint = 0
    )+scale_x_continuous(breaks = c(-1,-.5,.1))+
    labs(x = "",
         y = "", 
         fill = "Median Estimate") +
    theme(
        strip.text.x = element_text(
            face = "bold.italic"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor   = element_blank(),
        legend.position = "right",
        legend.title.position = "top"
    )
```

### * Barplot Thresholds

```{r}
pmap(
  expand_grid(
    consistency = c(0.3, 0.5, 0.7, 0.9),
    p_cutoff    = c(0.1, 0.05)
),
    \(consistency,p_cutoff){
      gsva_th_stability_df |>
    filter(p.value<p_cutoff,
           sign_consistency>consistency) |>
    group_by(cell_clean) |> 
    reframe(med=median(estimate),
            n=n(),
            cell_clean=paste(cell_clean," (",n,")",sep="")) |> 
        distinct() |> 
        (\(df){
        df |> 
            ggplot(aes(
                x = med,
                y = fct_reorder(cell_clean,med),
                fill = med
            )) +
            geom_col(color = "black", 
                     linewidth = 0.005) +
            theme_bw() +
            scale_fill_gradient2(
                low ="blue4",
                mid="white",
                high = "red4",
                midpoint = 0
            )+
        scale_x_continuous(breaks = c(-1,-.5,.1))+
            labs(x = "",title="Th-Airway Obstruction",
                 y = "", 
                 caption=paste("P <",p_cutoff, "& Effect Consistency >",consistency*100,"%"),
                 fill = "Median Estimate") +
            theme(
                strip.text.x = element_text(
                    face = "bold.italic"),
                panel.grid.major.x = element_blank(),
                panel.grid.minor   = element_blank(),
                legend.position = "right",
                legend.title.position = "top",
                plot.title = element_text(
                  size=12
                )
            ) })()
    }) |> 
  wrap_plots(ncol = 2)

gsva_th_stability_df |>
    filter(p.value<0.05,
           sign_consistency>0.9) |>
    group_by(cell_clean) |> 
    reframe(med=median(estimate)) |> (\(df){
        df |> 
            ggplot(aes(
                x = med,
                y = fct_reorder(cell_clean,med),
                fill = med
            )) +
            geom_col(color = "black", 
                     linewidth = 0.005) +
            theme_bw() +
            scale_fill_gradient2(
                low ="blue4",
                mid="white",
                high = "red4",
                midpoint = 0
            )+scale_x_continuous(breaks = c(-1,-.5,.1))+
            labs(x = "",title="Th-Airway Obstruction Pathway", 
                 subtitle = paste("N:", nrow(gsva_th_stability_df |>
                                                 filter(p.value<0.05,
                                                        sign_consistency>0.9))," Exposures"),
                 y = "", caption="P < 0..05 & Effect Consistency > 90%",
                 fill = "Median Estimate") +
            theme(
                strip.text.x = element_text(
                    face = "bold.italic"),
                panel.grid.major.x = element_blank(),
                panel.grid.minor   = element_blank(),
                legend.position = "right",
                legend.title.position = "top"
            ) })()
    
```


### Circos

```{r}
gsva_cor_res |> 
  mutate(th=gsub("\\."," ",th)) |> 
    mutate(combo=paste(th,gsva)) |> 
    group_by(combo) |> 
    reframe(med_cor=median(cor),th,gsva) |>
    distinct() |> 
    arrange(med_cor) |> 
    dplyr::select(th,gsva,med_cor) |> 
    tidygraph::as_tbl_graph() |> 
    ggraph::ggraph(
        layout = "linear",
        circular = TRUE
    ) +
    ggraph::geom_edge_arc(aes(
        color = med_cor,
        width = abs(med_cor)*10
    )) +
    ggraph::geom_node_point(shape = 21,
                            size = 4,
                            alpha = 0.8
    ) +
    ggraph::geom_node_text(
        aes(
            label = stringr::str_wrap(name, width = 15),
            x = 1.1 * x,
            y = 1.1 * y,
            angle = ggraph::node_angle(x, y)
        ),
        hjust = "outward",
        fontface = "bold.italic",
        lineheight = 0.9,
        check_overlap = TRUE
    ) +
    ggraph::theme_graph() +
    ggraph::scale_edge_color_gradient2(
        high = "#C33764",
        mid = "white",
        low = "#1D2671",
        midpoint = 0,
        guide = ggraph::guide_edge_colorbar()
    ) +
    ggplot2::coord_fixed(xlim = c(-2, 2), ylim = c(-2, 2)) +
    ggplot2::guides(edge_width = "none") +
    ggplot2::labs(
        edge_color = "Median Correlation"
    )+
    theme(legend.position = "bottom",
          legend.direction = "horizontal")
```

### Heatmap

```{r}
#| fig-width: 10
#| fig-height: 10
library(dplyr)
library(ggplot2)
library(forcats)

gsva_cor_res |>
  #filter(pval<0.05) |> 
  mutate(
    th = case_when(
      th == "Memory.CD4.T.cell.TFH"        ~ "Tfh (memory CD4)",
      th == "Memory.CD4.T.cell.Th1"        ~ "Th1 (memory CD4)",
      th == "Memory.CD4.T.cell.Th2"        ~ "Th2 (memory CD4)",
      th == "Memory.CD4.T.cell.Th17"       ~ "Th17 (memory CD4)",
      th == "Memory.CD4.T.cell.Th1.Th17"   ~ "Th1/Th17 (memory CD4)",
      th == "Terminal.effector.memory.CD4.T.cell" ~ "Terminal effector memory CD4",
      th == "naive.CD4.T.cell"             ~ "Naive CD4",
      th == "T.reg"                        ~ "Treg",
      TRUE                                   ~ th
    )
  ) |> 
    group_by(th, gsva) |>
    summarise(
        med_cor = median(cor, na.rm = TRUE),
        .groups = "drop"
    ) |>
  mutate(th=fct_reorder(th,med_cor),
         gsva=fct_reorder(gsva,med_cor)) |> 
    ggplot(aes(
        x = gsva,
        y = th,
        fill = med_cor
    )) +
    geom_tile(color = "grey90") +
    scale_fill_gradient2(
        low = "#1D2671",
        mid = "white",
        high = "#C33764",
        midpoint = 0
    ) +
    theme_classic() +
    theme(
        axis.text.x = element_text(angle = 65, hjust = 1),
        axis.text.y = element_text(),
        legend.position = "bottom"
    ) +
    labs(
        x = "",
        y = "",
        fill = "Median Correlation"
    )

```

### Gene sets

```{r}
library(msigdbr)
focused_set <- list(
  "Monocyte Chemotaxis"=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_MONOCYTE_CHEMOTAXIS",gs_name)) |> 
  pull(gene_symbol),
  
"T cell Chemotaxis"=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_T_CELL_CHEMOTAXIS",gs_name)) |> 
  pull(gene_symbol),
"Cytokine Production"=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_CYTOKINE_PRODUCTION",gs_name)) |> 
  pull(gene_symbol),
"Type I Interferon Production"=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_TYPE_I_INTERFERON_PRODUCTION",gs_name)) |> 
  pull(gene_symbol),
"Type II Interferon Production" =msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_TYPE_II_INTERFERON_PRODUCTION",gs_name)) |> 
  pull(gene_symbol),
"Response to Type II Interferon"=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_RESPONSE_TO_TYPE_II_INTERFERON",gs_name)) |> 
  pull(gene_symbol),
"Response to Type I Interferon"=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_RESPONSE_TO_TYPE_I_INTERFERON",gs_name)) |> 
  pull(gene_symbol),
"T-cell-Mediated Cytoxicity"=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_T_CELL_MEDIATED_CYTOTOXICITY",gs_name)) |> 
  pull(gene_symbol),
"T-cell Proliferation"=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_T_CELL_PROLIFERATION",gs_name)) |> 
  pull(gene_symbol),
"Leukocyte Proliferation"=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_LEUKOCYTE_PROLIFERATION",gs_name)) |> 
  pull(gene_symbol),
"Cell-Cell Adhesion"=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_CELL_CELL_ADHESION",gs_name)) |> 
  pull(gene_symbol),
"Cell-Endothelial Adhesion"=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_LEUKOCYTE_ADHESION_TO_VASCULAR_ENDOTHELIAL_CELL",gs_name)) |> 
  pull(gene_symbol),
"Cell-Cell Adhesion via Integrin"=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_CELL_CELL_ADHESION_MEDIATED_BY_INTEGRIN",gs_name)) |> 
  pull(gene_symbol),
"Antigen Receptor Mediated Signaling"=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_ANTIGEN_RECEPTOR_MEDIATED_SIGNALING_PATHWAY",gs_name)) |> 
  pull(gene_symbol),
"Antigen Processing and Presentation via MHCII"=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_ANTIGEN_PROCESSING_AND_PRESENTATION_OF_EXOGENOUS_PEPTIDE_ANTIGEN_VIA_MHC_CLASS_II",gs_name)) |> 
  pull(gene_symbol),
"Oxidative Response"=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_RESPONSE_TO_OXIDATIVE_STRESS",gs_name)) |> 
  pull(gene_symbol),
"Endoplasmic Reticulum Stress"=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_RESPONSE_TO_ENDOPLASMIC_RETICULUM_STRESS",gs_name)) |> 
  pull(gene_symbol),
"Integrated Stress Response"=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_INTEGRATED_STRESS_RESPONSE_SIGNALING",gs_name)) |> 
  pull(gene_symbol),
"Stress Granule Assembly"=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_STRESS_GRANULE_ASSEMBLY",gs_name)) |> 
  pull(gene_symbol)
)

# TODO:
# cell proliferation, adhesion, type A/B interferons, antigen presentation, stress pathways (UPR, misfolded, stress)
```

### PC Dfs

```{r}
cd4_se <- tidyexposomics:::.update_assay_colData(
  fev1_fvc_expom,
  "CD4 T cell RNA") |> 
  (\(se){
    assay(se,"log") <- log2(assay(se,"counts")+1);
    se
  })()


cd16_se <- tidyexposomics:::.update_assay_colData(
  fev1_fvc_expom,
  "CD16 Monocyte RNA") |> 
  (\(se){
    assay(se,"log") <- log2(assay(se,"counts")+1);
    se
  })()



cd4_pc_df <- map2_dfr(
  focused_set, names(focused_set),
  ~ cd4_se |> 
    pivot_se() |>
      filter(gene_name %in% .x) |>
      dplyr::select(.sample, gene_name, log) |>
      pivot_wider(names_from = gene_name,
                  values_from = log,
                  values_fn = mean) |>
      column_to_rownames(".sample") |>
      as.matrix() |>
      {\(mat) {
        if (ncol(mat) < 2) return(NULL)
        tibble(
          .sample = rownames(mat),
          cd4_pc1 = prcomp(mat, center = TRUE, scale. = TRUE)$x[, 1],
          set = .y
        )
      }}()
) 

cd16_pc_df <- map2_dfr(
  focused_set, names(focused_set),
  ~ cd16_se |> 
    pivot_se() |>
      filter(gene_name %in% .x) |>
      dplyr::select(.sample, gene_name, log) |>
      pivot_wider(names_from = gene_name,
                  values_from = log,
                  values_fn = mean) |>
      column_to_rownames(".sample") |>
      as.matrix() |>
      {\(mat) {
        if (ncol(mat) < 2) return(NULL)
        tibble(
          .sample = rownames(mat),
          cd16_pc1 = prcomp(mat, center = TRUE, scale. = TRUE)$x[, 1],
          set = .y
        )
      }}()
) 

pc_df <- inner_join(
  cd4_pc_df,
  cd16_pc_df,
  by = c(".sample"=".sample","set"="set")
) |>
  left_join(
    cd4_se |>
      pivot_sample(),
    by = ".sample"
  ) 
```

### GSVA Dfs

```{r}
library(GSVA)
cd4_mat <- cd4_se |>
  pivot_se() |>
  dplyr::select(.sample, gene_name, log) |>
  pivot_wider(
    names_from  = gene_name,
    values_from = log,
    values_fn   = mean
  ) |>
  column_to_rownames(".sample") |>
  as.matrix()

cd16_mat <- cd16_se |>
  pivot_se() |>
  dplyr::select(.sample, gene_name, log) |>
  pivot_wider(
    names_from  = gene_name,
    values_from = log,
    values_fn   = mean
  ) |>
  column_to_rownames(".sample") |>
  as.matrix()

cd4_gsva_par <- gsvaParam(
  t(cd4_mat),            # matrix: samples x genes
  focused_set,         # list of gene sets
  #method = "ssgsea",   # or "gsva", "zscore", "plage"
  kcdf = "Gaussian",   # use "Gaussian" for log-transformed RNA-seq data
  absRanking = TRUE    # use original ssGSEA normalization
)

cd16_gsva_par <- gsvaParam(
  t(cd16_mat),            # matrix: samples x genes
  focused_set,         # list of gene sets
  #method = "ssgsea",   # or "gsva", "zscore", "plage"
  kcdf = "Gaussian",   # use "Gaussian" for log-transformed RNA-seq data
  absRanking = TRUE    # use original ssGSEA normalization
)

cd4_gsva_df <- gsva(cd4_gsva_par) |> 
  t() |> 
  as.data.frame() |> 
  rownames_to_column(".sample") |> 
  left_join(
    cd4_se |>
      pivot_sample(),
    by = ".sample"
  ) 

cd16_gsva_df <- gsva(cd16_gsva_par) |> 
  t() |> 
  as.data.frame() |> 
  rownames_to_column(".sample") |> 
  left_join(
    cd16_se |>
      pivot_sample(),
    by = ".sample"
  ) 

```

### GSVA Testing

```{r}
cd4_gsva_res <- map(names(focused_set),~{
    fit <- glm(
      as.formula(paste("pftfev1fvc_actual~ `",.x, "`+ gli_age + gli_sex + gli_height + fis",sep = "")), data = cd4_gsva_df)
    tidy(fit) 
      # filter(term==paste("`",.x,"`",sep = ""))
  }) |> 
  bind_rows() |> 
  mutate(cell="CD4",
         term=gsub("`","",term))

cd16_gsva_res <- map(names(focused_set),~{
    fit <- glm(
      as.formula(paste("pftfev1fvc_actual~ `",.x, "`+ gli_age + gli_sex + gli_height + fis",sep = "")), data = cd16_gsva_df)
    tidy(fit) 
      #filter(term==paste("`",.x,"`",sep = ""))
  }) |> 
  bind_rows() |> 
  mutate(cell="CD16",
         term=gsub("`","",term)) 

gsva_res <- bind_rows(
  cd4_gsva_res,
  cd16_gsva_res
)
```

### PC Testing

```{r}
# For each set: avg_vst ~ DHT + (1 | cell_line)
cd4_pc_res <- pc_df |>
  group_by(set) |>
  group_modify(~{
    fit <- glm(pftfev1fvc_actual~ cd4_pc1 + gli_age + gli_sex + gli_height + fis, data = .x)
    tidy(fit) 
  }) |>
  ungroup() |> 
  filter(term=="cd4_pc1")

cd16_pc_res <- pc_df |>
  group_by(set) |>
  group_modify(~{
    fit <- glm(pftfev1fvc_actual ~ cd16_pc1 + gli_age + gli_sex + gli_height + fis, data = .x)
    tidy(fit) 
  }) |>
  ungroup() |> 
  filter(term=="cd16_pc1")

module_res <- bind_rows(cd4_pc_res,cd16_pc_res)
```

## Broad GSVA Sets

```{r}
library(GSVA)
cd4_se <- tidyexposomics:::.update_assay_colData(
  fev1_fvc_expom,
  "CD4 T cell RNA") |> 
  (\(se){
    assay(se,"log") <- log2(assay(se,"counts")+1);
    se
  })()


cd16_se <- tidyexposomics:::.update_assay_colData(
  fev1_fvc_expom,
  "CD16 Monocyte RNA") |> 
  (\(se){
    assay(se,"log") <- log2(assay(se,"counts")+1);
    se
  })()

cd4_mat <- cd4_se |>
  pivot_se() |>
  dplyr::select(.sample, gene_name, log) |>
  pivot_wider(
    names_from  = gene_name,
    values_from = log,
    values_fn   = mean
  ) |>
  column_to_rownames(".sample") |>
  as.matrix()

cd16_mat <- cd16_se |>
  pivot_se() |>
  dplyr::select(.sample, gene_name, log) |>
  pivot_wider(
    names_from  = gene_name,
    values_from = log,
    values_fn   = mean
  ) |>
  column_to_rownames(".sample") |>
  as.matrix()

go_df <- msigdbr::msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
    filter(grepl("ANTIGEN|ADHESION|INTERFERON|STRESS|UNFOLDED|PROLIFERATION|CELL_CYCLE|CELL_DIVISION|DNA_DAMAGE|INTERLEUKIN|CYTOKINE|CHEMOKINE|MITOCHONDRIA|OXIDATIVE|ATP|APOPTOSIS|CELL_DEATH",gs_name)) |> 
  filter(grepl("^GO",gs_name))

focused_set <- map(unique(go_df$gs_name),
                   ~{go_df |> 
                       filter(gs_name %in% .x) |>
                       pull(gene_symbol)}) |>
  set_names(unique(go_df$gs_name))

# focused_set <- map(unique(enr_res_sig$parentTerm),
#                    ~{
#                      enr_res_sig |> 
#                        filter(parentTerm==.x) |> 
#                        separate_rows(sep=", ",ids) |> 
#                        pull(ids) |> 
#                        unique()
#                    }) |>
#   set_names(unique(enr_res_sig$parentTerm))



focused_set <- focused_set[lengths(focused_set)>20]


cd4_gsva_par <- gsvaParam(
  t(cd4_mat),            # matrix: samples x genes
  focused_set,         # list of gene sets
  kcdf = "Gaussian",   # use "Gaussian" for log-transformed RNA-seq data
  absRanking = TRUE    # use original ssGSEA normalization
)

cd16_gsva_par <- gsvaParam(
  t(cd16_mat),            # matrix: samples x genes
  focused_set,         # list of gene sets
  kcdf = "Gaussian",   # use "Gaussian" for log-transformed RNA-seq data
  absRanking = TRUE    # use original ssGSEA normalization
)

cd4_gsva_df <- gsva(cd4_gsva_par) |> 
  t() |> 
  as.data.frame() |> 
  rownames_to_column(".sample") |> 
  left_join(
    cd4_se |>
      tidybulk::pivot_sample(),
    by = ".sample"
  ) 

cd16_gsva_df <- gsva(cd16_gsva_par) |> 
  t() |> 
  as.data.frame() |> 
  rownames_to_column(".sample") |> 
  left_join(
    cd16_se |>
      tidybulk::pivot_sample(),
    by = ".sample"
  ) 

cd4_gsva_res <- map(names(focused_set),~{
    fit <- glm(
      as.formula(paste("pftfev1fvc_actual~ `",.x, "`+ gli_age + gli_sex + gli_height + fis",sep = "")), data = cd4_gsva_df)
    tidy(fit) |> 
      filter(grepl("GO",term))
  }) |> 
  bind_rows() |> 
  mutate(cell="CD4",
         term=gsub("`","",term))

cd16_gsva_res <- map(names(focused_set),~{
    fit <- glm(
      as.formula(paste("pftfev1fvc_actual~ `",.x, "`+ gli_age + gli_sex + gli_height + fis",sep = "")), data = cd16_gsva_df)
    tidy(fit)  |> 
      filter(grepl("GO",term))
  }) |> 
  bind_rows() |> 
  mutate(cell="CD16",
         term=gsub("`","",term)) 

gsva_res <- bind_rows(
  cd4_gsva_res,
  cd16_gsva_res
) |> 
  mutate(category = case_when(
  grepl("ANTIGEN", term) ~ "Antigen Presentation",
  grepl("ADHESION", term) ~ "Cell Adhesion",
  grepl("INTERLEUKIN|CYTOKINE|CHEMOKINE", term) ~ "Cytokine Signaling",
  grepl("INTERFERON", term) ~ "Interferon Response",
  grepl("PROLIFERATION|CELL_CYCLE|CELL_DIVISION", term) ~ "Proliferation",
  grepl("DNA_DAMAGE", term) ~ "DNA Damage Response",
  grepl("STRESS|UNFOLDED", term) ~ "ER/Stress Response",
  grepl("MITOCHONDRIA|OXIDATIVE|ATP", term) ~ "Mitochondrial Metabolism",
  grepl("APOPTOSIS|CELL_DEATH", term) ~ "Apoptosis",
  TRUE ~ "Other"
))

  # mutate(category = case_when(
  #   str_detect(term, regex("ANTIGEN", ignore_case = TRUE))      ~ "Antigen Presentation",
  #   str_detect(term, regex("ADHESION", ignore_case = TRUE))     ~ "Cell Adhesion",
  #   str_detect(term, regex("INTERFERON", ignore_case = TRUE))   ~ "Interferon Response",
  #   str_detect(term, regex("STRESS|UNFOLDED", ignore_case = TRUE)) ~ "Stress Response",
  #   str_detect(term, regex("PROLIFERATION", ignore_case = TRUE))~ "Proliferation",
  #   str_detect(term, regex("DNA_DAMAGE", ignore_case = TRUE))   ~ "DNA Damage",
  #   str_detect(term, regex("INTERLEUKIN", ignore_case = TRUE))  ~ "Interleukin Signaling",
  #   str_detect(term, regex("CYTOKINE", ignore_case = TRUE))     ~ "Cytokine Signaling",
  #   str_detect(term, regex("CHEMOKINE", ignore_case = TRUE))    ~ "Chemokine Signaling",
  #   TRUE ~ "Other"
  # ))
```

### Median Effect Size

```{r}
gsva_res |>
    group_by(cell, category) |>
    summarize(
        median_est = median(estimate),
        median_logp = median(-log10(p.value)),
        n_terms = n(),
        .groups = "drop"
    ) |>
    ggplot(aes(
        x = cell,
        y = fct_reorder2(category, cell,median_est),
        fill = median_est
    )) +
    geom_tile(color = "white") +
    scale_fill_gradient2(
        low = "blue4",
        mid = "grey85",
        high = "red4",
        midpoint = 0,
        name = "Median Estimate"
    ) +
    theme_bw() +
    labs(
        x = "",
        y = ""
    ) +
    scale_x_discrete(labels=c(
        "CD4"       = expression(CD4^"+" ~ "T cell"),
        "CD16"       = expression(CD16^"+" ~ "Monocyte")))+
    theme(
        axis.text.y = element_text(size = 10),
        axis.text.x = element_text(size = 10, angle = 90, hjust = 1,vjust=.35),
        panel.grid = element_blank()
    )
```


### Correlation

```{r}
# Assume this gives: .sample, Th1, Th2, Treg, etc.
th_profiles <- decon |> 
    tidybulk::pivot_sample() |>
    dplyr::select(.sample, contains("cell"), contains("T.reg"))

# Join with GSVA scores
cd4_merged <- cd4_gsva_df |> dplyr::select(starts_with("GO"),.sample) |> 
    inner_join(th_profiles, by = ".sample")

# Get list of GSVA terms and Th features
gsva_cols <- cd4_gsva_df |> 
  dplyr::select(starts_with("GO"),.sample) |> 
  colnames() |> 
  setdiff(".sample")
th_cols   <- th_profiles |>
  colnames() |> 
  setdiff(".sample")

# Correlation per GSVA × Th combination
library(purrr)

gsva_cor_res <- crossing(gsva = gsva_cols, th = th_cols) |>
    mutate(
        cor = map2_dbl(gsva,
                       th, ~ cor(cd4_merged[[.x]],
                                 cd4_merged[[.y]],
                                 method = "spearman")),
        pval = map2_dbl(gsva,
                        th, ~ cor.test(cd4_merged[[.x]], 
                                       cd4_merged[[.y]],
                                       method = "spearman")$p.value)
    ) |>
    mutate(padj = p.adjust(pval, method = "fdr")) |> 
  mutate(th=gsub("\\."," ",th)) |> 
  mutate(category = case_when(
  grepl("ANTIGEN", gsva) ~ "Antigen Presentation",
  grepl("ADHESION", gsva) ~ "Cell Adhesion",
  grepl("INTERLEUKIN|CYTOKINE|CHEMOKINE", gsva) ~ "Cytokine Signaling",
  grepl("INTERFERON", gsva) ~ "Interferon Response",
  grepl("PROLIFERATION|CELL_CYCLE|CELL_DIVISION", gsva) ~ "Proliferation",
  grepl("DNA_DAMAGE", gsva) ~ "DNA Damage Response",
  grepl("STRESS|UNFOLDED", gsva) ~ "ER/Stress Response",
  grepl("MITOCHONDRIA|OXIDATIVE|ATP", gsva) ~ "Mitochondrial Metabolism",
  grepl("APOPTOSIS|CELL_DEATH", gsva) ~ "Apoptosis",
  TRUE ~ "Other"
))
```

### Circos

```{r}
gsva_cor_res |> 
    mutate(combo=paste(th,category)) |> 
    group_by(combo) |> 
    reframe(med_cor=median(cor),th,category) |>
    distinct() |> 
    arrange(med_cor) |> 
    dplyr::select(th,category,med_cor) |> 
    tidygraph::as_tbl_graph() |> 
    ggraph::ggraph(
        layout = "linear",
        circular = TRUE
    ) +
    ggraph::geom_edge_arc(aes(
        color = med_cor,
        width = abs(med_cor)*20
    )) +
    ggraph::geom_node_point(shape = 21,
                            size = 4,
                            alpha = 0.8
    ) +
    ggraph::geom_node_text(
        aes(
            label = stringr::str_wrap(name, width = 15),
            x = 1.1 * x,
            y = 1.1 * y,
            angle = ggraph::node_angle(x, y)
        ),
        hjust = "outward",
        fontface = "bold.italic",
        lineheight = 0.9,
        check_overlap = TRUE
    ) +
    ggraph::theme_graph() +
    ggraph::scale_edge_color_gradient2(
        high = "#C33764",
        mid = "white",
        low = "#1D2671",
        midpoint = 0,
        guide = ggraph::guide_edge_colorbar()
    ) +
    ggplot2::coord_fixed(xlim = c(-2, 2), ylim = c(-2, 2)) +
    ggplot2::guides(edge_width = "none") +
    ggplot2::labs(
        edge_color = "Median Correlation"
    )+
  theme(legend.position = "bottom",
        legend.direction = "horizontal")
```

## Focused Sets

```{r}

```

### Th Subtypes

```{r}
library(msigdbr)

msig_c7 <- msigdbr(
    species = "Homo sapiens",
    category = "C7"
)

th2_genes <- msig_c7 |> 
  filter(grepl("TH2_UP",gs_name)) |> 
  pull(gene_symbol)

th1_genes <- msig_c7 |> 
  filter(grepl("TH1_UP",gs_name)) |> 
  pull(gene_symbol)

tfh_genes <- msig_c7 |> 
  filter(grepl("TFH_UP",gs_name)) |> 
  pull(gene_symbol)


th17_genes <- msig_c7 |> 
  filter(grepl("TH17_UP",gs_name)) |> 
  pull(gene_symbol)

treg_genes <- msig_c7 |> 
  filter(grepl("NATURAL_TREG_UP",gs_name)) |> 
  pull(gene_symbol)


immune_genes <- list(
  Th1 = th1_genes,
  Th2 = th2_genes,
  Th17 = th17_genes,
  Treg = treg_genes,
  Tfh = tfh_genes
)

path_df <- map2_dfr(
  immune_genes, names(immune_genes),
  ~ cd4_se |> 
    pivot_se() |>
      filter(gene_name %in% .x) |>
      dplyr::select(.sample, gene_name, log) |>
      pivot_wider(names_from = gene_name,
                  values_from = log,
                  values_fn = mean) |>
      column_to_rownames(".sample") |>
      as.matrix() |>
      {\(mat) {
        if (ncol(mat) < 2) return(NULL)
        tibble(
          .sample = rownames(mat),
          pc1 = prcomp(mat, center = TRUE, scale. = TRUE)$x[, 1],
          set = .y
        )
      }}()
) |>
  left_join(
    cd4_se |>
      pivot_sample(),
    by = ".sample"
  ) 

# For each set: avg_vst ~ DHT + (1 | cell_line)
th_res <- path_df |>
  group_by(set) |>
  group_modify(~{
    fit <- glm(pc1 ~ pftfev1fvc_actual + gli_age + gli_sex + gli_height + fis, data = .x)
    tidy(fit) 
  }) |>
  ungroup()
```


```{r}


th1_genes <- c(
  "TBX21", "IFNG", "STAT1", "STAT4", "IL12RB1", "IL12RB2",
  "CXCR3", "CCR5"
)

th2_genes <- c(
  "GATA3", "IL4", "IL5", "IL13", "STAT6",
  "CCR4", "CCR8", "IL1RL1"
)

th17_genes <- c(
  "RORC", "IL17A", "IL17F", "IL22",
  "STAT3", "IL23R", "CCR6"
)

treg_genes <- c(
  "FOXP3", "IL2RA", "CTLA4",
  "IKZF2", "TGFB1", "IL10"
)

tfh_genes <- c(
  "BCL6", "CXCR5", "PDCD1",
  "ICOS", "IL21", "SH2D1A",
  "TNFSF13B","TNFSF13"
)

type2_cytokines <- c(
  "IL4", "IL5", "IL13", "IL9",
  "CCL17", "CCL22", "IL1RL1"
)

type1_ifn_genes <- c(
  "IFNG", "TNF", "IL2",
  "STAT1", "IRF1", "CXCL9", "CXCL10"
)

il17_axis_genes <- c(
  "IL17A", "IL17F", "IL22",
  "CXCL8", "CSF3", "S100A8", "S100A9"
)

monocyte_effector_genes <- c(
  "IL6", "TNF", "IL1B",
  "CCL2", "CCL3", "CCL4",
  "NFKBIA"
)

exhaustion_genes <- c(
  "PDCD1", "CTLA4", "LAG3", "HAVCR2", "TIGIT",
  "TOX", "NR4A1", "NR4A2", "NR4A3",
  "BATF"
)

chronic_activation_genes <- c(
  "HIF1A", "DDIT3", "XBP1",
  "ATF4", "JUN", "FOS",
  "GADD45A"
)

library(msigdbr)

msig_c7 <- msigdbr(
    species = "Homo sapiens",
    category = "C7"
)

th2_genes <- msig_c7 |> 
  filter(grepl("TH2_UP",gs_name)) |> 
  pull(gene_symbol)

th1_genes <- msig_c7 |> 
  filter(grepl("TH1_UP",gs_name)) |> 
  pull(gene_symbol)

tfh_genes <- msig_c7 |> 
  filter(grepl("TFH_UP",gs_name)) |> 
  pull(gene_symbol)


th17_genes <- msig_c7 |> 
  filter(grepl("TH17_UP",gs_name)) |> 
  pull(gene_symbol)

treg_genes <- msig_c7 |> 
  filter(grepl("NATURAL_TREG_UP",gs_name)) |> 
  pull(gene_symbol)


immune_genes <- list(
  Th1 = th1_genes,
  Th2 = th2_genes,
  Th17 = th17_genes,
  Treg = treg_genes,
  Tfh = tfh_genes,
  #Type2_Cytokines = type2_cytokines,
  Type1_IFN = type1_ifn_genes,
  IL17_Axis = il17_axis_genes,
  Monocyte_Effector = monocyte_effector_genes,
  Exhaustion = exhaustion_genes,
  Chronic_Activation = chronic_activation_genes
) |> 
  purrr::imap_dfr(~ tibble(
    group = .y,
    gene  = .x
  )) |> 
  left_join(
    fev1_fvc_expom@metadata |> 
      pluck("differential_analysis",
            "differential_abundance") |>
      filter(exp_name %in% c(
        "CD4 T cell RNA",
        "CD16 Monocyte RNA")),
    by=c("gene"="feature_map"),
    relationship = "many-to-many"
  )
```


```{r}
library(msigdbr)

msig_c7 <- msigdbr(
    species = "Homo sapiens",
    category = "C7"
)

th2_genes <- msig_c7 |> 
  filter(grepl("TH2_UP",gs_name)) |> 
  pull(gene_symbol)

th1_genes <- msig_c7 |> 
  filter(grepl("TH1_UP",gs_name)) |> 
  pull(gene_symbol)

tfh_genes <- msig_c7 |> 
  filter(grepl("TFH_UP",gs_name)) |> 
  pull(gene_symbol)


th17_genes <- msig_c7 |> 
  filter(grepl("TH17_UP",gs_name)) |> 
  pull(gene_symbol)

treg_genes <- msig_c7 |> 
  filter(grepl("NATURAL_TREG_UP",gs_name)) |> 
  pull(gene_symbol)


immune_genes <- list(
  Th1 = th1_genes,
  Th2 = th2_genes,
  Th17 = th17_genes,
  Treg = treg_genes,
  Tfh = tfh_genes
)|> 
  purrr::imap_dfr(~ tibble(
    group = .y,
    gene  = .x
  )) |> 
  left_join(
    fev1_fvc_expom@metadata |> 
      pluck("differential_analysis",
            "differential_abundance") |>
      filter(exp_name %in% c(
        "CD4 T cell RNA",
        "CD16 Monocyte RNA")),
    by=c("gene"="feature_map"),
    relationship = "many-to-many"
  )
```

### CD4+ T cell Profile

```{r}
immune_genes |>
  filter(!is.na(gene_name),
         !(grepl("Th",group) & grepl("CD16",exp_name)),
         !(grepl("Monocyte",group) & grepl("CD4",exp_name)),
         grepl("CD4",exp_name)
         ) |> 
  mutate(
    neglogp = -log10(P.Value),
    gene_label = reorder_within(gene_name, logFC, exp_name),
    exp_name = fct_reorder(exp_name, logFC, median, .desc = TRUE)
  ) |>
  ggplot(aes(
    x = logFC,
    y = gene_label,
    color = logFC,
    size = neglogp
  )) +
  geom_point(alpha = 0.8) +
  facet_grid(group ~ exp_name, scales = "free", space = "free") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  tidytext::scale_y_reordered() +
  theme_bw() +
  theme(strip.text.y = element_text(angle = 0,
                                  face="bold.italic"),
        strip.text.x = element_text(face="bold.italic"),
        axis.text.y = element_text(face = "italic")) +
  scale_color_gradient2(
    low = "blue4",
    mid = "grey85",
    high = "red4",
    midpoint = 0
  ) +
  labs(
    x = "Effect size (logFC)",
    y = NULL,
    color = expression("Log"[2] * "FC"),
    size = expression("-Log"[10] * "P")
  )

```


### CD16+ Monocyte Profile

```{r}
immune_genes |>
  filter(!is.na(gene_name),
         !(grepl("Th|Treg|IL17|Tfh",group) & grepl("CD16",exp_name)),
         !(grepl("Monocyte",group) & grepl("CD4",exp_name)),
         grepl("CD16",exp_name)
         ) |> 
  mutate(
    neglogp = -log10(P.Value),
    gene_label = reorder_within(gene_name, logFC, exp_name),
    exp_name = fct_reorder(exp_name, logFC, median, .desc = TRUE)
  ) |>
  ggplot(aes(
    x = logFC,
    y = gene_label,
    color = logFC,
    size = neglogp
  )) +
  geom_point(alpha = 0.8) +
  facet_grid(group ~ exp_name, scales = "free", space = "free") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  tidytext::scale_y_reordered() +
  theme_bw() +
  theme(strip.text.y = element_text(angle = 0,
                                  face="bold.italic"),
        strip.text.x = element_text(face="bold.italic"),
        axis.text.y = element_text(face = "italic")) +
  scale_color_gradient2(
    low = "blue4",
    mid = "grey85",
    high = "red4",
    midpoint = 0
  ) +
  labs(
    x = "Effect size (logFC)",
    y = NULL,
    color = expression("Log"[2] * "FC"),
    size = expression("-Log"[10] * "P")
  )

```


### General Boxplot

```{r}

msig_c5 <- msigdbr(
    species = "Homo sapiens",
    category = "C5"
) |> 
  filter(grepl("CHEMOTAXIS",gs_name))

x <- list(
  
  monocyte=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_MONOCYTE_CHEMOTAXIS",gs_name)) |> 
  pull(gene_symbol),
  
tcell=msigdbr(
    species = "Homo sapiens",
    category = "C5") |> 
  filter(grepl("GOBP_T_CELL_CHEMOTAXIS",gs_name)) |> 
  pull(gene_symbol)

)|> 
  purrr::imap_dfr(~ tibble(
    group = .y,
    gene  = .x
  )) |> 
  left_join(
    fev1_fvc_expom@metadata |> 
      pluck("differential_analysis",
            "differential_abundance") |>
      filter(exp_name %in% c(
        "CD4 T cell RNA",
        "CD16 Monocyte RNA")),
    by=c("gene"="feature_map"),
    relationship = "many-to-many"
  )

immune_genes |> 
  filter(!is.na(gene_name)) |> 
  ggplot(aes(
    x = logFC,
    y = group,
    fill = exp_name,
    color = exp_name
)) +
    geom_boxplot(outlier.shape = NA, alpha = 0.7) +
    #geom_jitter(height = 0.15, alpha = 0.2) +
  scale_fill_manual(values=c(
    "#6190E8","#3f2b96"
  ))+
  scale_color_manual(values=c(
    "#6190E8","#3f2b96"
  ))+
    geom_vline(xintercept = 0, linetype = "dashed") +
    theme_bw()+
  theme(legend.position = "bottom",
        legend.direction = "vertical")+
  labs(
    y = NULL,
    fill = "",
    color= "",
    x = expression("Log"[2] * "FC")
  )

```

## Th Transcription Factors

```{r}
th_tf_df <- fev1_fvc_expom |> tidyexposomics:::.update_assay_colData("CD4 T cell RNA") |>
  pivot_se() |> 
  filter(feature_map %in%
           c("TBX21","GATA3","RORC","FOXP3")) 

th_tf_df |> 
  mutate(fev1fvc_category=factor(
    fev1fvc_category,
    
  ))
    ggplot(aes(
        x=fev1fvc_category,
        y=log2(counts +1),
        fill=fev1fvc_category
    ))+
    geom_boxplot()+
    theme_bw()+
  scale_fill_manual(values = c(
    "#C33764",
    "#654ea3",
    "#24243e"
  ))+
  facet_grid(~ feature_map)+
  geom_pwc()+
  labs(
    x="",
    y=expression("Log"[2]*" (Exp + 1)")
  )+
  rotate_x_text(angle=45)
```

## Overflow

```{r}
# enr_res_rna <- map(
#   c(deg_lst,dtu_lst) |> (\(lst)lst[grepl("RNA",names(lst))])(),
#   ~{
#     run_fenr(
#       selected_genes = .x,
#       universe_genes = rna_features,
#       db = "GO",
#       species = "goa_human",
#       feature_col = "gene_symbol"
#     )
#   }
# ) |> 
#   bind_rows(.id = "group")

# enr_res_iso <- map(
#   c(deg_lst,dtu_lst) |> (\(lst)lst[grepl("Isoform",names(lst))])(),
#   ~{
#     run_fenr(
#       selected_genes = .x,
#       universe_genes = isoform_features,
#       db = "GO",
#       species = "goa_human",
#       feature_col = "gene_symbol"
#     )
#   }
# ) |> 
#   bind_rows(.id = "group")



# cd4_ref <- read.csv("../data/cd4_ref.csv")
# 
# cd4_ref_meta <- cd4_ref |> 
#   dplyr::select("specimen.specimenGuid",
#                 "subject.ageAtFirstDraw",
#                 "cell_type",
#                 "sex",
#                 "subject.ethnicity" ) |> 
#   mutate(combo=paste(specimen.specimenGuid,
#                      cell_type)) |> 
#   distinct() |> 
#   (\(df){
#     rownames(df) = df$combo
#     df
#   })()
# 
# cd4_ref_counts <- cd4_ref |> 
#   mutate(combo=paste(specimen.specimenGuid,
#                      cell_type)) |> 
#   dplyr::select(-c("specimen.specimenGuid",
#                 "subject.ageAtFirstDraw",
#                 "cell_type.1",
#                 "cell_type",
#                 "sex",
#                 "subject.ethnicity") ) |> 
#   column_to_rownames("combo") |> 
#   t()
# 
# identical(rownames(cd4_ref_meta),colnames(cd4_ref_counts))
# 
# cd4_ref_se <- SummarizedExperiment(
#   assays = SimpleList(counts=cd4_ref_counts),
#   colData = DataFrame(cd4_ref_meta)
# )
# 
# cd4_ref_se <- cd4_ref_se |> 
#   adjust_abundance(.factor_unwanted = sex,
#                    .factor_of_interest =  cell_type) |> 
#   adjust_abundance(.factor_unwanted = subject.ageAtFirstDraw,
#                    .factor_of_interest =  cell_type) |> 
#   adjust_abundance(.factor_unwanted = specimen.specimenGuid,
#                    .factor_of_interest =  cell_type)
# 
# library(Matrix)
# 
# expr <- cd4_ref_se@assays@data@listData[["counts_adjusted"]]
# 
# cell_types <- colData(cd4_ref_se)$cell_type
# 
# cell_by_gene <- rowsum(
#   t(expr),
#   group = cell_types
# ) / as.vector(table(cell_types))
# 
# cell_by_gene <- t(cell_by_gene)
# 
# library(immunedeconv)
# 
# cd4_se <- tidyexposomics:::.update_assay_colData(
#   fev1_fvc_expom,
#   "CD4 T cell RNA") |> 
#   (\(se){
#     assay(se,"log") <- log2(assay(se,"counts")+1);
#     se
#   })() |> 
#   deconvolve_cellularity(
#     .abundance = "log",
#     method="cibersort",
#     reference = cell_by_gene[,c(
#   "CD4-positive, alpha-beta memory T cell",
#   "central memory CD4-positive, alpha-beta T cell",
#   "effector memory CD4-positive, alpha-beta T cell",
#   "naive thymus-derived CD4-positive, alpha-beta T cell",
#   "memory regulatory T cell",
#   "naive regulatory T cell",
#   "memory CCR4-positive regulatory T cell"
# )]
#   )
# 
# 
# # cd4_se <- tidyexposomics:::.update_assay_colData(
# #   fev1_fvc_expom,
# #   "CD4 T cell RNA") |> 
# #   (\(se){
# #     assay(se,"log") <- log2(assay(se,"counts")+1);
# #     se
# #   })() |> 
# #   deconvolve_cellularity(
# #     feature_column = "gene_name",
# #     reference = get_X_cibersort() |>
# #       dplyr::select(
# #         all_of(
# #           get_X_cibersort() |>
# #             colnames() |>
# #             (\(chr) chr[grepl("T cells CD4|T cells follicular|T cells regulatory",chr)] )())))
# 
# decon_res <- map(
#   pivot_sample(decon) |> 
#     colnames() |> 
#     (\(chr)chr[grepl("cell|T.reg",chr)])(),
#   ~ {glm( as.formula(paste0(
#     "pftfev1fvc_actual ~ ",
#     .x,
#     " + gli_age + gli_sex + gli_height + fis")),
#     data=pivot_sample(decon)) |>
#       broom::tidy() |>
#       mutate(cell=.x)}) |>
#   bind_rows() |>
#   filter(grepl("cell|T.reg",term))
```


## Summary

* Isoform assays have the greatest number of differentially abundant features, followed by gene expression assays, followed by adductomics and then proteomics.

* Isoforms have greater absolute logFC changes followed by gene expression assays. 

* Isoforms overlap with their gene deg counterparts and isoforms overlap with each other as well. However, the overlap is minimal (Jacarrd Coef. < .15).

## Save Data

```{r}
# Save MultiAssayExperiment Objects
saveRDS(fev1_fvc_expom,file="../results/deg/fev1_fvc_expom.rds")

saveRDS(fev1_fvc_expom,file="../results/sens/fev1_fvc_expom.rds")

# Save Enrichment Data
saveRDS(enr_res_sig,file="../results/deg/enr_res_sig.rds")

# Save Th Data
saveRDS(gsva_assoc_res,file="../results/deg/gsva_assoc_res.rds")
saveRDS(gsva_th_stability_res,file="../results/deg/gsva_th_stability_res.rds")
saveRDS(gsva_th_stability_df,file="../results/deg/gsva_th_stability_df.rds")

```

## Session Info

```{r}
sessionInfo()
```

