---
title: "Exposome AIRWAYS Analysis"
subtitle: "Multiomics Integration"
author: "Jason Laird"
date: today 
date-format: short
engine: knitr
execute: 
  warning: false
  message: false
  cache: true
format: 
  html:
    code-fold: true
    theme: 
      - simplex
      - custom.scss
    fontcolor: '#0c051a'
---

## Setup

### Load Libraries

```{r}
#| warning: false
#| message: false

library(tidyverse)
library(janitor)
library(ggpubr)
library(tidyexposomics)
library(MultiAssayExperiment)
library(patchwork)
library(ggsci)
```

### Load Data

```{r}
# Codebook
aw_cb <- readRDS("../results/input_data/aw_cb.rds")

# MultiAssayExperiment Object
fev1_fvc_expom <- readRDS("../results/integration/fev1_fvc_expom.rds")

# Variable Map
fev1_fvc_expom_var_map <- readRDS(
  "../results/integration/fev1_fvc_expom_var_map.rds")

# Exposure variables
exp_vars <- readRDS("../results/qc/exp_vars.rds")
num_exp_vars <- readRDS("../results/qc/num_exp_vars.rds")

# All variables
vars <- aw_cb$variable |> 
  (\(chr)chr[chr %in% colnames(colData(fev1_fvc_expom))])()

```

## Exposure-Omics Association

### DEG Exposure Association

```{r}
#| fig-width: 12
#| fig-height: 13

# Correlate degs with exposures
fev1_fvc_expom_degs <- fev1_fvc_expom |>
    # Perform correlation analysis between degs
    # and exposures
    run_correlation(
        feature_type = "degs",
        robust = F,
        deg_pval_col = "P.Value",
        deg_pval_thresh = 0.05,
        deg_logfc_thresh = log2(1),
        deg_logfc_col = "logFC",
        exposure_cols = num_exp_vars,
        action = "add",
        correlation_cutoff = 0.3,
        pval_cutoff = 0.05,
        cor_pval_column = "p.value"
    ) |>
    # Perform correlation analysis between factor features
    run_correlation(
        feature_type = "degs",
        robust = F,
        deg_pval_col = "P.Value",
        deg_pval_thresh = 0.05,
        deg_logfc_thresh = log2(1),
        deg_logfc_col = "logFC",
        feature_cors = TRUE,
        action = "add",
        correlation_cutoff = 0.3,
        pval_cutoff = 0.05,
        cor_pval_column = "p.value"
    )

fev1_fvc_expom_degs |>
    plot_correlation_summary(
        mode = "summary",
        feature_type = "degs"
    ) &
  plot_layout(heights = c(5,2))
```

#### Proportion of Assay Correlations per Exposure Category

```{r}
#| fig-width: 6
#| fig-height: 4
fev1_fvc_expom_degs@metadata[["correlation"]][["degs"]] |> 
  tabyl(category, exp_name) |> 
  pivot_longer(-category,
               names_to = "exp_name", 
               values_to = "count") |> 
  ggplot(aes(x = category,
             y = count,
             fill = exp_name)) +
  geom_bar(stat = "identity", 
           position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = "",
    y = "Proportion",
    fill = "Exposure",
    title = "Proportion of Omics type by Exposure Category"
  ) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_npg()
```


#### Exposure Associations Normalized to # of Exposures

```{r}
#| fig-height: 3
#| fig-width: 6
fev1_fvc_expom_degs@metadata$correlation$degs |>
  group_by(category) |> 
  reframe(
    n=n(),
    unique_vars=length(unique(exposure)),
    prop=n/unique_vars
  ) |> 
  filter(unique_vars>1) |> 
  ggplot(aes(x = prop, 
             y = reorder(category,prop),
             fill = category)) +
     geom_col(color="black",linewidth=0.005) +
     labs(
       x = "No. of Assocations/No. Unique Variables",
       y = "",
       title = "Exposure Associations",
       subtitle = "Normalized to No. of Exposures"
     ) +
     theme_classic()+
  scale_fill_npg()+
  theme(legend.position = "none")
  
  
```


#### Number of Shared Features

#### Fix for Circos

```{r}
trace(plot_circos_correlation,edit=T)
#log10(edge_weight+10)
```


```{r}
#| fig-width: 10
#| fig-height: 10
fev1_fvc_expom_degs |>
    plot_circos_correlation(
        feature_type = "degs",
        shared_cutoff = 10,
        low = "blue4",
        mid="white",
        high = "red4",
        midpoint = 50
    )
```

#### Network Creation

#### Fix Network Creation 

```{r}

trace(run_create_network,edit=TRUE)
# Insert this fix:
 feature_type <- match.arg(feature_type)
  action <- match.arg(action)

  all_metadata <- MultiAssayExperiment::metadata(expomicset)
  cor_data <- all_metadata$correlation[[feature_type]]

  if (is.null(cor_data)) {
    stop(sprintf("No correlation data found for feature_type = '%s'", feature_type))
  }

  message("Creating network from correlation results.")

  # --- case 1: exposure-feature correlations
  if (all(c("feature", "exposure", "exp_name", "category") %in% colnames(cor_data))) {

    node_tbl <- dplyr::distinct(
      dplyr::bind_rows(
        dplyr::select(cor_data, name = feature, group = exp_name),
        dplyr::select(cor_data, name = exposure, group = category)
      )
    ) |>
      dplyr::mutate(vertex_id = paste(name, group, sep = "_"))

    edge_df <- cor_data |>
      dplyr::mutate(
        from = paste(exposure, category, sep = "_"),
        to   = paste(feature, exp_name, sep = "_")
      ) |>
      dplyr::select(from, to, correlation, FDR)

  # --- case 2: generic pairwise with var1/var2
  } else if (all(c("var1", "var2", "var1_type", "var2_type") %in% colnames(cor_data))) {

    node_tbl <- dplyr::distinct(
      dplyr::bind_rows(
        dplyr::select(cor_data, name = var1, group = var1_type),
        dplyr::select(cor_data, name = var2, group = var2_type)
      )
    ) |>
      dplyr::mutate(vertex_id = paste(name, group, sep = "_"))

    edge_df <- cor_data |>
      dplyr::mutate(
        from = paste(var1, var1_type, sep = "_"),
        to   = paste(var2, var2_type, sep = "_")
      ) |>
      dplyr::select(from, to, correlation, FDR)

  # --- case 3: correlations across experiments
  } else if (all(c("var1", "var2", "exp_name_1", "exp_name_2") %in% colnames(cor_data))) {

    node_tbl <- dplyr::distinct(
      dplyr::bind_rows(
        dplyr::select(cor_data, name = var1, group = exp_name_1),
        dplyr::select(cor_data, name = var2, group = exp_name_2)
      )
    ) |>
      dplyr::mutate(vertex_id = paste(name, group, sep = "_"))

    edge_df <- cor_data |>
      dplyr::mutate(
        from = paste(var1, exp_name_1, sep = "_"),
        to   = paste(var2, exp_name_2, sep = "_")
      ) |>
      dplyr::select(from, to, correlation, FDR)

  } else {
    stop("Unrecognized correlation format.")
  }

  # --- build graph safely with unique vertex_id ---
  vertices <- node_tbl |>
    dplyr::transmute(
      name = vertex_id,   # unique ID for igraph
      label = name,       # original biological name
      group = group       # grouping info
    )

  g <- igraph::graph_from_data_frame(
    edge_df,
    directed = FALSE,
    vertices = vertices
  )

  net_summary <- .summarize_graph(g)
  net_name <- paste0("network_", feature_type)

  if (action == "add") {
    MultiAssayExperiment::metadata(expomicset)[["network"]][[net_name]] <- list(
      graph = g,
      summary = net_summary
    )
    step_record <- list(run_create_network = list(
      timestamp = Sys.time(),
      params = list(feature_type = feature_type),
      notes = paste0("Created undirected network from correlation results for '",
                     feature_type, "'.")
    ))
    MultiAssayExperiment::metadata(expomicset)$summary$steps <-
      c(MultiAssayExperiment::metadata(expomicset)$summary$steps, step_record)

    message("Network added to metadata as: ", net_name)
    return(expomicset)

  } else {
    return(list(graph = g, summary = net_summary))
  }

```


```{r}
fev1_fvc_expom_degs <- fev1_fvc_expom_degs |>
    run_create_network(
        feature_type = "degs_feature_cor",
        action = "add"
    ) |>
    run_create_network(
        feature_type = "degs",
        action = "add"
    )
```



```{r}
#| fig-width: 10
#| fig-height: 10
fev1_fvc_expom_degs |>
    plot_network(
        network = "degs",
        top_n_nodes = 100,
        include_stats = TRUE,
        cor_thresh = 0.3,
        node_color_var = "group",
        label = TRUE,
        label_top_n = 10
    )

```


#### Fix Exposure Impact

```{r}

trace(run_exposure_impact,edit = TRUE)
.check_suggested(pkg = "tidygraph")

  feature_type <- match.arg(feature_type)
  action <- match.arg(action)

  md <- MultiAssayExperiment::metadata(expomicset)

  # --- checks -----------------------------------------------------------------
  if (is.null(md$correlation[[feature_type]])) {
    stop("Correlation results missing for feature_type = ", feature_type,
         ". Please run run_correlation(..., feature_type = '", feature_type, "').")
  }

  # prefer "..._feature_cor" but fall back to plain "network_<feature_type>"
  graph_key <- paste0("network_", feature_type, "_feature_cor")
  if (is.null(md$network[[graph_key]])) {
    fallback_key <- paste0("network_", feature_type)
    if (!is.null(md$network[[fallback_key]])) {
      graph_key <- fallback_key
    } else {
      stop("Network missing for feature_type = ", feature_type,
           ". Please run run_create_network(..., feature_type = '", feature_type, "').")
    }
  }

  correlation_df <- md$correlation[[feature_type]]
  net_graph      <- md$network[[graph_key]]$graph

  # Optional p-val filter if column exists
  if (!is.null(pval_col) && pval_col %in% names(correlation_df)) {
    correlation_df <- dplyr::filter(correlation_df, .data[[pval_col]] <= pval_thresh)
  }

  # --- centralities from graph -------------------------------------------------
  node_centrality <- tidygraph::as_tbl_graph(net_graph) |>
    tidygraph::activate(nodes) |>
    dplyr::mutate(
      centrality_degree      = tidygraph::centrality_degree(),
      centrality_eigen       = tidygraph::centrality_eigen(),
      centrality_closeness   = tidygraph::centrality_closeness(),
      centrality_betweenness = tidygraph::centrality_betweenness()
    ) |>
    tibble::as_tibble()
  # node_centrality has columns:
  #   - name  (vertex_id used internally by igraph)
  #   - label (original biological name)
  #   - group (exp_name/category/type)
  #   - centrality_* (computed above)

  # --- construct join keys that mirror create_network() ------------------------
  # Case A: exposure-feature schema: feature, exposure, exp_name, category
  if (all(c("feature", "exposure", "exp_name", "category") %in% names(correlation_df))) {

    df <- correlation_df |>
      dplyr::mutate(
        feature_vertex_id  = paste(.data$feature,  .data$exp_name, sep = "_"),
        exposure_vertex_id = paste(.data$exposure, .data$category, sep = "_")
      )

    exposure_col <- "exposure"

  # Case B: generic var1/var2 with explicit types (expect var1 == exposure)
  } else if (all(c("var1", "var2", "var1_type", "var2_type") %in% names(correlation_df))) {

    df <- correlation_df |>
      dplyr::filter(.data$var1_type == "exposure", .data$var2_type != "exposure") |>
      dplyr::mutate(
        exposure            = .data$var1,
        feature             = .data$var2,
        feature_vertex_id   = paste(.data$var2, .data$var2_type, sep = "_"),
        exposure_vertex_id  = paste(.data$var1, .data$var1_type, sep = "_")
      )

    exposure_col <- "exposure"

  # Case C: cross-experiment pairings (use exp_name_1/exp_name_2 as groups)
  } else if (all(c("var1", "var2", "exp_name_1", "exp_name_2") %in% names(correlation_df))) {

    df <- correlation_df |>
      dplyr::mutate(
        # Naming here keeps "exposure" just as a grouping label for summary;
        # adapt as needed for your schema.
        exposure            = .data$var1,
        feature             = .data$var2,
        feature_vertex_id   = paste(.data$var2, .data$exp_name_2, sep = "_"),
        exposure_vertex_id  = paste(.data$var1, .data$exp_name_1, sep = "_")
      )

    exposure_col <- "exposure"

  } else {
    stop("Unrecognized correlation format: cannot construct vertex ids to match the graph.")
  }

  # --- join feature nodes to their centralities via vertex_id ------------------
  df2 <- df |>
    dplyr::left_join(
      node_centrality,
      by = c("feature_vertex_id" = "name")  # name is vertex_id in the graph
    )

  # --- aggregate to exposure-level impact -------------------------------------
  impact_df <- df2 |>
    dplyr::filter(!is.na(.data$centrality_degree)) |>
    dplyr::group_by(.data[[exposure_col]]) |>
    dplyr::summarise(
      mean_degree      = mean(centrality_degree,      na.rm = TRUE),
      mean_eigen       = mean(centrality_eigen,       na.rm = TRUE),
      mean_closeness   = mean(centrality_closeness,   na.rm = TRUE),
      mean_betweenness = mean(centrality_betweenness, na.rm = TRUE),
      n_features       = dplyr::n(),
      .groups = "drop"
    ) |> 
    left_join(
      expomicset@metadata$codebook,
      by=c("exposure"="variable")
    )

  result <- list(exposure_impact = impact_df)

  if (action == "add") {
    all_metadata <- MultiAssayExperiment::metadata(expomicset)
    all_metadata$network$exposure_impact[[feature_type]] <- result
    MultiAssayExperiment::metadata(expomicset) <- all_metadata

    step_record <- list(run_exposure_impact = list(
      timestamp = Sys.time(),
      params    = list(feature_type = feature_type, pval_col = pval_col, pval_thresh = pval_thresh),
      notes     = paste("Computed exposure impact using", feature_type, "correlation network.")
    ))
    MultiAssayExperiment::metadata(expomicset)$summary$steps <-
      c(MultiAssayExperiment::metadata(expomicset)$summary$steps, step_record)

    return(expomicset)
  } else {
    return(result)
  }

  
```


#### Network Impact Analysis

```{r}
#| fig-width: 4
#| fig-height: 12


fev1_fvc_expom_degs <- fev1_fvc_expom_degs |>
    run_exposure_impact(feature_type = "degs",
                        pval_col = "P.Value",
                        pval_thresh = 0.05)

# Plot exposure-omics impact results
fev1_fvc_expom_degs |>
    plot_exposure_impact(
        feature_type = "degs",
        min_per_group = 10,
        bar_cols = "none"
    )
```


#### Test Enrichment

```{r}
#| fig-height: 18
#| fig-width: 12


# Run enrichment analysis on factor features correlated with exposures
a <- fev1_fvc_expom_degs |>
    run_enrichment(
        feature_type = c("degs_cor"),
        feature_col = "feature_map",
        db = c("GO"),
        species = "goa_human",
        fenr_col = "gene_symbol",
        padj_method = "none",
        pval_thresh = 0.1,
        min_set = 1,
        max_set = 800,
        clustering_approach = "diana",
        action = "add"
    )

b <- fev1_fvc_expom_degs |>
    run_enrichment(
        feature_type = c("omics"),
        variable_map = fev1_fvc_expom_degs@metadata |> 
          pluck("differential_analysis","differential_abundance") |> 
          filter(P.Value<0.05) |> 
          dplyr::select(
            feature=feature_map,
            exp_name=exp_name
          ),
        feature_col = "feature_map",
        db = c("GO"),
        species = "goa_human",
        fenr_col = "gene_symbol",
        padj_method = "none",
        pval_thresh = 1,
        min_set = 1,
        max_set = 800,
        clustering_approach = "diana",
        action = "add"
    )

a |>
    plot_enrichment(
        feature_type = "degs_cor",
        plot_type = "dotplot",
        top_n = 10,
        add_top_genes = TRUE,
        top_n_genes = 5
    )
```



### Multi-omics Exposure Association

```{r}
#| fig-width: 12
#| fig-height: 13


# top_factor_features <- fev1_fvc_expom |>
#     extract_results(result = "multiomics_integration") |>
#     pluck("top_factor_features") |>
#     dplyr::select(
#         variable = feature,
#         exp_name
#     )

asthma_sev_features <- fev1_fvc_expom_degs@metadata$differential_analysis$differential_abundance |>
  filter(P.Value<.05 & logFC<0 & abs(logFC)>log(1)) |> 
  dplyr::select(variable=feature,exp_name) 
  # filter(exp_name %in% c(
  #   asthma_sev_features |>
  #     tabyl(exp_name) |> 
  #     filter(n>10) |> 
  #     pull(exp_name)
  # ))

# Correlate top factor features with exposures
fev1_fvc_expom_int <- fev1_fvc_expom |>
    # Perform correlation analysis between factor features
    # and exposures
    run_correlation(
        feature_type = "omics",
        variable_map = asthma_sev_features,
        exposure_cols = num_exp_vars,
        action = "add",
        correlation_cutoff = 0.3,
        pval_cutoff = 0.05,
        cor_pval_column = "p.value"
    ) |>
    # Perform correlation analysis between factor features
    run_correlation(
        feature_type = "omics",
        variable_map = asthma_sev_features,
        feature_cors = TRUE,
        action = "add",
        correlation_cutoff = 0.3,
        pval_cutoff = 0.05,
        cor_pval_column = "p.value"
    )

fev1_fvc_expom_int |>
    plot_correlation_summary(
        mode = "summary",
        feature_type = "omics"
    ) &
  plot_layout(heights = c(5,2))
```

#### Proportion of Assay Correlations per Exposure Category

```{r}
#| fig-width: 6
#| fig-height: 4
fev1_fvc_expom_int@metadata$correlation$omics |> 
  tabyl(category, exp_name) |> 
  pivot_longer(-category,
               names_to = "exp_name", 
               values_to = "count") |> 
  ggplot(aes(x = category,
             y = count,
             fill = exp_name)) +
  geom_bar(stat = "identity", 
           position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = "",
    y = "Proportion",
    fill = "Exposure",
    title = "Proportion of Omics type by Exposure Category"
  ) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_npg()
```

#### Exposure Associations Normalized to # of Exposures

```{r}
#| fig-height: 3
#| fig-width: 6
fev1_fvc_expom_int@metadata$correlation$omics |>
  group_by(category) |> 
  reframe(
    n=n(),
    unique_vars=length(unique(exposure)),
    prop=n/unique_vars
  ) |> 
  filter(unique_vars>1) |> 
  ggplot(aes(x = prop, 
             y = reorder(category,prop),
             fill = category)) +
     geom_col() +
     labs(
       x = "No. of Assocations/No. Unique Variables",
       y = "",
       title = "Exposure Associations",
       subtitle = "Normalized to No. of Exposures"
     ) +
     theme_minimal(base_size = 14)+
  scale_fill_npg()+
  theme(legend.position = "none")
  
  
```

#### Number of Shared Features


```{r}
#| fig-width: 10
#| fig-height: 10
fev1_fvc_expom_int |>
    plot_circos_correlation(
        feature_type = "omics",
        shared_cutoff = 10,
        low = "blue4",
        mid="white",
        high = "red4",
        midpoint = 50
    )
```
#### Network Creation

```{r}
fev1_fvc_expom_int <- fev1_fvc_expom_int |>
    run_create_network(
        feature_type = "omics_feature_cor",
        action = "add"
    ) |>
    run_create_network(
        feature_type = "omics",
        action = "add"
    )
```

```{r}
#| fig-width: 10
#| fig-height: 10
fev1_fvc_expom_int |>
    plot_network(
        network = "omics",
        top_n_nodes = 100,
        include_stats = TRUE,
        cor_thresh = 0.3,
        node_color_var = "group",
        label = TRUE,
        label_top_n = 10
    )

```

#### Network Impact Analysis

```{r}
#| fig-width: 4
#| fig-height: 12


fev1_fvc_expom_int <- fev1_fvc_expom_int |>
    run_exposure_impact(feature_type = "omics",
                        pval_col = "P.Value",
                        pval_thresh = 0.05)

# Plot exposure-omics impact results
fev1_fvc_expom_int |>
    plot_exposure_impact(
        feature_type = "omics",
        min_per_group = 10,
        bar_cols = "none"
    )
```


#### Test Enrichment

```{r}
#| fig-height: 18
#| fig-width: 12


# Run enrichment analysis on factor features correlated with exposures
a <- fev1_fvc_expom_int |>
    run_enrichment(
        feature_type = c("omics_cor"),
        # variable_map = asthma_sev_features |> 
        #   dplyr::select(exp_name,
        #                 feature=variable),
        feature_col = "feature_map",
        db = c("GO"),
        species = "goa_human",
        fenr_col = "gene_symbol",
        padj_method = "none",
        pval_thresh = 0.1,
        min_set = 1,
        max_set = 800,
        clustering_approach = "diana",
        action = "add"
    )

a |>
    plot_enrichment(
        feature_type = "omics_cor",
        plot_type = "dotplot",
        top_n = 10,
        add_top_genes = TRUE,
        top_n_genes = 5
    )

b <- a |> pivot_feature()

bg <- b |> 
  filter(.exp_name=="CD4 T cell Isoforms") |> 
  pull(feature_map) |> 
  unique()

sel <- b |> 
  filter(.feature %in% 
           (a@metadata$correlation$omics |>
              filter(category=="Indoor Pollution") |>
              pull(feature))) |> 
  pull(feature_map) 

c <- run_fenr(selected_genes = sel,
              universe_genes = bg,
              db = "GO",
              species = "goa_human",
              feature_col = "gene_symbol")
```
## Chemical Effects

```{r}
x <- map(exp_vars,~{
  message("Working on: ",.x)
  res <- fev1_fvc_expom |>
    run_differential_abundance(
        formula = as.formula(paste0("~ ",.x, " + gli_age + gli_sex + black + fiscat4 + bmizscore + season + med_ics")),
        method = "limma_trend",
        scaling_method = "none",
        action = "get"
    ) |> 
    filter(P.Value<0.05,
           abs(logFC)>log2(1)) |> 
    mutate(exposure=.x)
})

names(x) <- exp_vars

y <- bind_rows(x,.id="exposure")

z <- c(
  x,
  list("FEV1/FVC"=fev1_fvc_expom_degs@metadata$differential_analysis$differential_abundance |>
         filter(P.Value<0.05,abs(logFC)>log2(1)) )
) |> 
  map(~{
    .x |> 
      mutate(combo=paste(exp_name,feature_map,sep = " ")) |> 
      pull(combo) |> 
      unique()
  }) |> 
  get_pairwise_overlaps()

z |> 
    filter(target=="FEV1/FVC") |> 
    tidygraph::as_tbl_graph() |> 
    ggraph::ggraph(
        layout = "linear",
        circular = TRUE
    ) +
    ggraph::geom_edge_arc(aes(
        color = jaccard,
        width = log10(num_shared+1)
    )) +
    ggraph::geom_node_point(shape = 21,
                            size = 4,
                            alpha = 0.8
    ) +
    ggraph::geom_node_text(
        aes(
            label = name,
            x = 1.1 * x,
            y = 1.1 * y,
            angle = ggraph::node_angle(x, y)
        ),
        hjust = "outward",
        fontface = "bold.italic",
        check_overlap = TRUE
    ) +
    ggraph::theme_graph() +
    ggraph::scale_edge_color_gradient2(
        high = "#0f0c29",
        mid = "white",
        low = "#43C6AC",
        midpoint = 0.1,
        guide = ggraph::guide_edge_colorbar()
    ) +
    ggplot2::coord_fixed(xlim = c(-2, 2), ylim = c(-2, 2)) +
    ggplot2::guides(edge_width = "none") +
    ggplot2::labs(
        edge_color = "Jaccard Coef."
    )
```


```{r}

yy=y |> 
  left_join(
    fev1_fvc_expom_degs@metadata$codebook,
    by=c("exposure"="variable"))

yyy <- yy |> 
  inner_join(
    fev1_fvc_expom@metadata |> 
      pluck("differential_analysis",
            "differential_abundance") |> 
      filter(P.Value<0.05,
             abs(logFC)>log2(1)) |>
      dplyr::select(feature,exp_name) |>
      distinct(),
    by=c("feature"="feature",
         "exp_name"="exp_name")
  )

net <- fev1_fvc_expom |>
    # Perform correlation analysis between factor features
    run_correlation(
        feature_type = "omics",
        variable_map = asthma_sev_features,
        feature_cors = TRUE,
        action = "add",
        correlation_cutoff = 0.3,
        pval_cutoff = 0.05,
        cor_pval_column = "p.value"
    )

net <- net |>
    run_create_network(
        feature_type = "omics_feature_cor",
        action = "add"
    )
  

cor_df <- fev1_fvc_expom_int@metadata |> 
  pluck("correlation",
        "omics_feature_cor") |>
  mutate(bw=case_when(
    exp_name_1 != exp_name_2 ~ "across",
    exp_name_1 == exp_name_2 ~ "within"
  ))

c <- yy |> 
  mutate(combo=paste(exp_name,feature_map,sep = " ")) |> 
  group_by(category) |> 
  reframe(n=paste(
    names(table(combo)[table(combo)>1]),
  ),
  exp_name=exp_name)

c <- 
```


## Summary

*   

## Save Data

```{r}
# Save MultiAssayExperiment Objects
```

## Session Info

```{r}
sessionInfo()
```