---
title: "Exposome AIRWAYS Analysis"
subtitle: "Multiomics Integration"
author: "Jason Laird"
date: today 
date-format: short
engine: knitr
execute: 
  warning: false
  message: false
  cache: true
format: 
  html:
    code-fold: true
    theme: 
      - simplex
      - custom.scss
    fontcolor: '#0c051a'
---

## Setup

### Load Libraries

```{r}
#| warning: false
#| message: false

library(tidyverse)
library(janitor)
library(ggpubr)
library(tidyexposomics)
library(MultiAssayExperiment)
```

### Load Data

```{r}
# Codebook
aw_cb <- readRDS("../results/input_data/aw_cb.rds")

# MultiAssayExperiment Object
fev1_fvc_expom <- readRDS("../results/deg/fev1_fvc_expom.rds")

# Exposure variables
exp_vars <- readRDS("../results/qc/exp_vars.rds")
num_exp_vars <- readRDS("../results/qc/num_exp_vars.rds")

# All variables
vars <- aw_cb$variable |> 
  (\(chr)chr[chr %in% colnames(colData(fev1_fvc_expom))])()

```

## Multi-omics Integration

### FEV1/FVC

#### Integration

```{r}
# Perform multi-omics integration
fev1_fvc_expom <- fev1_fvc_expom |>
    run_multiomics_integration(
        method = "DIABLO",
        n_factors = 5,
        outcome = "pftprefev1fvc_zscore",
        action = "add"
    )
```


#### Association

```{r}
#| fig-width: 6
#| fig-height: 10


# Identify factors that correlate with the outcome
fev1_fvc_expom <- fev1_fvc_expom |>
    run_association(
        source = "factors",
        outcome = "pftfev1fvc_actual",
        #feature_set = exp_vars,
        covariates = c(
            "gli_age",
          "gli_sex",
          "black",
          "fiscat4",
          "bmizscore",
          "season",
          "med_ics"
        ),
        action = "add",
        family = "gaussian"
    )

# Plot the association
fev1_fvc_expom |>
    plot_association(
        source = "factors",
        subtitle = "Covariates: Age, Sex, Income",
        filter_col = "p.value",
        filter_thresh = 0.1,
        r2_col = "adj_r2"
    )
```

#### Extract Top/Shared Factor Features

```{r}
# Extract top features that contribute to a factor
fev1_fvc_expom <- fev1_fvc_expom |>
    extract_top_factor_features(
        method = "percentile",
        pval_col = "p.value",
        pval_thresh = 0.1,
        percentile = 0.95,
        action = "add"
    )

fev1_fvc_expom_var_map <- fev1_fvc_expom |> 
  extract_results(result = "multiomics_integration") |> 
  pluck("top_factor_features") |> 
  inner_join(fev1_fvc_expom |>
               pivot_feature(),
             by=c("feature"=".feature",
                  "exp_name"=".exp_name")) |> 
  dplyr::select(
    feature=feature_map,
    exp_name
  ) |> 
  distinct()


```


## Overlap w. DEGs

```{r}
#| fig-width: 10
#| fig-height: 10


source("./get_pairwise_overlaps.R")


deg_lst <- fev1_fvc_expom |> 
  extract_results("differential_analysis") |> 
  pluck("differential_abundance") |> 
  filter(P.Value<0.05) |> 
  (\(df)split(df,df$exp_name))() |> 
  map(~.x |> 
        pull(feature_map) |> 
        unique()) 
names(deg_lst) <- paste("DEG: ",names(deg_lst))

mult_int_lst <- split(
             fev1_fvc_expom_var_map,
             fev1_fvc_expom_var_map$exp_name) |> 
             map(~.x$feature)
names(mult_int_lst) <- paste("Integration: ",names(mult_int_lst))


c(deg_lst,mult_int_lst) |> 
  get_pairwise_overlaps() |> 
  tidygraph::as_tbl_graph() |> 
  tidygraph::activate(nodes) |> 
  mutate(status=ifelse(grepl("DEG",name),"DEG","Integration")) |> 
  ggraph::ggraph(
        layout = "linear",
        circular = TRUE
    ) +
        ggraph::geom_edge_arc(aes(
            color = jaccard,
            width = log10(num_shared+1)
        )) +
        ggraph::geom_node_point(
          aes(fill=status),
          shape = 21,
            size = 4,
            alpha = 0.8
        ) +
        ggraph::geom_node_text(
            aes(
                label = name,
                x = 1.1 * x,
                y = 1.1 * y,
                angle = ggraph::node_angle(x, y)
            ),
            hjust = "outward",
            fontface = "bold.italic",
            check_overlap = TRUE
        ) +
        ggraph::theme_graph() +
        scale_fill_cosmic()+
        ggraph::scale_edge_color_gradient2(
            low = "blue4",
            mid = "white",
            high = "red4",
            midpoint = 0.05,
            guide = ggraph::guide_edge_colorbar()
        ) +
        ggplot2::coord_fixed(xlim = c(-2, 2), ylim = c(-2, 2)) +
        ggplot2::guides(edge_width = "none") +
        ggplot2::labs(
            edge_color = "Jaccard Coef.",
            fill =""
        )
```


## Summary

*   Only CD4+ T-cell and Adductomics components were associated with FEV1/FVC, suggesting  that these may jointly influence lung function decline.

* CD4+ T-cell Isoform Integration features overlap more with CD4+ T-cell RNA integration features rather than CD4+ T-cell Isoform degs.

## Save Data

```{r}
# Save MultiAssayExperiment Objects
saveRDS(fev1_fvc_expom,file="../results/integration/fev1_fvc_expom.rds")

# Save the variable map
saveRDS(fev1_fvc_expom_var_map,file="../results/integration/fev1_fvc_expom_var_map.rds")
```

## Session Info

```{r}
sessionInfo()
```