---
title: "Exposome AIRWAYS Analysis"
subtitle: "Exposure-Omics Association Analysis"
author: "Jason Laird"
date: today 
date-format: short
engine: knitr
execute: 
  warning: false
  message: false
  cache: true
format: 
  html:
    code-fold: true
    theme: 
      - simplex
      - custom.scss
    fontcolor: '#0c051a'
---


## Setup 

### Load Libraries

```{r}
#| warning: false
#| message: false

library(tidyverse)
library(janitor)
library(ggpubr)
library(ggsci)
library(tidyexposomics)
library(MultiAssayExperiment)
library(furrr)
library(tidytext)

# Source QC helper functions
invisible(lapply(
  list.files(
    path = "./",
    pattern = "*.R", full.names = TRUE
  ),
  source
))
```

### Load Data

```{r}
# Codebook
aw_cb <- readRDS("../results/input_data/aw_cb.rds")

# MultiAssayExperiment Object
fev1_fvc_expom <- readRDS("../results/sens/fev1_fvc_expom.rds")

# Exposure variables
exp_vars <- readRDS("../results/qc/exp_vars.rds")
num_exp_vars <- readRDS("../results/qc/num_exp_vars.rds")

# All variables
vars <- aw_cb$variable |> 
  (\(chr)chr[chr %in% colnames(colData(fev1_fvc_expom))])()

# Setting colors for exposure categories
cat_colors <- scales::alpha(get_palette("aaas",13), 0.5) |> 
  (\(colors){
    names(colors) <- aw_cb |> 
      filter(variable %in% exp_vars) |> 
      pull(category) |> 
      unique()
    colors
  })()
```

## Differential Abundance - Exposures

```{r}
# Run differential abundance analysis for all exposures
plan(multisession)

exp_omics_res <- future_map(
  exp_vars,
  ~{
  message("Working on: ",.x)
  res <- fev1_fvc_expom |>
    run_differential_abundance(
        formula = as.formula(paste0("~ ",.x, " + gli_age + gli_sex + gli_height + fis")),
        method = "limma_trend",
        scaling_method = "none",
        action = "get"
    ) |> 
    filter(P.Value<0.05,
           abs(logFC)>log2(1)) |> 
    mutate(exposure=.x)
})

names(exp_omics_res) <- exp_vars

# exp_omics_res_df <- bind_rows(exp_omics_res,.id="exposure") |> 
#   left_join(aw_cb,
#             by=c("exposure"="variable"))

exp_omics_res_df <- bind_rows(exp_omics_res) |> 
  left_join(aw_cb,
            by=c("exposure"="variable"))
```

## Sensitivity Analysis

```{r}
# ran this on the cluster for speed
# exp_omics_sens_res <- map(
#   exp_vars,
#   ~{
#   message("Working on: ",.x)
#     fev1_fvc_expom |> 
#       run_sensitivity_analysis(
#         base_formula = as.formula(paste0("~ ",.x, " + gli_age + gli_sex + gli_height + fis")),
#         methods = c("limma_trend"),
#         scaling_methods = c("none"),
#         covariates_to_remove = c(
#             "gli_age",
#             "gli_sex",
#             "gli_height",
#             "fis"
#         ),
#         pval_col = "P.Value",
#         logfc_col = "logFC",
#         logFC_threshold = log2(1),
#         pval_threshold = 0.05,
#         stability_metric = "stability_score",
#         bootstrap_n = 100,
#         action = "get"
#     ) |> 
#       pluck("feature_stability") |> 
#       filter(effect_consistency>.7) |> 
#       mutate(exposure=.x)
# })
# 
# saveRDS(exp_omics_sens_res, file="../results/exposure_omics/exp_omics_sens_res.rds")

  # fev1_fvc_expom |>
  #   run_differential_abundance(
  #       formula = as.formula(paste0("~ ",.x, " + gli_age + gli_sex + gli_height + fis")),
  #       method = "limma_trend",
  #       scaling_method = "none",
  #       action = "get"
  #   ) |> 
  #   filter(P.Value<0.05,
  #          abs(logFC)>log2(1)) |> 
  #   mutate(exposure=.x)

exp_omics_sens_res_df <- exp_omics_sens_res |> 
  bind_rows() |> 
  inner_join(exp_omics_res_df,
             by=c("feature"="feature",
                  "exp_name"="exp_name",
                  "exposure"="exposure"))

```


### Overlap

```{r}
asthma_exp_comparison <- c(
  exp_omics_res,
  list("FEV1/FVC"=fev1_fvc_expom@metadata |> 
         pluck("differential_analysis",
               "differential_abundance") |>
         filter(P.Value<0.05,abs(logFC)>log2(1)) )
) |> 
  map(~{
    .x |> 
      mutate(combo=paste(exp_name,feature_map,sep = " ")) |> 
      pull(combo) |> 
      unique()
  }) |> 
  get_pairwise_overlaps()

asthma_exp_comparison |> 
    filter(target=="FEV1/FVC") |> 
    tidygraph::as_tbl_graph() |> 
    ggraph::ggraph(
        layout = "linear",
        circular = TRUE
    ) +
    ggraph::geom_edge_arc(aes(
        color = jaccard,
        width = log10(num_shared+1)
    )) +
    ggraph::geom_node_point(shape = 21,
                            size = 4,
                            alpha = 0.8
    ) +
    ggraph::geom_node_text(
        aes(
            label = name,
            x = 1.1 * x,
            y = 1.1 * y,
            angle = ggraph::node_angle(x, y)
        ),
        hjust = "outward",
        fontface = "bold.italic",
        check_overlap = TRUE
    ) +
    ggraph::theme_graph() +
  ggraph::scale_edge_color_gradient2(
        high = "#5f2c82",
        mid = "white",
        low = "#49a09d",
        midpoint = 0.1,
        guide = ggraph::guide_edge_colorbar()
    ) +
    # ggraph::scale_edge_color_gradient2(
    #     high = "#0f0c29",
    #     mid = "white",
    #     low = "#43C6AC",
    #     midpoint = 0.1,
    #     guide = ggraph::guide_edge_colorbar()
    # ) +
    ggplot2::coord_fixed(xlim = c(-2, 2), ylim = c(-2, 2)) +
    ggplot2::guides(edge_width = "none") +
    ggplot2::labs(
        edge_color = "Jaccard Coef."
    )
```

### Lollipop Plot

```{r}
jc_asthma <- asthma_exp_comparison |> 
    filter(target=="FEV1/FVC") |> 
  left_join(aw_cb,
            by=c("source"="variable")) |> 
    # Order categories by median effect size
    mutate(
        category = forcats::fct_reorder(
            category, jaccard, .fun = median, .desc = TRUE),
        
        # Order terms by estimate (within categories)
        exposure = forcats::fct_reorder(source, jaccard)
    ) 

median_df <- jc_asthma |> 
  group_by(category) |> 
  summarize(median_jaccard = median(jaccard, na.rm = TRUE), .groups = "drop")

jc_asthma <- jc_asthma |> 
  left_join(median_df, by = "category")

jc_asthma |> 
    ggplot(aes(x = jaccard, 
               y = exposure,
               fill = jaccard)) +
    
    # Add lollipop stems
    geom_segment(aes(x = 0, xend = jaccard, y = exposure, yend = exposure),
                 color = "grey70", size = 0.6) +
    
    # Lollipop heads
    geom_point(shape = 21, size = 2.5, stroke = 0.8) +
    
    # Facet by category with colored strips
    ggh4x::facet_grid2(category ~ ., 
                       scales = "free",
                       space = "free", 
                       strip = ggh4x::strip_themed(
                           background_y = ggh4x::elem_list_rect(
                               fill = cat_colors[levels(jc_asthma$category)]))) +
    
    # Color gradients for fill and stroke
    scale_fill_gradient2(
        low = "#c94b4b",
        mid = "grey85",
        high = "#4b134f",
        midpoint = .1
    ) +
    
    # Reference line at 0
    geom_vline(xintercept = 0, linetype = "dashed") +
    
    # Labels and theme
    labs(
        title = NULL,
        x = "Jaccard Coef.",
        y = "",
        fill = "Jaccard Coef."
    ) +
    theme_bw(base_size = 13) +
    theme(
        panel.grid.minor = element_blank(),
        strip.text.y = element_text(angle = 0, face = "bold.italic")
    )
```
### Overlap within Groups

```{r}
jc_cat <- asthma_exp_comparison |> 
    filter(!target=="FEV1/FVC") |> 
    left_join(aw_cb |> 
                  dplyr::select(variable,
                                source_category=category),
              by=c("source"="variable")) |> 
    left_join(aw_cb |> 
                  dplyr::select(variable,
                                target_category=category),
              by=c("target"="variable")) |> 
    filter(source_category==target_category) 

# jc_cat |> 
#     ggplot(aes(
#         x = jaccard,
#         y = fct_reorder(source_category,jaccard),
#         fill = source_category,
#         color = source_category
#     )) +
#     geom_boxplot(outlier.shape = NA, alpha = 0.7) +
#     #geom_jitter(height = 0.15, alpha = 0.2) +
#     scale_fill_manual(values=cat_colors)+
#     scale_color_manual(values=cat_colors)+
#     theme_bw()+xlim(c(0,.5))+
#     theme(legend.position = "none",axis.title.x = element_text(size = 10))+
#     labs(
#         y = NULL,
#         fill = "",
#         color= "",
#         x = "Jaccard Coef."
#     )

jc_cat |> 
  group_by(source_category) |> 
  reframe(med=median(jaccard)) |> 
  ggplot(aes(
    x = med,
    y = fct_reorder(source_category,med),
    fill = source_category,
    color = source_category
  )) +
  geom_col(color="black",linewidth = 0.005)+
  theme_classic()+
  scale_fill_manual(values=cat_colors)+
  scale_color_manual(values=cat_colors)+
  scale_x_continuous(breaks = c(0,0.1,0.2))+
  theme(
    legend.position = "none",
    axis.title.x = element_text(size = 10))+
  labs(
    y = NULL,
    fill = "",
    color= "",
    x = "Median Jaccard Coef."
  )
```
### Overlap with FEV1/FVC

```{r}
jc_cat_fvc <- asthma_exp_comparison |> 
    filter(target=="FEV1/FVC") |> 
    left_join(aw_cb |> 
                  dplyr::select(variable,
                                source_category=category),
              by=c("source"="variable")) |> 
    left_join(aw_cb |> 
                  dplyr::select(variable,
                                target_category=category),
              by=c("target"="variable")) 

# jc_cat_fvc |> 
#     ggplot(aes(
#         x = jaccard,
#         y = fct_reorder(source_category,jaccard),
#         fill = source_category,
#         color = source_category
#     )) +
#     geom_boxplot(outlier.shape = NA, alpha = 0.7) +
#     #geom_jitter(height = 0.15, alpha = 0.2) +
#     scale_fill_manual(values=cat_colors)+
#     scale_color_manual(values=cat_colors)+
#     theme_bw()+xlim(c(0,.3))+
#     theme(legend.position = "none",axis.title.x = element_text(size = 10))+
#     labs(
#         y = NULL,
#         fill = "",
#         color= "",
#         title="Overlap with FEV1/FVC Features",
#         x = "Jaccard Coef."
#     )

jc_cat_fvc |> 
  group_by(source_category) |> 
  reframe(med=median(jaccard)) |> 
  ggplot(aes(
    x = med,
    y = fct_reorder(source_category,med),
    fill = source_category,
    color = source_category
  )) +
  geom_col(color="black",linewidth = 0.005)+
  theme_classic()+
  scale_fill_manual(values=cat_colors)+
  scale_color_manual(values=cat_colors)+
  scale_x_continuous(breaks = c(0,0.1,0.2))+
  theme(
    legend.position = "none",
    axis.title.x = element_text(size = 10))+
  labs(
    y = NULL,
    fill = "",
    color= "",
    x = "Median Jaccard Coef."
  )
```

### Exposure Omics Impact

```{r}
exp_omics_sens_res_df |> 
    group_by(category, exp_name, exposure) |> 
    summarise(n_degs = n(), 
              .groups = "drop") |> 
    group_by(category, exp_name) |> 
    summarise(median_degs = median(n_degs),
              .groups = "drop") |>
    mutate(
        category = fct_reorder(category, median_degs, .fun = sum)
    ) |>
    ggplot(aes(x = category, y = median_degs, fill = exp_name)) +
    geom_col(alpha=.8,color="black",linewidth = 0.005) +
    coord_flip() +
    scale_fill_manual(values=c('#3e2b6f',
                               "#bdb3d4"	,
                               '#f27070'	,
                               "#f1a04b"	,
                               #"#ffd966",
                               "#1f0322",
                               "#8a1c48"	,
                               "#d96241"	,
                               "#f1cbcc"	,
                               "#99bbaa")) +
    labs(x = "", 
         y = paste("Median No. of ","Differential Features",sep="\n"),
         fill = "",
         title = "") +
    theme_classic()+
  theme(legend.position = "bottom")
```

## Enrichment Analysis

### Set Genesets

```{r}
enr_sets <- exp_omics_sens_res_df |> 
  mutate(direction = ifelse(
    logFC > 0, "Upregulated", "Downregulated"),
    combo=paste(exposure,exp_name,direction,sep=";")) |> 
  (\(df)split(df,df$combo))() |> 
  map(~{
    .x |> 
      pull(feature_map) |> 
      unique()
  }) |> 
  (\(lst) lst[lengths(lst)>20])()

rna_features <- pivot_feature(fev1_fvc_expom) |>
  filter(.exp_name =="CD4 T cell RNA") |>
  pull(feature_map)

isoform_features <-  pivot_feature(fev1_fvc_expom) |>
  filter(.exp_name =="CD4 T cell Isoforms") |>
  pull(feature_map)

proteomics_features <- pivot_feature(fev1_fvc_expom) |>
  filter(.exp_name =="Proteomics") |>
  pull(feature_map)
```


### Enrichment

```{r}
plan(multisession)

exp_enr_res_rna <- future_map(
  enr_sets |> 
    (\(lst)lst[grepl("RNA",names(lst))])(),
  ~{
    options(GO_ANNOTATION_URL = "http://current.geneontology.org/annotations/goa_human.gaf.gz")
    
    run_fenr(
      selected_genes = .x,
      universe_genes = rna_features,
      db = "GO",
      species = "goa_human",
      feature_col = "gene_symbol"
    )
  },
  .progress = TRUE
) |> 
  bind_rows(.id = "group")


exp_enr_res_iso <- future_map(
  enr_sets |> 
    (\(lst)lst[grepl("Isoform",names(lst))])(),
  ~{
    options(GO_ANNOTATION_URL = "http://current.geneontology.org/annotations/goa_human.gaf.gz")
    
    run_fenr(
      selected_genes = .x,
      universe_genes = isoform_features,
      db = "GO",
      species = "goa_human",
      feature_col = "gene_symbol"
    )
  },
  .progress = TRUE
) |> 
  bind_rows(.id = "group")

exp_enr_res_prot <- future_map(
  enr_sets |> 
    (\(lst)lst[grepl("Proteomics",names(lst))])(),
  ~{
    options(GO_ANNOTATION_URL = "http://current.geneontology.org/annotations/goa_human.gaf.gz")
    
    run_fenr(
      selected_genes = .x,
      universe_genes = isoform_features,
      db = "GO",
      species = "goa_human",
      feature_col = "gene_symbol"
    )
  },
  .progress = TRUE
) |> 
  bind_rows(.id = "group")

exp_enr_res_sig <- exp_enr_res_rna |> 
  bind_rows(exp_enr_res_iso) |> 
  bind_rows(exp_enr_res_prot) |> 
  filter(p_adjust<.1,
         n_with_sel>5,
         N_with< 1000)

parent_terms <- find_parent_terms(exp_enr_res_sig)

exp_enr_res_sig <- exp_enr_res_sig |> 
  left_join(parent_terms |> 
              dplyr::select(term,parentTerm),
            by=c("term_name"="term"))
```
### Shared Terms

```{r}
exp_enr_res_sig <- exp_enr_res_sig |> 
  separate(col = "group",
           into = c("exposure","exp_name","direction"),
           sep=";",
           remove = F) |> 
  mutate(cell=case_when(
    grepl("CD4",group) ~ "CD4",
    grepl("CD16",group) ~ "CD16",
    .default = "Other"
  )) |> 
  left_join(aw_cb |> 
              dplyr::select(variable,category),
            by=c("exposure"="variable"))
  # separate_rows(ids,sep = ", ") |> 
  # (\(df) split(df,df$exposure))() |> 
```

### No. of DEGs by Exposure

```{r}
exp_omics_sens_res_df |>
    group_by(exp_name,exposure) |>
    reframe(n=n()) |>
    mutate(log_n=log2(n)) |> 
    ggplot(aes(
        y=fct_reorder(exposure,log_n),
        x=fct_reorder(exp_name,-log_n),
        fill=log_n))+
    geom_tile()+
    scale_fill_gradientn(
        colors = c(
            "#3a1c71",
            "#d76d77",
            "#ffaf7b"
        ),
        na.value = "grey80")+
  theme_classic()+
  rotate_x_text(angle=45)+
  labs(
    x="",
    y="",
    fill=expression("Log"[2]*" No. of DEGs")
  )
```

### * Barplot No. of DEGs by Exposure

```{r}
exp_omics_sens_res_df |>
    group_by(exposure) |>
    reframe(n=n()) |>
    ggplot(aes(
        y=fct_reorder(exposure,n),
        x=n,
        fill= n))+
    geom_col(
      color="black",linewidth = 0.005
    )+
    scale_fill_gradientn(
        colors = c(
            "#3a1c71",
            "#d76d77",
            "#ffaf7b"
        ),
        na.value = "grey80")+
  theme_classic()+
  rotate_x_text(angle=45)+
  labs(
    x="",
    y="",
    fill=expression("No. of Differential Features")
  )
```


### Common Pathways

```{r}
common_pathways_vec <-
  exp_enr_res_sig |>
  filter(grepl("RNA",exp_name)) |> 
  filter(p_adjust < 0.05) |>
  group_by(parentTerm,cell) |>
  reframe(
    n_exposures  = n_distinct(exposure),
    n_cells      = n_distinct(exp_name),
    median_logp  = median(-log10(p_adjust+1e-12)),
    median_count = median(n_with_sel),
    enrichment_score = median_logp * median_count
  ) |>
  distinct() |> 
  filter(n_exposures>=3) |>  
  # group_by(cell) |>
  # arrange(desc(n_exposures), desc(enrichment_score)) |> 
  # slice_head(n=20) |> 
  pull(parentTerm) |> 
  unique()

common_pathways <-
  exp_enr_res_sig |>
  #filter(grepl("RNA",exp_name)) |> 
  filter(p_adjust < 0.05) |>
  group_by(parentTerm,exp_name,direction) |>
  reframe(
    n_exposures  = n_distinct(exposure),
    n_cells      = n_distinct(exp_name),
    median_logp  = median(-log10(p_adjust+1e-12)),
    median_count = median(n_with_sel),
    enrichment_score = median_logp * median_count,
    enrichment_score = ifelse(
      direction == "Upregulated",
      enrichment_score,
      -enrichment_score
      )
  ) |>
  distinct() |> 
  arrange(desc(n_exposures), desc(enrichment_score)) |> 
  filter(parentTerm %in% common_pathways_vec) |> 
  mutate(
    cell=case_when(
      grepl("CD16",exp_name) ~ "CD16+ Monocyte",
      grepl("CD4",exp_name) ~ "CD4+ T cell",
      grepl("Proteomics",exp_name) ~ "Proteomics"
    ),
    mode=case_when(
      grepl("RNA",exp_name) ~ "RNA",
      grepl("Isoform",exp_name) ~ "Isoform",
      .default = "Protein-Level"
    )
  )

common_pathways |>
  arrange(enrichment_score) |> 
  mutate(
    parentTerm=fct_reorder(parentTerm,enrichment_score),
    combo=paste(cell,mode),
    #parentTerm=reorder_within(parentTerm,enrichment_score,combo)
    ) |> 
  group_by(parentTerm,cell,mode) |> 
  mutate(cons=ifelse(length(unique(direction))>1,"not","is")  ) |> 
  ungroup() |> 
  filter(cons=="is") |> 
    ggplot(aes(x=direction,
               y=parentTerm,
               fill=enrichment_score))+
    geom_tile()+
    scale_fill_gradient2(
        low ="blue4",
        mid="white",
        high = "red4",
        midpoint = 0,
        limits=c(-100,100),
        oob=scales::squish
    )+
    ggh4x::facet_nested(. ~ cell+mode)+
    theme_bw()+
  #scale_y_reordered()+
    theme(strip.text.x = element_text(face="bold.italic"),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank())+
    labs(
        x="",
        y="",
        fill ="Enrichment Score"
    )
```


### * RNA Common Pathways

```{r}
common_pathways |>
  arrange(enrichment_score) |> 
  mutate(
    parentTerm=fct_reorder(parentTerm,enrichment_score),
    combo=paste(cell,mode),
    #parentTerm=reorder_within(parentTerm,enrichment_score,combo)
    ) |> 
  group_by(parentTerm,cell,mode) |> 
  mutate(cons=ifelse(length(unique(direction))>1,"not","is")  ) |> 
  ungroup() |> 
  filter(cons=="is",
         mode=="RNA") |> 
    ggplot(aes(x=direction,
               y=parentTerm,
               fill=enrichment_score))+
    geom_tile()+
    scale_fill_gradient2(
        low ="blue4",
        mid="white",
        high = "red4",
        midpoint = 0,
        limits=c(-100,100),
        oob=scales::squish
    )+
    ggh4x::facet_nested(. ~ cell+mode)+
    theme_bw()+
  #scale_y_reordered()+
    theme(strip.text.x = element_text(face="bold.italic"),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank())+
    labs(
        x="",
        y="",
        fill ="Enrichment Score"
    )

```

### * Barplot

```{r}
common_pathways |>
  mutate(parentTerm=str_wrap(parentTerm,width=50)) |> 
    group_by(exp_name) |>
    slice_max(
        order_by = n_exposures,
        n = 10,
        with_ties = FALSE
    ) |>
    ungroup() |>
    mutate(
        parentTerm = reorder_within(
            parentTerm,
            enrichment_score,
            exp_name
        )
    ) |>
    ggplot(aes(
        x = enrichment_score,
        y = parentTerm,
        fill = direction
    )) +
    geom_col() +
    scale_y_reordered() +
    scale_fill_manual(values = c("blue4", "red4")) +
    facet_grid(
        exp_name ~ .,
        scales = "free_y",
        space = "free"
    ) +
    theme_bw()+
    theme(strip.text.y = element_text(angle = 0,
                                      face = "bold.italic"))+
    labs(
        x="Enrichment Score",
        y="",
        fill=""
    )
```
### * Tileplot

```{r}
common_pathways |>
  mutate(parentTerm=str_wrap(parentTerm,width=50),
         exp_name=case_when(
           exp_name == "CD16 Monocyte Isoforms" ~ "CD16+ Monocyte\nIsoforms",
           exp_name == "CD4 T cell Isoforms" ~ "CD4+ T cell\nIsoforms",
           exp_name == "CD4 T cell RNA" ~ "CD4+ T cell\nRNA",
           exp_name == "CD16 Monocyte RNA" ~ "CD16+ Monocyte\nRNA",
           exp_name == "Proteomics" ~ "Proteomics",
         )) |> 
    group_by(exp_name) |>
    slice_max(
        order_by = n_exposures,
        n = 10,
        with_ties = FALSE
    ) |>
    ungroup() |>
    mutate(
        parentTerm = reorder_within(
            parentTerm,
            enrichment_score,
            exp_name
        )
    ) |>
    ggplot(aes(
        x = 1,
        y = parentTerm,
        fill = enrichment_score
    )) +
    geom_tile() +
    scale_y_reordered() +
    scale_fill_gradient2(
        low ="blue4",
        mid="white",
        high = "red4",
        limits=c(-100,100),
        oob=scales::squish,
        midpoint = 0
    ) +
    facet_grid(
        exp_name ~ .,
        scales = "free_y",
        space = "free"
    ) +
    theme_bw()+
    theme(strip.text.y = element_text(angle = 0,
                                      face = "bold.italic"),
          axis.text.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.position = "bottom",
          legend.title.position = "top")+
    labs(
        x="",
        y="",
        fill="Enrichment Score"
    )
```


## Exposures Driving Terms

```{r}
exp_gene_long <-
  exp_enr_res_sig |>
  mutate(
    gene = str_split(
      ids,
      pattern = "\\s*[;,/|]\\s*|\\s+"
    )
  ) |>
  unnest(gene) |>
  mutate(gene = str_trim(gene)) |>
  filter(!is.na(gene), gene != "")

pathway_hits <-
  exp_gene_long |>
  filter(
    #parentTerm == "antigen binding",
    #parentTerm == "positive regulation of neutrophil chemotaxis",
    #parentTerm == "PML body",
    #parentTerm == "cytokine receptor activity",
    #parentTerm == "immunoglobulin complex",
    #parentTerm == "immunological synapse",
    #parentTerm == "T cell differentiation in thymus",
    #parentTerm == "toll-like receptor 4 signaling pathway",
    #parentTerm == "damaged DNA binding",
    p_adjust < 0.05
  ) |>
  distinct(
    parentTerm,
    gene,
    exposure,
    category,
    direction,
    cell,
    exp_name
  )

gene_summary <-
  pathway_hits |>
  group_by(gene) |>
  summarise(
    n_exposures  = n_distinct(exposure),
    n_categories = n_distinct(category),
    n_cells      = n_distinct(cell),
    directions   = paste(sort(unique(direction)), collapse = ", "),
    .groups = "drop"
  ) |>
  arrange(desc(n_exposures), desc(n_categories), desc(n_cells))

category_summary <-
  pathway_hits |>
  group_by(category) |>
  summarise(
    n_genes     = n_distinct(gene),
    n_exposures = n_distinct(exposure),
    n_hits      = n(),
    .groups = "drop"
  ) |>
  arrange(desc(n_genes), desc(n_exposures))

```


### * CD4 Heatmap

```{r}

gene_term_counts <-
  exp_enr_res_sig |>
  filter(grepl("CD4",group)) |> 
  filter(p_adjust < 0.05) |>
  mutate(
    gene = str_split(
      ids,
      pattern = "\\s*[;,/|]\\s*|\\s+"
    )
  ) |>
  unnest(gene) |>
  mutate(gene = str_trim(gene)) |>
  filter(!is.na(gene), gene != "") |>
  filter(
    parentTerm == "antigen binding" |
    parentTerm == "positive regulation of neutrophil chemotaxis" |
    parentTerm == "PML body" |
    parentTerm == "cytokine receptor activity" |
    parentTerm == "immunoglobulin complex" |
    parentTerm == "immunological synapse" |
    parentTerm == "T cell differentiation in thymus" |
    parentTerm == "toll-like receptor 4 signaling pathway" |
    parentTerm == "damaged DNA binding"
  ) |>
  distinct(parentTerm, gene, exposure) |>
  group_by(parentTerm, gene) |>
  summarise(
    n_exposures = n_distinct(exposure),
    .groups = "drop"
  )

top_genes_per_term <-
  gene_term_counts |>
  group_by(parentTerm) |>
  slice_max(
    order_by = n_exposures,
    n = 5,
    with_ties = FALSE
  ) |>
  mutate(
    gene = fct_reorder(gene, n_exposures)
  ) |>
  ungroup()

ggplot(
    top_genes_per_term,
    aes(
        y = parentTerm,
        x = gene,
        fill = n_exposures
    )
) +
    geom_tile() +
    scale_fill_gradientn(
        colors = c(
            "#3a1c71",
            "#d76d77",
            "#ffaf7b"
        ),
        na.value = "grey80")+
    theme_minimal(base_size = 12) +
    theme(
        axis.text.x = element_text(face="italic",
                                   angle = 45,
                                   hjust = 1,
                                   vjust = 1),
        panel.grid = element_blank(),
        
    ) +
    labs(
        x = "",
        y = "",
        title = expression("CD4"^"+"*" T cell"),
        fill = "No. of Exposures"
    )
```


### * CD16 Heatmap

```{r}

gene_term_counts <-
    exp_enr_res_sig |>
    filter(grepl("CD16",group)) |> 
    filter(p_adjust < 0.05) |>
    mutate(
        gene = str_split(
            ids,
            pattern = "\\s*[;,/|]\\s*|\\s+"
        )
    ) |>
    unnest(gene) |>
    mutate(gene = str_trim(gene)) |>
    filter(!is.na(gene), gene != "") |>
    filter(
        parentTerm == "antigen binding" |
            parentTerm == "positive regulation of neutrophil chemotaxis" |
            parentTerm == "PML body" |
            parentTerm == "cytokine receptor activity" |
            parentTerm == "immunoglobulin complex" |
            parentTerm == "immunological synapse" |
            parentTerm == "T cell differentiation in thymus" |
            parentTerm == "toll-like receptor 4 signaling pathway" |
            parentTerm == "damaged DNA binding"
    ) |>
    distinct(parentTerm, gene, exposure) |>
    group_by(parentTerm, gene) |>
    summarise(
        n_exposures = n_distinct(exposure),
        .groups = "drop"
    )

top_genes_per_term <-
    gene_term_counts |>
    group_by(parentTerm) |>
    slice_max(
        order_by = n_exposures,
        n = 5,
        with_ties = FALSE
    ) |>
    mutate(
        gene = fct_reorder(gene, n_exposures)
    ) |>
    ungroup()

ggplot(
    top_genes_per_term,
    aes(
        y = parentTerm,
        x = gene,
        fill = n_exposures
    )
) +
    geom_tile() +
    scale_fill_gradientn(
        colors = c(
            "#3a1c71",
            "#d76d77",
            "#ffaf7b"
        ),
        na.value = "grey80")+
    theme_minimal(base_size = 12) +
    theme(
        axis.text.x = element_text(face="italic",
                                   angle = 45,
                                   hjust = 1,
                                   vjust = 1),
        panel.grid = element_blank(),
        
    ) +
    labs(
        x = "",
        y = "",
        title = expression("CD16"^"+"*" Monocyte"),
        fill = "No. of Exposures"
    )
```

### * Combined

```{r}
gene_term_counts <-
  exp_enr_res_sig |>
  filter(
    grepl("CD4|CD16", group),
    p_adjust < 0.05
  ) |>
  mutate(
    cell_type = case_when(
      grepl("CD4", group)  ~ "CD4+ T cell",
      grepl("CD16", group) ~ "CD16+ Monocyte"
    ),
    gene = str_split(
      ids,
      pattern = "\\s*[;,/|]\\s*|\\s+"
    )
  ) |>
  unnest(gene) |>
  mutate(gene = str_trim(gene)) |>
  filter(!is.na(gene), gene != "") |>
  filter(
    parentTerm == "antigen binding" |
      parentTerm == "positive regulation of neutrophil chemotaxis" |
      parentTerm == "PML body" |
      parentTerm == "cytokine receptor activity" |
      parentTerm == "immunoglobulin complex" |
      parentTerm == "defense response to virus" |
      parentTerm == "immunological synapse" |
      parentTerm == "cell-cell contact zone" |
      parentTerm == "T cell differentiation in thymus" |
      parentTerm == "toll-like receptor 4 signaling pathway" |
      parentTerm == "damaged DNA binding"
  ) |>
  distinct(cell_type, parentTerm, gene, exposure) |>
  group_by(cell_type, parentTerm, gene) |>
  summarise(
    n_exposures = n_distinct(exposure),
    .groups = "drop"
  )

top_genes_per_term <-
  gene_term_counts |>
  group_by(cell_type, parentTerm) |>
  slice_max(
    order_by = n_exposures,
    n = 5,
    with_ties = FALSE
  ) |>
  mutate(
    gene = fct_reorder(gene, n_exposures)
  ) |>
  ungroup()

ggplot(
  top_genes_per_term,
  aes(
    x = gene,
    y = parentTerm,
    fill = n_exposures
  )
) +
  geom_tile() +
  scale_fill_gradientn(
    colors = c(
      "#3a1c71",
      "#d76d77",
      "#ffaf7b"
    ),
    na.value = "grey80"
  ) +
  facet_wrap(~ cell_type, ncol = 1, scales = "free_y") +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(
      face = "italic",
      angle = 45,
      hjust = 1,
      vjust = 1
    ),
    panel.grid = element_blank(),
    strip.text = element_text(face = "bold")
  ) +
  labs(
    x = "",
    y = "",
    title = "Top 5 genes per pathway by number of exposures",
    fill = "No. of Exposures"
  )


```

## * Top Genes

```{r}
top_exp_genes <- exp_omics_sens_res_df |>
  mutate(feature_map=case_when(
    exp_name == "Adduct Burden" ~ feature_id,
    .default = feature_map
  )) |> 
  filter(!grepl("^ENSG|Y_RNA|Metazoa|-AS1$|-AS2$",feature_map)) |> 
  dplyr::select(feature_map,exposure,exp_name) |>
  distinct() |>  
  group_by(feature_map,exp_name) |>
  reframe(n_exposures=n_distinct(exposure)) |>
  group_by(feature_map,exp_name) |>
  mutate(total=sum(n_exposures)) |>
  ungroup() |> 
  group_by(exp_name) |> 
  arrange(desc(n_exposures)) |> 
  slice_head(n=5)

top_exp_genes |> 
  mutate(exp_name=fct_reorder(exp_name,n_exposures),
         feature_map=fct_reorder(feature_map,n_exposures)) |> 
  ggplot(
  aes(
    x = exp_name,
    y = feature_map,
    fill = n_exposures
  )
) +
  geom_tile(color="black",linewidth = 0.005) +
  scale_fill_gradientn(
    colors = c(
      "#3a1c71",
      "#d76d77",
      "#ffaf7b"
    ),
    na.value = "grey80"
  ) +
  #facet_wrap(~ cell_type, ncol = 1, scales = "free_y") +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.y = element_text(
      face = "italic"
    ),
    panel.grid = element_blank(),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom",
    legend.title.position = "top"
  ) +
  rotate_x_text(angle = 45)+
  labs(
    x = "",
    y = "",
    title = "",
    fill = "No. of Exposures"
  )
```




### hot mess

```{r}
exp_process_summary <-
  exp_enr_res_sig |>
  filter(p_adjust < 0.05) |>
  group_by(exposure, term_id, term_name) |>
  summarise(
    n_cell_contexts = n_distinct(exp_name),
    n_directions    = n_distinct(direction),
    mean_enrichment = mean(enrichment),
    max_odds_ratio  = max(odds_ratio),
    min_p           = min(p_adjust),
    median_p        = median(p_adjust),
    median_count    = median(n_with_sel),
    .groups = "drop"
  ) |>
  arrange(exposure, min_p)

exp_term_matrix <-
  exp_enr_res_sig |>
  filter(p_adjust < 0.05) |>
  distinct(exposure, parentTerm) |>
  mutate(present = 1) |>
  pivot_wider(
    names_from = exposure,
    values_from = present,
    values_fill = 0
  ) |> 
  filter(!is.na(parentTerm)) |> 
  column_to_rownames("parentTerm")

exp_pathway_mat <-
  exp_enr_res_sig |>
  filter(p_adjust < 0.05) |>
  group_by(exposure, parentTerm) |>
  summarise(
    median_logp  = median(-log10(p_adjust+1e-12)),
    median_count = median(n_with_sel),
    enrichment_score = median_logp * median_count,
    .groups = "drop"
  ) |>
  dplyr::select(-c(median_logp,median_count)) |> 
  pivot_wider(
    names_from  = exposure,
    values_from = enrichment_score,
    values_fill = 0
  ) |> 
  filter(!is.na(parentTerm)) |> 
  column_to_rownames("parentTerm")

exp_pathway_mat_scaled <- t(scale(t(exp_pathway_mat)))
mat_scaled[is.na(mat_scaled)] <- 0

exp_pathway_dist_exp <- dist(t(exp_pathway_mat_scaled), method = "euclidean")

exp_pathway_dist_hc_exp   <- hclust(exp_pathway_dist_exp, method = "ward.D2")

factoextra::fviz_nbclust(
    exp_pathway_mat_scaled,
    FUNcluster = factoextra::hcut,
    method = "gap",
    hc_method = "ward.D2",
    nboot = 100
)

clusters <- data.frame(
  exposure=names(cutree(exp_pathway_dist_hc_exp, k = 9)),
  cluster=cutree(exp_pathway_dist_hc_exp, k = 9))
  
exp_enr_clustered <-
  exp_enr_res_sig |>
  inner_join(clusters, by = "exposure") |>
  filter(p_adjust < 0.05)


top_terms_per_cluster <-
  exp_enr_clustered |>
  group_by(cluster, cell,parentTerm) |>
  summarise(
    n_exposures = n_distinct(exposure),
    median_logp  = median(-log10(p_adjust+1e-12)),
    median_count = median(n_with_sel),
    enrichment_score = median_logp * median_count,
    .groups = "drop"
  ) |>
  group_by(cluster) |>
  slice_max(enrichment_score, n = 10) |>
  ungroup()


```


### Shared terms among exposures

```{r}
exp_enr_g <- exp_enr_res_sig |> 
  (\(df)split(df,df$exposure))() |> 
  map(~{
    .x |> 
      pull(parentTerm) |> 
      unique()
  }) |> 
  get_pairwise_overlaps() |> 
  dplyr::select(source,target,jaccard) |> 
  tidygraph::as_tbl_graph() |> 
  mutate(degree=tidygraph::centrality_degree()) |> 
  left_join(aw_cb |> 
              dplyr::select(variable,category),
            by=c("name"="variable")) |> 
  mutate(lab=ifelse(degree>60,name,NA))

exp_enr_g |> 
    ggraph(layout="fr") +
    geom_edge_link(aes(
      edge_width = jaccard),
      alpha = 0.7,
      color = "grey80") +
    geom_node_point(aes(
      size = degree,
      color = category),
      alpha = 0.9) +
  geom_node_point(aes(
      size = degree,
      color = category),
                  color = "black",
                  stroke = 0.005,
                  shape = 21,
                  fill = NA) +
    geom_node_label(
      aes(label = lab),
      size = 3,
      repel = TRUE) +
  scale_color_manual(values=cat_colors)+
    scale_size_continuous(range = c(2, 10)) +
    theme_void()
```


## Save Data

```{r}
saveRDS(exp_omics_res,file="../results/exposure_omics/exp_omics_res.rds")

saveRDS(exp_omics_sens_res_df,file="../results/exposure_omics/exp_omics_sens_res_df.rds")

saveRDS(exp_enr_res_sig,file="../results/exposure_omics/exp_enr_res_sig.rds")

```

