---
title: "Exposome AIRWAYS Analysis"
subtitle: "Exposure-Omics Th Association Analysis"
author: "Jason Laird"
date: today 
date-format: short
engine: knitr
execute: 
  warning: false
  message: false
  cache: true
format: 
  html:
    code-fold: true
    theme: 
      - simplex
      - custom.scss
    fontcolor: '#0c051a'
---


## Setup 

### Load Libraries

```{r}
#| warning: false
#| message: false

library(tidyverse)
library(janitor)
library(ggpubr)
library(ggsci)
library(tidyexposomics)
library(MultiAssayExperiment)
library(furrr)
library(patchwork)
library(msigdbr)
library(GSVA)

# Source QC helper functions
invisible(lapply(
  list.files(
    path = "./",
    pattern = "*.R", full.names = TRUE
  ),
  source
))
```

### Load Data

```{r}
# Codebook
aw_cb <- readRDS("../results/input_data/aw_cb.rds")

# MultiAssayExperiment Object
fev1_fvc_expom <- readRDS("../results/sens/fev1_fvc_expom.rds")

decon <- readRDS("../results/deconvolution/decon.rds")

# Exposure variables
exp_vars <- readRDS("../results/qc/exp_vars.rds")
num_exp_vars <- readRDS("../results/qc/num_exp_vars.rds")

# Load Exposure-omics association results
exp_omics_sens_res_df <- readRDS("~/jhu/projects/exposomePilot/results/exposure_omics/exp_omics_sens_res_df.rds")

# All variables
vars <- aw_cb$variable |> 
  (\(chr)chr[chr %in% colnames(colData(fev1_fvc_expom))])()

# Setting colors for exposure categories
cat_colors <- scales::alpha(get_palette("aaas",13), 0.5) |> 
  (\(colors){
    names(colors) <- aw_cb |> 
      filter(variable %in% exp_vars) |> 
      pull(category) |> 
      unique()
    colors
  })()
```

## Th Correlation

```{r}
# prep the data
ps <- tidybulk::pivot_sample(decon)

cell_cols <- colnames(ps)[grepl("cell|T.reg", colnames(ps))]

cell_mat <- ps |>
  dplyr::select(all_of(cell_cols)) |>
  as.matrix()

# Zero handling (required for CLR)
cell_mat[cell_mat == 0] <- 1e-6

# CLR transform
clr_mat <- compositions::clr(cell_mat)
colnames(clr_mat) <- paste0("clr_", cell_cols)

meta <- ps |>
  dplyr::select(
    all_of(exp_vars),
    gli_age,
    gli_sex,
    gli_height,
    fis
  )

model_df <- bind_cols(meta, as.data.frame(clr_mat))

# fit the models
decon_res <- expand_grid(
  exposure = exp_vars,
  clr_col  = colnames(clr_mat)
) |>
  mutate(
    model = map2(
      exposure,
      clr_col,
      ~ {
        fml <- as.formula(paste(
          .x,
          "~",
          .y,
          "+ gli_age + gli_sex + gli_height + fis"
        ))
        lm(fml, data = model_df)
      }
    )
  ) |>
  mutate(
    tidy = map(model, tidy)
  ) |>
  unnest(tidy) |>
  filter(term == clr_col) |>
  mutate(
    cell = sub("^clr_", "", clr_col)
  ) |>
  select(
    exposure,
    cell,
    estimate,
    std.error,
    statistic,
    p.value
  )

# clean the labels
decon_res <- decon_res |>
  mutate(
    cell_clean = case_when(
      cell == "Memory.CD4.T.cell.TFH"               ~ "Tfh (memory CD4)",
      cell == "Memory.CD4.T.cell.Th1"               ~ "Th1 (memory CD4)",
      cell == "Memory.CD4.T.cell.Th2"               ~ "Th2 (memory CD4)",
      cell == "Memory.CD4.T.cell.Th17"              ~ "Th17 (memory CD4)",
      cell == "Memory.CD4.T.cell.Th1.Th17"          ~ "Th1/Th17 (memory CD4)",
      cell == "Terminal.effector.memory.CD4.T.cell" ~ "Terminal effector memory CD4",
      cell == "naive.CD4.T.cell"                    ~ "Naive CD4",
      cell == "T.reg"                               ~ "Treg",
      TRUE                                          ~ cell
    )
  )

# adjust p-values
decon_res <- decon_res |>
  group_by(exposure) |>
  mutate(fdr = p.adjust(p.value, method = "fdr")) |>
  ungroup()
```

### Th Association Stability

```{r}
# base data
stability_df <- bind_cols(
  tidybulk::pivot_sample(decon),
  as.data.frame(clr_mat)
)

# run sensitivity for each CLR cell outcome
th_stability_res <- map(
  colnames(clr_mat),
  ~ {
    run_exwas_sensitivity(
      df = stability_df,
      exposures = exp_vars,
      outcome = .x,
      covariates = c("gli_age", "gli_sex", "gli_height", "fis"),
      covariates_to_remove = c("gli_age", "gli_sex", "gli_height", "fis"),
      family = gaussian(),
      bootstrap_n = 100
    )$sensitivity_df |>
      mutate(
        clr_col = .x,
        cell = sub("^clr_", "", .x)
      )
  }
) |> 
  bind_rows()

# summarise stability (mirrors your EXWAS logic)
th_stability_df <- th_stability_res |>
  mutate(
    signif = (p_value <= 0.05),
    sign = sign(estimate)
  ) |>
  group_by(exposure, cell) |>
  summarise(
    n_tests = n(),
    n_signif = sum(signif, na.rm = TRUE),
    prop_signif = n_signif / n_tests,
    median_effect = median(estimate, na.rm = TRUE),
    median_p = median(p_value, na.rm = TRUE),
    sd_effect = sd(estimate, na.rm = TRUE),
    sign_consistency =
      ifelse(all(is.na(sign)), NA_real_,
             abs(sum(sign, na.rm = TRUE)) / n_tests),
    stability_score = prop_signif * sign_consistency,
    .groups = "drop"
  ) |>
  mutate(
    cell_clean = case_when(
      cell == "Memory.CD4.T.cell.TFH"               ~ "Tfh (memory CD4)",
      cell == "Memory.CD4.T.cell.Th1"               ~ "Th1 (memory CD4)",
      cell == "Memory.CD4.T.cell.Th2"               ~ "Th2 (memory CD4)",
      cell == "Memory.CD4.T.cell.Th17"              ~ "Th17 (memory CD4)",
      cell == "Memory.CD4.T.cell.Th1.Th17"          ~ "Th1/Th17 (memory CD4)",
      cell == "Terminal.effector.memory.CD4.T.cell" ~ "Terminal effector memory CD4",
      cell == "naive.CD4.T.cell"                    ~ "Naive CD4",
      cell == "T.reg"                               ~ "Treg",
      TRUE                                          ~ cell
    )
  ) |> 
  left_join(
    decon_res,
    by=c(
      "cell_clean"="cell_clean",
      "cell"="cell",
      "exposure"="exposure"
    )
  )


```


###  Heatmap

```{r}
th_stability_df |>
    group_by(exposure) |>
  mutate(med=mean(estimate)) |> 
  ungroup() |> 
  mutate(p_lab=ifelse(p.value<0.1,"lab",NA)) |> 
    mutate(
        exposure=fct_reorder(exposure,med),
        cell_clean=fct_reorder(cell_clean,-med)) |> 
    ggplot(aes(
        x = cell_clean,
        y = exposure,
        fill = estimate
    )) +
    geom_tile(color = "grey90") +
    scale_fill_gradient2(
        low = "#1D2671",
        mid = "white",
        high = "#C33764",
        midpoint = 0,limits=c(-.3,.3),oob=scales::squish
    ) +
    theme_bw() +
    theme(
        axis.text.x = element_text(angle = 90, hjust = 1,vjust = .35),
        axis.text.y = element_text(),
        legend.position = "right"
    ) +
    labs(
        x = "",
        y = "",
        fill = "Estimate"
    )
```

### * Full Heatmap

```{r}
th_stability_df |>
    group_by(exposure) |>
    mutate(
        med   = mean(estimate),
        n_sig = sum(p.value < 0.05)
    ) |>
    ungroup() |>
    group_by(cell_clean) |>
    mutate(
        n_sig_cell = sum(p.value < 0.05)
    ) |>
    ungroup() |>
    mutate(
        exposure   = forcats::fct_reorder(exposure, med),
        cell_clean = forcats::fct_reorder(cell_clean, -n_sig_cell),
        sig        = p.value < 0.05
    ) |>
    (\(df) {
        
        ## ---- HEATMAP ---------------------------------------------------------
        p_heat <-
            ggplot(df, aes(
                x = cell_clean,
                y = exposure,
                fill = estimate
            )) +
            geom_tile(color = "grey90") +
            geom_point(
                data = \(d) d |> filter(sig),
                color = "black",
                size = 1.8
            ) +
            scale_fill_gradient2(
                low = "#1D2671",
                mid = "white",
                high = "#C33764",
                midpoint = 0,
                limits = c(-0.3, 0.3),
                oob = scales::squish
            ) +
            scale_x_discrete(expand = c(0, 0)) +   ## <<< FIX
            theme_bw() +
            theme(
                axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.35),
                legend.position = "bottom"
            ) +
            labs(x = "", y = "", fill = "Estimate")
        
        ## ---- RIGHT BAR -------------------------------------------------------
        p_bar_right <-
            ggplot(
                distinct(df, exposure, n_sig),
                aes(x = n_sig, y = exposure)
            ) +
            geom_col(fill = "grey30") +
            theme_classic() + scale_x_continuous(breaks = c(0,1,2))+
            theme(
                axis.text.y = element_blank(),
                axis.ticks.y = element_blank()
            ) +
            labs(x = "", y = "")
        
        ## ---- TOP BAR ---------------------------------------------------------
        p_bar_top <-
            ggplot(
                distinct(df, cell_clean, n_sig_cell),
                aes(x = cell_clean, y = n_sig_cell)
            ) +
            geom_col(fill = "grey30") +
            theme_classic() +
            theme(
                axis.text.x  = element_blank(),
                axis.ticks.x = element_blank(),
            ) +
            labs(x = "", y = "")
        
        ## ---- COMPOSE ---------------------------------------------------------
        top    <- (p_bar_top | plot_spacer()) +
            plot_layout(widths = c(3, 1))            ## <<< FIX
        
        bottom <- (p_heat | p_bar_right) +
            plot_layout(widths = c(4, 1))            ## <<< FIX
        
        top / bottom +
            plot_layout(heights = c(1, 12))
    })()

```

### * Exposures Barplot

```{r}
th_stability_df |> 
  filter(sign_consistency>0.7) |> 
    group_by(exposure) |>
    reframe(nsig=sum(p.value<0.05),
            med=median(abs(estimate)))|> 
    filter(nsig>0) |> 
    ggplot(aes(
        x=med,
        y=reorder(exposure,med),
        fill=med
    ))+
    # geom_point()+
    # geom_segment(aes(
    #     x=0,
    #     xend=med,
    #     y=exposure
    # ))+
  geom_col(color="black",linewidth = 0.005)+
    scale_fill_gradientn(
        colors = c(
            "#3a1c71",
            "#d76d77"
        ),
        na.value = "grey80")+
    theme_classic()+
  scale_x_continuous(
    breaks = c(0,.1,.2)
  )+
    labs(
        x="Median |Estimate|",
        y="",
        fill="Median |Estimate|"
    )
```

### * Th Barplot

```{r}
th_stability_df |>
    filter(p.value<0.1,
           sign_consistency>0.7) |>
    group_by(cell_clean) |> 
    reframe(med=median(estimate)) |>
    ggplot(aes(
        x = med,
        y = fct_reorder(cell_clean,med),
        fill = med
    )) +
    geom_col(color = "black", 
             linewidth = 0.005) +
    theme_bw() +
    scale_fill_gradient2(
        low ="blue4",
        mid="white",
        high = "red4",
        midpoint = 0
    )+scale_x_continuous(breaks = c(-1,-.5,.1))+
    labs(x = "",
         y = "", 
         fill = "Median Estimate") +
    theme(
        strip.text.x = element_text(
            face = "bold.italic"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor   = element_blank(),
        legend.position = "right",
        legend.title.position = "top"
    )
```
### * Th Barplot Thresholds

```{r}
pmap(
  expand_grid(
    consistency = c(0.3, 0.5, 0.7, 0.9),
    p_cutoff    = c(0.1, 0.05)
),
    \(consistency,p_cutoff){
      th_stability_df |>
    filter(p.value<p_cutoff,
           sign_consistency>consistency) |>
    group_by(cell_clean) |> 
    reframe(med=median(estimate),
            n=n(),
            cell_clean=paste(cell_clean," (",n,")",sep="")) |> 
        distinct() |> 
        (\(df){
        df |> 
            ggplot(aes(
                x = med,
                y = fct_reorder(cell_clean,med),
                fill = med
            )) +
            geom_col(color = "black", 
                     linewidth = 0.005) +
            theme_bw() +
            scale_fill_gradient2(
                low ="blue4",
                mid="white",
                high = "red4",
                midpoint = 0
            )+
        scale_x_continuous(breaks = c(-1,-.5,.1))+
            labs(x = "",title="Th-Exposures",
                 y = "", 
                 caption=paste("P <",p_cutoff, "& Effect Consistency >",consistency*100,"%"),
                 fill = "Median Estimate") +
            theme(
                strip.text.x = element_text(
                    face = "bold.italic"),
                panel.grid.major.x = element_blank(),
                panel.grid.minor   = element_blank(),
                legend.position = "right",
                legend.title.position = "top",
                plot.title = element_text(
                  size=12
                )
            ) })()
    }) |> 
  wrap_plots(ncol = 2)

th_stability_df |>
    filter(p.value<0.1,
           sign_consistency>0.9) |>
    group_by(cell_clean) |> 
    reframe(med=median(estimate)) |> (\(df){
        df |> 
            ggplot(aes(
                x = med,
                y = fct_reorder(cell_clean,med),
                fill = med
            )) +
            geom_col(color = "black", 
                     linewidth = 0.005) +
            theme_bw() +
            scale_fill_gradient2(
                low ="blue4",
                mid="white",
                high = "red4",
                midpoint = 0
            )+scale_x_continuous(breaks = c(-1,-.5,.1))+
            labs(x = "",title="Th-Exposures", 
                 subtitle = paste("N:", nrow(th_stability_df |>
                                                 filter(p.value<0.1,
                                                        sign_consistency>0.9))," Exposures"),
                 y = "", caption="P < 0.1 & Effect Consistency > 90%",
                 fill = "Median Estimate") +
            theme(
                strip.text.x = element_text(
                    face = "bold.italic"),
                panel.grid.major.x = element_blank(),
                panel.grid.minor   = element_blank(),
                legend.position = "right",
                legend.title.position = "top"
            ) })()
    
```


### * Heatmap

```{r}
th_stability_df |>
    filter(p.value<0.1,
           sign_consistency>0.7) |> 
    mutate(
        exposure=fct_reorder(exposure,estimate),
        cell_clean=fct_reorder(cell_clean,estimate),
    ) |>
    ggplot(
        aes(
            x = cell_clean,
            y = exposure,
            fill = estimate
        )
    ) +
    geom_tile(
        color = "black",
        linewidth = 0.2,
        height = 0.9
    ) +
    scale_fill_gradient2(
        low ="blue4",
        mid="white",
        high = "red4",
        midpoint = 0,
        limits=c(-0.2,.2),
        oob=scales::squish
    )+
    labs(
        x = "",
        y = "",
        fill = "Estimate"
    ) +
    theme_bw() +
    theme(
        # rotate labels
        axis.text.x = element_text(
            angle = 90,
            vjust = 0.5,
            hjust = 1
        ),
        
        # remove x-axis formatting
        axis.title.x = element_blank(),
        axis.line.x  = element_blank(),
        
        axis.text.y = element_text(size = 9),
        legend.position = "right",
        legend.title.position = "top"
    )
# th_stability_df |>
#   filter(p.value<0.1,
#          sign_consistency>0.7) |> 
#     inner_join(
#         path_stability_df |> 
#           filter(p.value<0.1,
#                  sign_consistency>0.7) |> 
#           janitor::tabyl(exposure),
#         by = "exposure"
#     ) |>
#     mutate(
#         exposure_label = if_else(
#             n > 1,
#             paste0(strrep("*", n), " ", exposure),
#             exposure
#         ),
#         exposure_label=fct_reorder(exposure_label,estimate),
#         cell_clean=fct_reorder(cell_clean,estimate),
#     ) |>
#   ggplot(
#         aes(
#             x = cell_clean,
#             y = exposure_label,
#             fill = estimate
#         )
#     ) +
#     geom_tile(
#         color = "black",
#         linewidth = 0.2,
#         height = 0.9
#     ) +
#     scale_fill_gradient2(
#                 low ="blue4",
#                 mid="white",
#                 high = "red4",
#                 midpoint = 0,
#                 limits=c(-0.2,.2),
#                 oob=scales::squish
#             )+
#     labs(
#         x = "",
#         y = "",
#         fill = "Estimate",
#         caption = "* Appears in more than one phenotype."
#     ) +
#     theme_bw() +
#     theme(
#         # rotate labels
#         axis.text.x = element_text(
#             angle = 90,
#             vjust = 0.5,
#             hjust = 1
#         ),
#         
#         # remove x-axis formatting
#         axis.title.x = element_blank(),
#         axis.line.x  = element_blank(),
#         
#         axis.text.y = element_text(size = 9),
#         legend.position = "right",
#         legend.title.position = "top"
#     )
```

### Cell Summary

```{r}
th_stability_df |> 
  filter(sign_consistency>0.7) |> 
    group_by(cell_clean) |>
    reframe(nsig=sum(p.value<0.05),
            mean_effect_up=median(estimate[estimate>0]),
            mean_effect_down=median(estimate[estimate<0]),
            med=median(estimate))|> 
    filter(nsig>0) |> 
    mutate(cell_clean=fct_reorder(cell_clean,med)) |> 
    (\(df) {
        heat <- df |> 
            ggplot(aes(
                x=1,
                y=cell_clean,
                fill=med
            ))+
            geom_tile()+
            theme_bw()+
            labs(
                x="",
                y="",
                fill="Median Estimate"
            )+
            scale_fill_gradient2(
                low ="blue4",
                mid="white",
                high = "red4",
                midpoint = 0
            )+
            theme(axis.text.x = element_blank(),
                  axis.ticks.x = element_blank(),
                  legend.position = "bottom",
                  legend.title.position = "top")
        
        bar <- df |> 
            ggplot(aes(
                x=nsig,
                y=cell_clean
            ))+
            geom_col(fill="grey25")+
            theme_classic()+
            labs(
                x="No. of Sig. Associations",
                y=""
            )+
            theme(axis.text.y = element_blank(),
                  axis.ticks.y = element_blank())
        
        (heat|bar)+plot_layout(
            widths = c(1,3),
            guides = "collect")
    })()
    
```

### Flipped

```{r}
decon_res |> 
    group_by(cell_clean) |>
    reframe(nsig=sum(p.value<0.05),
            mean_effect_up=median(estimate[estimate>0]),
            mean_effect_down=median(estimate[estimate<0]),
            med=median(estimate))|> 
    filter(nsig>0) |> 
    mutate(cell_clean=fct_reorder(cell_clean,med)) |> 
    (\(df) {
        heat <- df |> 
            ggplot(aes(
                y=1,
                x=cell_clean,
                fill=med
            ))+
            geom_tile()+
            theme_bw()+
            labs(
                x="",
                y="",
                fill="Median Estimate"
            )+
            scale_fill_gradient2(
                low ="blue4",
                mid="white",
                high = "red4",
                midpoint = 0
            )+
            theme(axis.text.y = element_blank(),
                  axis.ticks.y = element_blank(),
                  axis.text.x = element_text(
                      angle=90,
                      hjust = 1,vjust = .3
                  ))
        
        bar <- df |> 
            ggplot(aes(
                y=nsig,
                x=cell_clean
            ))+
            geom_col(fill="grey25")+
            theme_classic()+
            labs(
                y="No. of Sig. Associations",
                x=""
            )+
            theme(axis.text.x = element_blank(),
                  axis.ticks.x = element_blank())
        
        (bar/heat)+plot_layout(
            heights = c(2,1),
            guides = "collect")
    })()
```
## Pathway Associations

### Extract SEs

```{r}
cd4_se <- tidyexposomics:::.update_assay_colData(
  fev1_fvc_expom,
  "CD4 T cell RNA") |> 
  (\(se){
    assay(se,"log") <- log2(assay(se,"counts")+1);
    se
  })()


cd16_se <- tidyexposomics:::.update_assay_colData(
  fev1_fvc_expom,
  "CD16 Monocyte RNA") |> 
  (\(se){
    assay(se,"log") <- log2(assay(se,"counts")+1);
    se
  })()
```

### Extract Genesets

```{r}
msig_c5 <- msigdbr(
    species = "Homo sapiens",
    category = "C5"
)

gene_sets <- list(
  # activation
  "Monocyte Activation" =  msig_c5 |> 
    filter(gs_name == "GOBP_MONOCYTE_ACTIVATION") |> 
    pull(gene_symbol),
  "T cell Activation" =  msig_c5 |> 
    filter(gs_name == "GOBP_T_CELL_ACTIVATION") |> 
    pull(gene_symbol),
  
  # chemotaxis
  "Monocyte Chemotaxis" =  msig_c5 |> 
    filter(gs_name == "GOBP_MONOCYTE_CHEMOTAXIS") |> 
    pull(gene_symbol),
  "T cell Chemotaxis" =  msig_c5 |> 
    filter(gs_name == "GOBP_T_CELL_CHEMOTAXIS") |> 
    pull(gene_symbol),
  
  # tolerance/exhaustion
  "Tolerance" =  msig_c5 |> 
    filter(gs_name == "GOBP_TOLERANCE_INDUCTION") |> 
    pull(gene_symbol),
  
  # senescence
  "Senescence" =  msig_c5 |> 
    filter(gs_name == "GOBP_CELLULAR_SENESCENCE") |> 
    pull(gene_symbol)

  
)

```

### GSVA Dfs

```{r}
cd4_mat <- cd4_se |>
  pivot_se() |>
  dplyr::select(.sample, gene_name, log) |>
  pivot_wider(
    names_from  = gene_name,
    values_from = log,
    values_fn   = mean
  ) |>
  column_to_rownames(".sample") |>
  as.matrix()

cd16_mat <- cd16_se |>
  pivot_se() |>
  dplyr::select(.sample, gene_name, log) |>
  pivot_wider(
    names_from  = gene_name,
    values_from = log,
    values_fn   = mean
  ) |>
  column_to_rownames(".sample") |>
  as.matrix()

cd4_gsva_par <- gsvaParam(
  t(cd4_mat),            # matrix: samples x genes
  gene_sets,         # list of gene sets
  #method = "ssgsea",   # or "gsva", "zscore", "plage"
  kcdf = "Gaussian",   # use "Gaussian" for log-transformed RNA-seq data
  absRanking = TRUE    # use original ssGSEA normalization
)

cd16_gsva_par <- gsvaParam(
  t(cd16_mat),            # matrix: samples x genes
  gene_sets,         # list of gene sets
  #method = "ssgsea",   # or "gsva", "zscore", "plage"
  kcdf = "Gaussian",   # use "Gaussian" for log-transformed RNA-seq data
  absRanking = TRUE    # use original ssGSEA normalization
)

cd4_gsva_df <- gsva(cd4_gsva_par) |> 
  t() |> 
  as.data.frame() |> 
  rownames_to_column(".sample") |> 
  dplyr::select(-c("Monocyte Activation",
                   "Monocyte Chemotaxis")) |> 
  dplyr::rename( "T cell Senescence"=Senescence,
               "T cell Tolerance"=Tolerance)

cd16_gsva_df <- gsva(cd16_gsva_par) |> 
  t() |> 
  as.data.frame() |> 
  rownames_to_column(".sample") |> 
  dplyr::select(-c("T cell Activation",
                   "T cell Chemotaxis")) |> 
  dplyr::rename( "Monocyte Senescence"=Senescence,
               "Monocyte Tolerance"=Tolerance)

gsva_df <- cd4_gsva_df |> 
  inner_join(cd16_gsva_df,
             by=".sample") |> 
  left_join(
    cd16_se |>
      tidybulk::pivot_sample(),
    by = ".sample"
  )
```

### Assocation Testing

```{r}
# build all pathway × exposure combinations
gsva_grid <- expand_grid(
  gsva_pathway = setdiff(
  names(gsva_df),
  names(tidybulk::pivot_sample(cd4_se))
),
  exposure     = exp_vars
)

# fit models
gsva_res <- gsva_grid |>
  mutate(
    model = map2(
      gsva_pathway,
      exposure,
      ~ {
        fml <- as.formula(paste(
          "`",
          .x,
          "`~",
          "`", .y, "` + gli_age + gli_sex + gli_height + fis",
          sep = ""
        ))
        glm(fml, data = gsva_df)
      }
    ),
    tidy = map(model, tidy)
  ) |>
  unnest(tidy) |>
  filter(term == exposure) |>
  select(
    pathway = gsva_pathway,
    exposure,
    estimate,
    std.error,
    statistic,
    p.value
  )
```

### Stability

```{r}
# run sensitivity for each pathway
paths <- setdiff(
  names(gsva_df),
  names(tidybulk::pivot_sample(cd4_se))
)
paths <- gsub(" ","_",paths)

colnames(gsva_df) <- gsub(" ","_",colnames(gsva_df))

plan(multisession)

path_stability_res <- future_map(
  paths,
  ~ {
    run_exwas_sensitivity(
      df = gsva_df,
      exposures = exp_vars,
      outcome = .x,
      covariates = c("gli_age", "gli_sex", "gli_height", "fis"),
      covariates_to_remove = c("gli_age", "gli_sex", "gli_height", "fis"),
      family = gaussian(),
      bootstrap_n = 100
    )$sensitivity_df |>
      mutate(
        pathway = .x
      )
  },.options = furrr_options(seed = 123)
) |> 
  bind_rows()

# summarise stability (mirrors your EXWAS logic)
path_stability_df <- path_stability_res |>
  mutate(
    signif = (p_value <= 0.05),
    sign = sign(estimate)
  ) |>
  group_by(exposure, pathway) |>
  summarise(
    n_tests = n(),
    n_signif = sum(signif, na.rm = TRUE),
    prop_signif = n_signif / n_tests,
    median_effect = median(estimate, na.rm = TRUE),
    median_p = median(p_value, na.rm = TRUE),
    sd_effect = sd(estimate, na.rm = TRUE),
    sign_consistency =
      ifelse(all(is.na(sign)), NA_real_,
             abs(sum(sign, na.rm = TRUE)) / n_tests),
    stability_score = prop_signif * sign_consistency,
    .groups = "drop"
  ) |> 
  mutate(pathway = gsub("_"," ",pathway)) |> 
  left_join(
    gsva_res,
    by=c(
      "pathway"="pathway",
      "exposure"="exposure"
    )
  )
```

### * Heatmap

```{r}
path_stability_df |>
  filter(p.value<0.1,
         sign_consistency>0.7) |> 
    mutate(
        exposure=fct_reorder(exposure,estimate),
        pathway=fct_reorder(pathway,estimate),
    ) |>
  ggplot(
        aes(
            x = pathway,
            y = exposure,
            fill = estimate
        )
    ) +
    geom_tile(
        color = "black",
        linewidth = 0.2,
        height = 0.9
    ) +
    scale_fill_gradient2(
                low ="blue4",
                mid="white",
                high = "red4",
                midpoint = 0,
                limits=c(-0.2,.2),
                oob=scales::squish
            )+
    labs(
        x = "",
        y = "",
        fill = "Estimate"
    ) +
    theme_bw() +
    theme(
        # rotate labels
        axis.text.x = element_text(
            angle = 90,
            vjust = 0.5,
            hjust = 1
        ),
        
        # remove x-axis formatting
        axis.title.x = element_blank(),
        axis.line.x  = element_blank(),
        
        axis.text.y = element_text(size = 9),
        legend.position = "right",
        legend.title.position = "top"
    )
# path_stability_df |>
#   filter(p.value<0.1,
#          sign_consistency>0.7) |> 
#     inner_join(
#         path_stability_df |> 
#           filter(p.value<0.1,
#                  sign_consistency>0.7) |> 
#           janitor::tabyl(exposure),
#         by = "exposure"
#     ) |>
#     mutate(
#         exposure_label = if_else(
#             n > 1,
#             paste0(strrep("*", n), " ", exposure),
#             exposure
#         ),
#         exposure_label=fct_reorder(exposure_label,estimate),
#         pathway=fct_reorder(pathway,estimate),
#     ) |>
#   ggplot(
#         aes(
#             x = pathway,
#             y = exposure_label,
#             fill = estimate
#         )
#     ) +
#     geom_tile(
#         color = "black",
#         linewidth = 0.2,
#         height = 0.9
#     ) +
#     scale_fill_gradient2(
#                 low ="blue4",
#                 mid="white",
#                 high = "red4",
#                 midpoint = 0,
#                 limits=c(-0.2,.2),
#                 oob=scales::squish
#             )+
#     labs(
#         x = "",
#         y = "",
#         fill = "Estimate",
#         caption = "* Appears in more than one phenotype."
#     ) +
#     theme_bw() +
#     theme(
#         # rotate labels
#         axis.text.x = element_text(
#             angle = 90,
#             vjust = 0.5,
#             hjust = 1
#         ),
#         
#         # remove x-axis formatting
#         axis.title.x = element_blank(),
#         axis.line.x  = element_blank(),
#         
#         axis.text.y = element_text(size = 9),
#         legend.position = "right",
#         legend.title.position = "top"
#     )
```

### * Exposure Barplot

```{r}
path_stability_df |>
    filter(p.value<0.1,
           sign_consistency>0.7) |>
    group_by(exposure) |> 
    reframe(med=median(estimate)) |>
    ggplot(aes(
        x = med,
        y = fct_reorder(exposure,med),
        fill = med
    )) +
    geom_col(color = "black", 
             linewidth = 0.005) +
    theme_bw() +
    scale_fill_gradient2(
        low ="blue4",
        mid="white",
        high = "red4",
        midpoint = 0
    )+scale_x_continuous(breaks = c(-1,-.5,.1))+
    labs(x = "",
         y = "", 
         fill = "Median Estimate") +
    theme(
        strip.text.x = element_text(
            face = "bold.italic"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor   = element_blank(),
        legend.position = "right",
        legend.title.position = "top"
    )
```
### * Pathway Barplot

```{r}
path_stability_df |>
    filter(p.value<0.1,
           sign_consistency>0.7) |>
  group_by(pathway) |> 
  reframe(med=median(estimate)) |>
  ggplot(aes(
    x = med,
    y = fct_reorder(pathway,med),
    fill = med
  )) +
  geom_col(color = "black", 
           linewidth = 0.005) +
  theme_bw() +
  scale_fill_gradient2(
                low ="blue4",
                mid="white",
                high = "red4",
                midpoint = 0
            )+
  labs(x = "",
       y = "", 
       fill = "Median Estimate") +
  theme(
    strip.text.x = element_text(
      face = "bold.italic"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    legend.position = "bottom",
    legend.title.position = "top"
  )
```

### * Pathway Barplot Thresholds

```{r}

pmap(
  expand_grid(
    consistency = c(0.3, 0.5, 0.7, 0.9),
    p_cutoff    = c(0.1, 0.05)
),
    \(consistency,p_cutoff){
      path_stability_df |>
    filter(p.value<p_cutoff,
           sign_consistency>consistency) |>
    group_by(pathway) |> 
    reframe(med=median(estimate),
            n=n(),
            pathway=paste(pathway," (",n,")",sep="")) |> 
        distinct() |> 
        (\(df){
        df |> 
            ggplot(aes(
                x = med,
                y = fct_reorder(pathway,med),
                fill = med
            )) +
            geom_col(color = "black", 
                     linewidth = 0.005) +
            theme_bw() +
            scale_fill_gradient2(
                low ="blue4",
                mid="white",
                high = "red4",
                midpoint = 0
            )+
        scale_x_continuous(breaks = c(-1,-.5,.1))+
            labs(x = "",title="Cell State-Exposures",
                 y = "", 
                 caption=paste("P <",p_cutoff, "& Effect Consistency >",consistency*100,"%"),
                 fill = "Median Estimate") +
            theme(
                strip.text.x = element_text(
                    face = "bold.italic"),
                panel.grid.major.x = element_blank(),
                panel.grid.minor   = element_blank(),
                legend.position = "right",
                legend.title.position = "top",
                plot.title = element_text(
                  size=12
                )
            ) })()
    }) |> 
  wrap_plots()

    
```

## By Cell Marker - Senescence and Exhaustion/Tolerance

```{r}
gene_sets <- list(
  tcell_exhaustion   = c(
    "TIGIT",
    "PDCD1",
    "CTLA4",
    "LAG3",
    "HAVCR2",
    "ENTPD1",
    "CD244",
    "TOX",
    "TOX2",
    "NR4A1",
    "NR4A2",
    "BATF",
    "IRF4"
  ),
  endotoxin_tolerance = c(
    "LILRB1",
    "LILRB2",
    "LILRB4",
    "CD274",
    "PDCD1LG2",
    "CIITA",
    "HLA-DRA",
    "HLA-DRB1",
    "HLA-DPB1",
    "SIGLEC10",
    "LAIR1",
    "S100A8",
    "S100A9",
    "CTSD"
  ),
  monocyte_senescence         = c(
    "CDKN2A",
    "CDKN1A",
    "TP53",
    "RB1",
    "IL1B",
    "IL6",
    "TNF",
    "CXCL8",
    "CCL2",
    "CCL3",
    "MMP9",
    "TGFB1",
    "SERPINE1"
  ),
  t_cell_senescence = c(
    "CDKN2A",
    "CDKN1A",
    "TP53",
    "RB1",
    "KLRG1",
    "B3GAT1",
    "GZMB",
    "IFNG",
    "TNF"
  ),
  monocyte_activation = c(
  # Immediate early / inflammatory
  "IL1B",
  "TNF",
  "IL6",
  "NFKBIA",
  "JUN",
  "FOS",

  # Antigen presentation / costimulation
  "CD74",
  "HLA-DRA",
  "HLA-DRB1",
  "HLA-DPB1",
  "CIITA",

  # Pattern recognition / activation state
  "TLR2",
  "TLR4",
  "FCGR1A",   # CD64
  "FCGR3A",   # CD16
  "LYZ"
),
  tcell_activation = c(
  # Immediate activation
  "IL2",
  "IL2RA",    # CD25
  "CD69",
  "TNFRSF9",  # 4-1BB
  "TNFRSF4",  # OX40

  # TCR signaling / transcriptional response
  "ZAP70",
  "LCK",
  "ICOS",
  "BATF",
  "IRF4",

  # Effector skew (non-exhaustion)
  "IFNG",
  "GZMB"
)
)

```


###  Summary Df

```{r}
expected_dir <- c(
  CIITA      = -1,
  `HLA-DRA`  = -1,
  `HLA-DRB1` = -1,
  `HLA-DPB1` = -1
)

exp_omics_sens_res_df <- exp_omics_sens_res_df |>
  mutate(
    expected_dir = coalesce(expected_dir[feature_map], 1),
    signed_logFC = logFC * expected_dir
  )

summary_df <- bind_rows(
  summarise_set(
    exp_omics_sens_res_df,
    gene_sets$tcell_exhaustion,
    "CD4",
    "T cell Exhaustion"
  ),
  
  summarise_set(
    exp_omics_sens_res_df,
    gene_sets$endotoxin_tolerance,
    "CD16",
    "Monocyte Tolerance"
  ),
  
  summarise_set(
    exp_omics_sens_res_df |>
      inner_join(
        senescence_gate(exp_omics_sens_res_df, "CD4"),
        by = "exposure"
      ),
    gene_sets$t_cell_senescence,
    "CD4",
    "T cell Senescence"
  ),
  
  summarise_set(
    exp_omics_sens_res_df |>
      inner_join(
        senescence_gate(exp_omics_sens_res_df, "CD16"),
        by = "exposure"
      ),
    gene_sets$monocyte_senescence,
    "CD16",
    "Monocyte Senescence"
  ),
  summarise_set(
    exp_omics_sens_res_df,
    gene_sets$monocyte_activation,
    "CD16",
    "Monocyte Activation"
  ),
  summarise_set(
    exp_omics_sens_res_df,
    gene_sets$tcell_activation,
    "CD4",
    "T cell Activation"
  )
)

```

###  Grouped

```{r}
# # Expected direction map, 1 is up and -1 is down
# 
# expected_dir <- c(
#   CIITA     = -1,
#   `HLA-DRA` = -1,
#   `HLA-DRB1` = -1,
#   `HLA-DPB1` = -1
# )
# 
# # Default direction = +1
# # we are looking for it to go up so we flip the sign
# # for some genes if down is consistent with the 
# # phenotype
# exp_omics_sens_res_df <- exp_omics_sens_res_df |>
#   mutate(
#     expected_dir = ifelse(
#       feature_map %in% names(expected_dir),
#       expected_dir[feature_map],
#       1
#     ),
#     signed_logFC = logFC * expected_dir
#   )
# 
# # CD4 T-cell exhaustion direction-aware
# exhaustion_res <- exp_omics_sens_res_df |>
#   filter(
#     feature_map %in% gene_sets$tcell_exhaustion,
#     grepl("CD4", exp_name)
#   ) |>
#   group_by(exposure) |>
#   reframe(
#     med_signed = median(signed_logFC, 
#                         na.rm = TRUE),
#     frac_expected_dir = mean(signed_logFC > 0,
#                              na.rm = TRUE),
#     n_genes = n_distinct(feature_map)
#   ) |>
#   filter(
#     n_genes > 1,
#     frac_expected_dir >= 0.5,
#     med_signed > 0
#     
#   ) |> 
#   mutate(set="T cell Exhaustion")
# 
# 
# # CD16 monocyte endotoxin tolerance direction-aware
# tolerance_res <- exp_omics_sens_res_df |>
#   filter(
#     feature_map %in% gene_sets$endotoxin_tolerance,
#     grepl("CD16", exp_name)
#   ) |>
#   group_by(exposure) |>
#   reframe(
#     med_signed = median(
#       signed_logFC, na.rm = TRUE),
#     frac_expected_dir = mean(
#       signed_logFC > 0, na.rm = TRUE),
#     n_genes = n_distinct(feature_map)
#   ) |>
#   filter(
#     n_genes > 1,
#     frac_expected_dir >= 0.5,
#     med_signed > 0
#   ) |> 
#   mutate(set="Monocyte Tolerance")
# 
# 
# # CD4 T-cell senescence gated on cell-cycle arrest
# tcell_sen_gate <- exp_omics_sens_res_df |>
#   filter(
#     feature_map %in% c("CDKN2A", "CDKN1A"),
#     grepl("CD4", exp_name)
#   ) |>
#   group_by(exposure) |>
#   reframe(
#     gate_signal = median(logFC, na.rm = TRUE)
#   ) |>
#   filter(gate_signal > 0)
# 
# tcell_senescence_res <- exp_omics_sens_res_df |>
#   filter(
#     feature_map %in% gene_sets$t_cell_senescence,
#     grepl("CD4", exp_name)
#   ) |>
#   inner_join(tcell_sen_gate, by = "exposure") |>
#   group_by(exposure) |>
#   reframe(
#     med_signed = median(
#       signed_logFC, na.rm = TRUE),
#     frac_expected_dir = mean(
#       signed_logFC > 0, na.rm = TRUE),
#     n_genes = n_distinct(feature_map)
#   ) |>
#   filter(
#     n_genes > 1,
#     frac_expected_dir >= 0.5,
#     med_signed > 0
#   ) |> 
#   mutate(set="T cell Senescence")
# 
# # CD16 monocyte senescence gated on cell-cycle arrest
# mono_sen_gate <- exp_omics_sens_res_df |>
#   filter(
#     feature_map %in% c("CDKN2A", "CDKN1A"),
#     grepl("CD16", exp_name)
#   ) |>
#   group_by(exposure) |>
#   reframe(
#     gate_signal = median(logFC, na.rm = TRUE)
#   ) |>
#   filter(gate_signal > 0)
# 
# mono_senescence_res <- exp_omics_sens_res_df |>
#   filter(
#     feature_map %in% gene_sets$monocyte_senescence,
#     grepl("CD16", exp_name)
#   ) |>
#   inner_join(mono_sen_gate, by = "exposure") |>
#   group_by(exposure) |>
#   reframe(
#     med_signed = median(
#       signed_logFC, na.rm = TRUE),
#     frac_expected_dir = mean(
#       signed_logFC > 0, na.rm = TRUE),
#     n_genes = n_distinct(feature_map)
#   ) |>
#   filter(
#     n_genes > 1,
#     frac_expected_dir >= 0.5,
#     med_signed > 0
#   ) |> 
#   mutate(set="Monocyte Senescence")
# 
# summary_df <- bind_rows(
#   exhaustion_res,
#   tolerance_res,
#   tcell_senescence_res,
#   mono_senescence_res)
```



### Heatmap

```{r}
summary_df |> 
    # group_by(exposure) |> 
    # mutate(n_sets=n_distinct(set)) |> 
    # ungroup() |> 
    mutate(
        set=fct_reorder(set,med_signed),
        exposure=fct_reorder(exposure,med_signed)
    ) |> 
    ggplot(aes(
        x=set,
        y=exposure,
        fill=med_signed
    ))+
    geom_tile()+
    theme_bw()+
    # scale_fill_gradient(
    #     low ="#E29587",
    #     high = "red4"
    # )+
  scale_fill_gradientn(
        colors = c(
            "#3a1c71",
            "#d76d77"
        ),
        na.value = "grey80")+
    labs(
        x="",
        y="",
        fill="Median Estimate"
    )+
    theme(
        axis.text.x = element_text(
            angle = 90,hjust = 1,vjust = 0.3)
    )
```

###  Barplot

```{r}
summary_df |>
  (\(df) {
    df |>
      inner_join(
        df |> janitor::tabyl(exposure),
        by = "exposure"
      )
  })() |>
  mutate(
    exposure_label = if_else(
      n > 1,
      paste0(strrep("*", n), " ", exposure),
      exposure
    )
  ) |>
  group_by(set) |>
  arrange(med_signed,
          .by_group = TRUE) |>
  mutate(
    exposure_in_set = paste0(
      exposure_label, "___", set),
    exposure_in_set = factor(
      exposure_in_set, levels =
        unique(exposure_in_set))
  ) |>
  ungroup() |>
  mutate(set=gsub(" Senescence","\nSenescence",set)) |>
  mutate(set=gsub(" Tolerance","\nTolerance",set)) |>
  mutate(set=gsub(" Exhaustion","\nExhaustion",set)) |> 
  ggplot(aes(
    x = med_signed,
    y = exposure_in_set,
    fill = med_signed
  )) +
  geom_col(color = "black", 
           linewidth = 0.005) +
  facet_grid(~ set, 
             scales = "free_y", 
             space = "free_y") +
  scale_y_discrete(
    labels = function(x) sub("___.*$", "", x)) +
  theme_bw() +
  scale_fill_gradientn(
    colors = c("#3a1c71", 
               "#d76d77"),
    na.value = "grey80"
  ) +
  labs(x = "",
       y = "", 
       fill = "Median Estimate",
       caption = "* Appears in more than phenotype.") +
  theme(
    strip.text.x = element_text(
      face = "bold.italic"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    legend.position = "bottom",
    legend.title.position = "top"
  )
```
###  Heatmap

```{r}
summary_df |>
    left_join(
        summary_df |> janitor::tabyl(exposure),
        by = "exposure"
    ) |>
    mutate(
        exposure_label = if_else(
            n > 1,
            paste0(strrep("*", n), " ", exposure),
            exposure
        )
    ) |>
    group_by(set) |>
    arrange(med_signed, .by_group = TRUE) |>
    mutate(
        exposure_label = factor(
            exposure_label,
            levels = unique(exposure_label)
        )
    ) |>
    ungroup() |> ggplot(
        aes(
            x = set,
            y = exposure_label,
            fill = med_signed
        )
    ) +
    geom_tile(
        color = "black",
        linewidth = 0.2,
        height = 0.9
    ) +
    scale_fill_gradientn(
        colors = c("#3a1c71", "#d76d77"),
        na.value = "grey80"
    ) +
    labs(
        x = "",
        y = "",
        fill = "Median Estimate",
        caption = "* Appears in more than one phenotype."
    ) +
    theme_bw() +
    theme(
        # rotate labels
        axis.text.x = element_text(
            angle = 90,
            vjust = 0.5,
            hjust = 1
        ),
        
        # remove x-axis formatting
        axis.title.x = element_blank(),
        axis.line.x  = element_blank(),
        
        axis.text.y = element_text(size = 9),
        legend.position = "right",
        legend.title.position = "top"
    )
```

## Save Data

```{r}
saveRDS(th_stability_res,file="../results/lineage/th_stability_res.rds")

saveRDS(th_stability_df,file="../results/lineage/th_stability_df.rds")

saveRDS(path_stability_res,file="../results/lineage/path_stability_res.rds")

saveRDS(path_stability_df,file="../results/lineage/path_stability_df.rds")
```

