---
title: "Exposome AIRWAYS Analysis"
subtitle: "ExWAS"
author: "Jason Laird"
date: today 
date-format: short
engine: knitr
execute: 
  warning: false
  message: false
  cache: true
format: 
  html:
    code-fold: true
    theme: 
      - simplex
      - custom.scss
    fontcolor: '#0c051a'
---

## Setup

### Load Libraries

```{r}
#| warning: false
#| message: false

library(tidyverse)
library(janitor)
library(ggpubr)
library(tidyexposomics)
library(MultiAssayExperiment)
library(biostats)
library(glmnet)

invisible(lapply(
  list.files(path = "./",
             pattern="*.R",full.names = TRUE),
  source))
```

### Load Data

```{r}
# MultiAssayExperiment Object
expom <- readRDS("../results/qc/expom.rds")

# Codebook - Filtered to include variables
# retained in MultiAssayExperiment Object
aw_cb <- readRDS("../results/input_data/aw_cb.rds") |> 
  filter(variable %in% colnames(colData(expom)))

# Exposure variables
exp_vars <- readRDS("../results/qc/exp_vars.rds")
num_exp_vars <- readRDS("../results/qc/num_exp_vars.rds")

# All variables
vars <- aw_cb$variable |> 
  (\(chr)chr[chr %in% colnames(colData(expom))])()

# Setting colors for exposure categories
cat_colors <- scales::alpha(get_palette("aaas",13), 0.5) |> 
  (\(colors){
    names(colors) <- aw_cb |> 
      filter(variable %in% exp_vars) |> 
      pull(category) |> 
      unique()
    colors
  })()
```

## Sample Clustering

```{r}
#| fig-width: 10
#| fig-height: 8


# Sample clustering
expom <- expom |>
    run_cluster_samples(
        exposure_cols = num_exp_vars,
        clustering_approach = "elbow",
        action = "add"
    )

expom |>
    plot_sample_clusters(
        exposure_cols = num_exp_vars
    )
```

## Exposure Correlations

```{r}
#| fig-width: 10
#| fig-height: 10

num_vars <- expom@colData |>
  as_tibble() |>
  select_if(is.numeric) |> 
  (\(df){df[,colnames(df) %in% vars]})() |> 
  colnames()

expom <- expom |>
    run_correlation(
        feature_type = "exposures",
        action = "add",
        exposure_cols = num_vars,
        correlation_cutoff = 0.3
    )

expom |>
    plot_circos_correlation(
        feature_type = "exposures",
        corr_threshold = 0.5,
        exposure_cols = num_vars,
        annotation_colors = get_palette("jco",k=42) |> 
          (\(colors){
            names(colors) <- unique(aw_cb$category)
            colors
          })()
        
    )
```

### Exposure Correlation Circos

```{r}
cors <- expom |>
    run_correlation(
        feature_type = "exposures",
        action = "get",
        exposure_cols = exp_vars,
        correlation_cutoff = 0,
        pval_cutoff = 1
    )


cor_g <- cors |>
    filter(
        p.value < 0.05,
        abs(correlation) > 0.3
    ) |>
    tidygraph::as_tbl_graph(directed = FALSE) |>
    tidygraph::activate(nodes) |>
    mutate(
        community = tidygraph::group_louvain(
            weights = abs(correlation)
        ),
        strength = tidygraph::centrality_degree(
            weights = abs(correlation)
        ),
        top_lab =if_else(strength > 10, name, NA_character_)
    ) |> 
    left_join(aw_cb,
              by=c("name"="variable")) |> 
  tidygraph::rename(Category=category)

cor_g <- cor_g |> 
    group_by(Category) |> 
    arrange(desc(strength), .by_group = TRUE) |>
    mutate(
        top_lab = if_else(name %in% c(
            cor_g |>
                as.data.frame() |> 
                group_by(Category) |> 
                arrange(desc(strength),
                        .by_group = TRUE) |> 
                tidygraph::slice_head(n=2) |>
                pull(name)
        ), name, NA_character_)
    ) |>
    ungroup()



cor_g |>
    ggraph::ggraph(
        layout = "fr"
    ) +
    ggraph::geom_edge_link(
        aes(
            color = correlation,
            width = abs(correlation)*5
        )
    ) +
    ggraph::geom_node_point(aes(fill=Category,size=strength),
                            shape = 21,
                            #size = 4,
                            alpha = 0.8
    ) +
    scale_fill_manual(values=cat_colors)+
    ggraph::geom_node_label(
        aes(
            label = top_lab
        ),
        fontface = "bold.italic",repel = T,
        na.rm = TRUE
    ) +
    ggraph::theme_graph() +
    ggraph::scale_edge_color_gradient2(
        high = "#C33764",
        mid  = "white",
        low  = "#1D2671",
        midpoint = 0.1,
        guide = ggraph::guide_edge_colorbar()
    ) +
    guides(
        fill = guide_legend(
            override.aes = list(size = 4)
        ),size = "none",
        edge_width = "none"
    )+
    ggplot2::labs(edge_color = "Correlation")

```


### Exposure Corr. Heatmap

```{r}
cors |>
    select(var1, var2, correlation) |>
    
    # --- Build correlation matrix ---
    pivot_wider(names_from = var2, values_from = correlation) |>
    column_to_rownames("var1") |>
    (\(mat){ mat[is.na(mat)] <- 0; mat })() |>
    
    # --- Wrap into pheatmap call ---
    (\(mat){
        
        # Row annotation (aw_cb$type mapped to matching variables)
        row_ann <- aw_cb |>
            select(variable, category) |>
            filter(variable %in% rownames(mat)) |>
            distinct() |>
            column_to_rownames("variable")
        
        # Column annotation
        col_ann <- aw_cb |>
            select(variable, category) |>
            filter(variable %in% colnames(mat)) |>
            distinct() |>
            column_to_rownames("variable")
        
        # Colors for main heatmap
        heat_cols <- colorRampPalette(c("#1d2671","white","#c33764"))(100)
        
        # Category colors for annotations
        cats <- unique(c(row_ann$category, col_ann$category))
        ann_cols <- list(
            # category = setNames(
            #     scales::alpha(get_palette("aaas",length(cats)), 0.5),
            #     cats
            # )
            category = cat_colors
        )
        
        pheatmap::pheatmap(
            mat,
            color = heat_cols,
            breaks = seq(-1, 1, length.out = 101),
            border_color = NA,
            annotation_row = row_ann,
            annotation_col = col_ann,
            annotation_colors = ann_cols,
            fontsize = 9
        )
    })()

```

### Category Correlations

```{r}
cor_sum_df <- cors |>
  select(var1, var2, correlation) |>

  # --- Bring categories into long form first ---
  left_join(aw_cb |> select(variable, category),
            by = c("var1" = "variable")) |>
  dplyr::rename(cat1 = category) |>
  left_join(aw_cb |> select(variable, category),
            by = c("var2" = "variable")) |>
  dplyr::rename(cat2 = category) |>

  # --- Drop rows without category info ---
  filter(!is.na(cat1), !is.na(cat2)) |>

  # --- Label within vs. between correlation types ---
  mutate(
    corr_type = if_else(cat1 == cat2, "Within-category", "Between-category"),
    pair_category = if_else(cat1 == cat2, cat1, "Between")
  ) 


cor_sum_df |> 
  mutate(combo=paste(cat1,cat2)) |> 
  group_by(combo) |> 
  reframe(med_cor=median(correlation),cat1,cat2) |>
  distinct() |>
  dplyr::select(cat1,cat2,med_cor) |> 
    tidygraph::as_tbl_graph() |> 
    ggraph::ggraph(
        layout = "linear",
        circular = TRUE
    ) +
    ggraph::geom_edge_arc(aes(
        color = med_cor,
        width = abs(med_cor)*10
    )) +
    ggraph::geom_node_point(shape = 21,
                            size = 4,
                            alpha = 0.8
    ) +
    ggraph::geom_node_text(
        aes(
            label = name,
            x = 1.1 * x,
            y = 1.1 * y,
            angle = ggraph::node_angle(x, y)
        ),
        hjust = "outward",
        fontface = "bold.italic",
        check_overlap = TRUE
    ) +
    ggraph::theme_graph() +
    ggraph::scale_edge_color_gradient2(
        high = "#C33764",
        mid = "white",
        low = "#1D2671",
        midpoint = 0.1,
        guide = ggraph::guide_edge_colorbar()
    ) +
    ggplot2::coord_fixed(xlim = c(-2, 2), ylim = c(-2, 2)) +
    ggplot2::guides(edge_width = "none") +
    ggplot2::labs(
        edge_color = "Median Correlation"
    )
```


### Within Category Correlations

```{r}
cor_sum_df |>
    filter(corr_type == "Within-category") |>
    ggplot(aes(x = reorder(cat1,correlation), 
               y = correlation,
               fill = cat1)) +
    geom_boxplot(alpha = 0.7,linewidth=0.005) +
    geom_jitter(aes(color=cat1),alpha=.1)+
    theme_classic(base_size = 12) +
    scale_fill_manual(values=cat_colors)+
    scale_color_manual(values=cat_colors)+
    labs(title = "",
         x = "",
         y = "Correlation") +
    guides(color="none",
           fill="none")+geom_hline(yintercept = 0,linetype="dashed",color="grey25")+
    theme(axis.text.x = element_text(angle = 45, hjust = 1))


cor_sum_df |>
    filter(corr_type == "Within-category") |>
    ggplot(aes(y = reorder(cat1,correlation), 
               x = correlation,
               fill = cat1)) +
    geom_boxplot(alpha = 0.7,linewidth=0.005) +
    geom_jitter(aes(color=cat1),alpha=.1)+
    theme_classic(base_size = 12) +
    scale_fill_manual(values=cat_colors)+
    scale_color_manual(values=cat_colors)+
    labs(title = "",
         y = "",
         x = "Correlation") +
    guides(color="none",
           fill="none")+
  geom_vline(xintercept = 0,
             linetype="dashed",
             color="grey25")+
    theme(axis.text.y = element_text(angle = 45, hjust = 1))
```


### Variable Correlations with FEV1/FVC

```{r}
#| fig-width: 4
#| fig-height: 3
extract_results(expom,result = "correlation") |>
    pluck("exposures") |> 
    filter(var1=="pftfev1fvc_actual",
           !grepl("^pft",var2)) |> 
    ggplot(aes(x = correlation, 
               y = reorder(var2,correlation),
               fill = correlation)) +
    geom_bar(stat = "identity",color="black",linewidth = 0.005)+
    scale_fill_gradient2(
        low = "blue4", 
        mid = "white",
        high = "red4",
        midpoint = 0
    ) +
    labs(
        x = "Correlation",
        y = NULL,
        fill = "Correlation"
    ) +
    theme_classic() +
    theme(
        legend.position = "right"
    )
      
```
### PM10 Correlations

```{r}
#| fig-width: 5
#| fig-height: 7
extract_results(expom,result = "correlation") |>
    pluck("exposures") |> 
    filter(var1=="pm10f",
           !grepl("^pft",var2)) |> 
    ggplot(aes(x = correlation, 
               y = reorder(var2,correlation),
               fill = correlation)) +
    geom_bar(stat = "identity",color="black",linewidth = 0.005)+
    scale_fill_gradient2(
        low = "blue4", 
        mid = "white",
        high = "red4",
        midpoint = 0
    ) +
    labs(
        x = "Correlation",
        y = NULL,
        fill = "Correlation"
    ) +
    theme_classic() +
    theme(
        legend.position = "right"
    )
```
## Cohort Characterization

### Table 1 - Cohort

```{r}
library(gtsummary)

expom |> 
    tidyexposomics::pivot_sample() |>
  mutate(gli_sex=case_when(
    gli_sex=="M" ~ "Male",
    gli_sex=="F" ~ "Female"
  ),
  black = case_when(
    black == "black" ~ "Black",
    black == "non-black" ~ "Non-Black"
  ),
  income5 = case_when(
    income5 == "RA/DK" ~ "No Response",
    .default = income5
  )) |> 
    transmute(
        Sex     = factor(gli_sex),
        `Age (yrs)`     = as.numeric(gli_age),
        `Height (cm)`  = as.numeric(gli_height),
        Ancestry       = factor(black, 
                                levels = c("Non-Black", "Black")),
        `FEV1/FVC` = pftfev1fvc_actual,
        BMI     = factor(bmicat4),
        `Household Income` = income5
    ) |> tbl_summary(
    statistic = list(
        all_continuous() ~ "{mean} ({sd})",
        all_categorical() ~ "{n} ({p}%)"
    ),
    digits = all_continuous() ~ 1,
    missing = "no"
) |> 
    modify_header(label = "**Characteristic**") |> 
    bold_labels()


```
### Table 2 - Difference in FEV1/FVC

```{r}
lm(
    `FEV1/FVC` ~ Sex + Ancestry + BMI + `Household Income` + `Age (yrs)` + `Height (cm)`,
    data = expom |> 
        tidyexposomics::pivot_sample() |>
        mutate(gli_sex=case_when(
            gli_sex=="M" ~ "Male",
            gli_sex=="F" ~ "Female"
        ),
        black = case_when(
            black == "black" ~ "Black",
            black == "non-black" ~ "Non-Black"
        ),
        income5 = case_when(
            income5 == "RA/DK" ~ "No Response",
            .default = income5
        )) |> 
        transmute(
            Sex     = factor(gli_sex),
            `Age (yrs)`     = as.numeric(gli_age),
            `Height (cm)`  = as.numeric(gli_height),
            Ancestry       = factor(black, 
                                    levels = c("Non-Black", "Black")),
            `FEV1/FVC` = pftfev1fvc_actual,
            BMI     = factor(bmicat4),
            `Household Income` = income5
        )
) |> 
  bold_labels() |> 
    gtsummary::tbl_regression()
```


## Cell Type Analysis

### CBC Differential

```{r}
# Perform ExWAS Analysis
cell_res <- expom |>
    run_association(
        source = "exposures",
        outcome = "pftfev1fvc_actual",
        feature_set = aw_cb |>
          filter(category=="CBC Differential") |> 
          pull(variable),
        covariates = c(
          "gli_age",
          "gli_sex",
          "gli_height",
          "fis"
        ),
        action = "get",
        family = "gaussian"
    ) |> 
  pluck("results_df") |> 
  filter(term %in% c(
    aw_cb |>
          filter(category=="CBC Differential") |> 
          pull(variable)
  ))

cell_res |> 
  # Order categories by median effect size
    mutate(
        sig = p.value< 0.05,
        term = case_when(
          sig == TRUE ~ paste0("* ",definition),
          .default =definition),
        # Order terms by estimate (within categories)
        term = forcats::fct_reorder(term, estimate)
    ) |> 
    ggplot(aes(x = estimate, 
               y = term,
               color = estimate,
               fill = estimate)) +
    geom_point(shape = 21, 
               size = 2.5, 
               stroke = 0.8) +
    geom_errorbarh(aes(xmin = estimate - 1.96 * std.error,
                       xmax = estimate + 1.96 * std.error),
                   height = 0.15) +
    scale_color_gradient2(
        low="blue4",
        mid="grey85",
        high="red4",
        midpoint = 0
    )+
    scale_fill_gradient2(
        low="blue4",
        mid="grey85",
        high="red4",
        midpoint = 0
    )+
    geom_vline(xintercept = 0, linetype = "dashed") +
    labs(
        title = NULL,
        x = "Effect size",
        y = "",
        fill ="Effect size",
        color ="Effect size"
    ) +
    theme_bw(base_size = 13) +
    theme(panel.grid.minor = element_blank(),
          strip.text.y = element_text(angle=0,
                                      face="bold.italic"))
```
 

### Cell/Cytokines

```{r}
# Perform ExWAS Analysis
cell_cyto_res <- expom |>
    run_association(
        source = "exposures",
        outcome = "pftfev1fvc_actual",
        feature_set = aw_cb |>
    filter(grepl(
      "Serum Cytokines|Nasal Cytokines|CBC Differential",category)) |> 
    filter(variable %in% colnames(tidyexposomics::pivot_sample(expom))) |> 
    pull(variable),
        covariates = c(
          "gli_age",
          "gli_sex",
          "gli_height",
          "fis"
        ),
        action = "get",
        family = "gaussian"
    ) |> 
  pluck("results_df") |> 
  filter(term %in% c(
    aw_cb |>
    filter(grepl(
      "Serum Cytokines|Nasal Cytokines|CBC Differential",category)) |> 
    filter(variable %in% colnames(tidyexposomics::pivot_sample(expom))) |> 
    pull(variable)
  ))

cell_cyto_res |> 
  # Order categories by median effect size
    mutate(
        sig = p.value< 0.05,
        term = case_when(
          sig == TRUE ~ paste0("* ",term),
          .default =term),
        # Order terms by estimate (within categories)
        term = forcats::fct_reorder(term, estimate),
        category = forcats::fct_reorder(
            category, estimate, .fun = median, .desc = TRUE),
    ) |> 
    ggplot(aes(x = estimate, 
               y = term,
               color = estimate,
               fill = estimate)) +
    geom_point(shape = 21, 
               size = 2.5, 
               stroke = 0.8) +
    geom_errorbarh(aes(xmin = estimate - 1.96 * std.error,
                       xmax = estimate + 1.96 * std.error),
                   height = 0.15) +
  ggh4x::facet_grid2(category ~ ., 
                       scales = "free",
                       space = "free") +
    scale_color_gradient2(
        low="blue4",
        mid="grey85",
        high="red4",
        midpoint = 0
    )+
    scale_fill_gradient2(
        low="blue4",
        mid="grey85",
        high="red4",
        midpoint = 0
    )+
    geom_vline(xintercept = 0, linetype = "dashed") +
    labs(
        title = NULL,
        x = "Effect size",
        y = "",
        fill ="Effect size",
        color ="Effect size"
    ) +
    theme_bw(base_size = 13) +
    theme(panel.grid.minor = element_blank(),
          strip.text.y = element_text(angle=0,
                                      face="bold.italic"))
```


## ExWAS

### Covariate Inclusion

```{r}
covars <- covar_screen(
  expom |> 
    tidyexposomics::pivot_sample(),
  outcome = "pftfev1fvc_actual",
  base_covars = c(
    "gli_age",
    "gli_sex"),
  candidate_covars = c(
    "fis",
    "black",
    "gli_height",
    "bmizscore",
    "somecol",
    "income5",
    "season",
    "med_ics")
  # candidate_covars=vars[!vars %in% c(exp_vars,"gli_age","gli_sex")]
  
)

covars$recommended_covariates
```

### FEV1 / FVC

Measuring airway obstruction

```{r}
#| fig-height: 12
#| fig-width: 5


# Perform ExWAS Analysis
fev1_fvc_expom <- expom |>
    run_association(
        source = "exposures",
        outcome = "pftfev1fvc_actual",
        feature_set = exp_vars,
        covariates = c(
          "gli_age",
          "gli_sex",
          "gli_height",
          "fis"
        ),
        action = "add",
        family = "gaussian"
    )

fev1_fvc_expom |>
    plot_association(
        subtitle = paste(
          "Covariates: ",
          "Age, Sex, Height, Food Insecurity"
        ,sep = "\n"),
        source = "exposures",
        terms = exp_vars,
        filter_thresh = 0.05,
        filter_col = "p.value",
        r2_col = "adj_r2"
    )+
  labs(
    caption="P < 0.05",
    title=NULL
  )+
  theme(plot.caption = element_text(face="italic"))
```
### Manhattan Plot 

```{r}
#| fig-width: 8
#| fig-height: 7
plot_manhattan(fev1_fvc_expom,
               facet_angle = 0,
               facet_cols = cat_colors)
```

### Group Effects

```{r}
#| fig-width: 8
#| fig-height: 10
assoc_df <- fev1_fvc_expom |> 
  extract_results(result = "association") |> 
  pluck("assoc_exposures",
        "results_df") |> 
  # Order categories by median effect size
    mutate(
        sig = p.value< 0.05,
        term = case_when(
          sig == TRUE ~ paste0("* ",term),
          .default =term),
        # Order terms by estimate (within categories)
        term = forcats::fct_reorder(term, estimate),
        category = forcats::fct_reorder(
            category, estimate, .fun = median, .desc = TRUE),
    ) 
    # # Order categories by median effect size
    # mutate(
    #     category = forcats::fct_reorder(
    #         category, estimate, .fun = median, .desc = TRUE),
    #     
    #     # Order terms by estimate (within categories)
    #     term = forcats::fct_reorder(term, estimate),
    #     
    #     # Significance flag for fill aesthetics
    #     sig = p_adjust < 0.05
    # ) 
assoc_df |> 
    ggplot(aes(x = estimate, 
               y = term,
               color = estimate,
               fill = estimate)) +
    geom_point(shape = 21, 
               size = 2.5, 
               stroke = 0.8) +
    geom_errorbarh(aes(xmin = estimate - 1.96 * std.error,
                       xmax = estimate + 1.96 * std.error),
                   height = 0.15) +
    #facet_grid(category~., scales = "free",space = "free") +
    ggh4x::facet_grid2(category ~ ., 
                       scales = "free",
                       space = "free", 
                       strip = ggh4x::strip_themed(
                           background_y = ggh4x::elem_list_rect(
                               fill = cat_colors[levels(assoc_df$category)]))) +
    # ggh4x::force_panelsizes(rows = c(
    #   2,2,2,2,2,1,1,2,1,1,3,3,3
    # )) + 
    scale_color_gradient2(
        low="blue4",
        mid="grey85",
        high="red4",
        midpoint = 0
    )+
    scale_fill_gradient2(
        low="blue4",
        mid="grey85",
        high="red4",
        midpoint = 0
    )+
    geom_vline(xintercept = 0, linetype = "dashed") +
    labs(
        title = NULL,
        x = "Effect size",
        y = "",
        fill ="Effect size",
        color ="Effect size"
    ) +
    theme_bw(base_size = 13) +
    theme(panel.grid.minor = element_blank(),
          strip.text.y = element_text(angle=0,
                                      face="bold.italic"))
```

### Stability

```{r}
exwas_stability_res <- run_exwas_sensitivity(
    df = tidyexposomics::pivot_sample(fev1_fvc_expom),
    exposures = exp_vars,
    outcome = "pftfev1fvc_actual",
    covariates = c("gli_age" , "gli_sex" , "gli_height" , "fis"),
    covariates_to_remove = c("gli_age" , "gli_sex" , "gli_height" , "fis"),
    family = gaussian(),
    bootstrap_n = 100
)

sens_df <- exwas_stability_res$sensitivity_df |>
    mutate(
        signif = (p_value <= 0.05) ,
        sign = sign(estimate)
    ) %>%
    group_by(exposure) %>%
    summarise(
        n_tests = n(),
        n_signif = sum(signif, na.rm = TRUE),
        prop_signif = n_signif / n_tests,
        median_effect = median(estimate, na.rm = TRUE),
        median_p=median(p_value, na.rm = TRUE),
        sd_effect = sd(estimate, na.rm = TRUE),
        sign_consistency =
            ifelse(all(is.na(sign)), NA_real_,
                   abs(sum(sign, na.rm = TRUE)) / n_tests),
        stability_score = prop_signif * sign_consistency,
        .groups = "drop"
    ) |> 
  left_join(
    aw_cb,
    by=c("exposure"="variable")
  )


```

#### Forest Plot

```{r}
sens_assoc_df <- sens_df |> 
    # Order categories by median effect size
    mutate(
        category = forcats::fct_reorder(
            category, median_effect, .fun = median, .desc = TRUE),
        
        # Order terms by estimate (within categories)
        exposure = forcats::fct_reorder(exposure, median_effect),
        
        # Significance flag for fill aesthetics
        sig = median_p < 0.05
    ) 
sens_assoc_df |> 
    ggplot(aes(x = median_effect, 
               y = exposure,
               color = median_effect,
               fill = median_effect)) +
    geom_point(shape = 21, 
               size = 2.5, 
               stroke = 0.8) +
    geom_errorbarh(aes(xmin = median_effect - 1.96 * sd_effect,
                       xmax = median_effect + 1.96 * sd_effect),
                   height = 0.15) +
    #facet_grid(category~., scales = "free",space = "free") +
    ggh4x::facet_grid2(category ~ ., 
                       scales = "free",
                       space = "free", 
                       strip = ggh4x::strip_themed(
                           background_y = ggh4x::elem_list_rect(
                               fill = cat_colors[levels(sens_assoc_df$category)]))) +
    # ggh4x::force_panelsizes(rows = c(
    #   2,2,2,2,2,1,1,2,1,1,3,3,3
    # )) + 
    scale_color_gradient2(
        low="blue4",
        mid="grey85",
        high="red4",
        midpoint = 0
    )+
    scale_fill_gradient2(
        low="blue4",
        mid="grey85",
        high="red4",
        midpoint = 0
    )+
    geom_vline(xintercept = 0, linetype = "dashed") +
    labs(
        title = NULL,
        x = "Effect size",
        y = "",
        fill ="Effect size",
        color ="Effect size"
    ) +
    theme_bw(base_size = 13) +
    theme(panel.grid.minor = element_blank(),
          strip.text.y = element_text(angle=0,
                                      face="bold.italic"))
```
#### Effect Consistency 

```{r}
sens_assoc_df |>
    # mutate(exposure_ord = fct_reorder(exposure, sign_consistency)) |>
  mutate(
    exp_match = ifelse()
    exposure_ord = fct_reorder(exposure,
                                    )) |> 
    ggplot(aes(
        y = exposure_ord,
        x = sign_consistency,
        color = median_effect,
        fill = median_effect
    )) +
    # stem
    geom_segment(
        aes(
            x = 0,
            xend = sign_consistency,
            yend = exposure_ord
        ),
        linewidth = 0.5
    ) +
    # head
    geom_point(
        shape = 21,
        size = 3,
        stroke = 0.9
    ) +
    scale_color_gradient2(
        low = "blue4",
        mid = "grey85",
        high = "red4",
        midpoint = 0,
        limits = c(-0.15, 0.15),
        oob = scales::squish
    ) +
    scale_fill_gradient2(
        low = "blue4",
        mid = "grey85",
        high = "red4",
        midpoint = 0,
        limits = c(-0.15, 0.15),
        oob = scales::squish
    ) +
    labs(
        x = "Effect consistency",
        y = NULL,
        color = "Effect size",
        fill = "Effect size"
    ) +
    theme_bw(base_size = 13) +
    theme(
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank()
    )

```

#### Median Effect Size

```{r}
sens_assoc_df |> 
    group_by(category) |> 
    mutate(med_cat=mode(median_effect)) |> 
    ungroup() |> 
    ggplot(aes(
        x = reorder(category,med_cat),
        y = median_effect,
        fill = category,
        color = category)) +
    geom_boxplot(outlier.shape = NA, 
                 alpha = 0.8,
                 color = "black",
                 linewidth=0.005) +
    geom_jitter(alpha=0.1)+
    labs(
        y = "Median Estimate",
        x = "",
        title = ""
    ) +
    scale_fill_manual(values=cat_colors)+
    scale_color_manual(values=cat_colors)+
    theme_classic(base_size = 12) +
    theme(legend.position = "none")+
    rotate_x_text(angle = 65)+
    ylim(c(-0.04,0.04))+
    geom_hline(yintercept = 0,
               linetype="dashed")
```

#### Sign Flip Change

```{r}
sens_assoc_df |> 
    group_by(category) |> 
    mutate(med_cat=median(sign_consistency)) |> 
    ungroup() |> 
    ggplot(aes(
        x = reorder(category,med_cat),
        y = sign_consistency,
        fill = category,
        color = category)) +
    geom_boxplot(outlier.shape = NA, 
                 alpha = 0.8,
                 color = "black",
                 linewidth=0.005) +
    geom_jitter(alpha=0.1)+
    labs(
        y = "Median Sign Consistency",
        x = "",
        title = ""
    ) +
    scale_fill_manual(values=cat_colors)+
    scale_color_manual(values=cat_colors)+
    theme_classic(base_size = 12) +
    theme(legend.position = "none")+
    rotate_x_text(angle = 65)
    # geom_hline(yintercept = 0.5,
    #            linetype="dashed")
```

#### Proportion Significant 

```{r}

```


### Sample Size

```{r}
assoc_df <- expom |>
    run_association(
        source = "exposures",
        outcome = "pftfev1fvc_actual",
        feature_set = exp_vars,
        action = "get",
        family = "gaussian"
    ) |> 
  pluck("results_df") 
  # df = n - (exposure + covars + intercept)
  # mutate(df = nrow(tidyexposomics::pivot_sample()(expom)) - 2) 
assoc_df <- fev1_fvc_expom@metadata$association$assoc_exposures$results_df

sample_size_results <- bind_rows(
  calc_sample_sizes(
  assoc_df=assoc_df,
  sd_y = 1,
  alpha = 0.05, 
  power = 0.70) |> 
    mutate(power="0.70"),
  calc_sample_sizes(
  assoc_df=assoc_df,
  sd_y = 1,
  alpha = 0.05, 
  power = 0.80) |> 
    mutate(power="0.80"),
  calc_sample_sizes(
  assoc_df=assoc_df,
  sd_y = 1,
  alpha = 0.05, 
  power = 0.90) |> 
    mutate(power="0.90")
)

sample_size_results |> 
  ggplot(aes(
    x = required_n,
    y = abs(beta),
    color = power
  ))+
  geom_point(alpha=0.7)+
  geom_line(alpha=0.5)+
  scale_x_log10(labels = scales::label_number(big.mark = ",")) +
  theme_classic()+
  scale_color_manual(values=c(
    "#483d5b",
    "#855174",
    "#c3687a"
  ))+
  labs(
    x="Sample Size",
    y="Abs(Estimate)",
    color="Power"
  )
```

## Elastic Net

```{r}
set.seed(42)
enet_res <-
  fev1_fvc_expom |>
  tidyexposomics::pivot_sample() |>
  select(pftfev1fvc_actual, 
         all_of(exp_vars), 
         all_of(c("gli_age","gli_sex","gli_height","fis"))) |>
  na.omit() |>
  (\(df) {
    
    X <- model.matrix(
      ~ . - pftfev1fvc_actual - 1,
      data = df
    )
    
    y <- df$pftfev1fvc_actual
    
    penalty <- ifelse(colnames(X) %in% exp_vars, 1, 0)
    
    cv.glmnet(
      x = X,
      y = y,
      family = "gaussian",
      alpha = 0.5,                 # elastic net
      penalty.factor = penalty,
      standardize = TRUE,
      nfolds = 10
    )
  })()

enet_tbl <-
  broom::tidy(enet_res$glmnet.fit) |>
  filter(lambda==enet_res$lambda.1se)
```

### * Forest Plot

```{r}
combo_assoc_df |>
    ggplot(aes(x = enet_estimate,
               y = term,
               color = enet_estimate)) +
    geom_segment(
        aes(x = 0, xend = enet_estimate,
            yend = term),
        linewidth = 0.8
    ) +
    geom_point(size = 2.5) +
    ggh4x::facet_grid2(
        category ~ .,
        scales = "free",
        space = "free",
        strip = ggh4x::strip_themed(
            background_y = ggh4x::elem_list_rect(
                fill = cat_colors[levels(assoc_df$category)]
            )
        )
    ) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    scale_color_gradient2(
        low = "blue4",
        mid = "grey85",
        high = "red4",
        midpoint = 0,
        limits = c(-0.15, 0.15),
        oob = scales::squish
    ) +
    scale_fill_gradient2(
        low = "blue4",
        mid = "grey85",
        high = "red4",
        midpoint = 0,
        limits = c(-0.15, 0.15),
        oob = scales::squish
    ) +
    labs(x = "Elastic-net coefficient", y = NULL,,color="Effect Size") +
    theme_bw(base_size = 13)+
    theme(panel.grid.minor = element_blank(),
          strip.text.y = element_text(angle = 0, face = "bold.italic"),
          legend.position = "right")
```


### * Aggregate Analysis

```{r}
combo_assoc_df <- fev1_fvc_expom |> 
  extract_results(result = "association") |> 
  pluck("assoc_exposures",
        "results_df") |> 
  dplyr::select(term,
                estimate,
                std.error,
                p.value,
                category) |> 
  left_join(sens_df |> 
              dplyr::select(exposure,
                            median_effect,
                            sign_consistency),
            by=c("term"="exposure")) |> 
  left_join(broom::tidy(enet_res$glmnet.fit) |>
              filter(lambda == enet_res$lambda.min,
                     term != "(Intercept)") |> 
              dplyr::select(term,
                            enet_estimate=estimate)) |> 
  mutate(
    sig=p.value<0.05,
    term = case_when(
      sig == TRUE ~ paste0("* ",term),
      .default =term),
    term = forcats::fct_reorder(term, estimate),
    category = forcats::fct_reorder(
      category, estimate, .fun = median, .desc = TRUE)
  )

p_assoc <-
  combo_assoc_df |>
  ggplot(aes(x = estimate,
             y = term,
             color = estimate,
             fill = estimate)) +
  geom_point(shape = 21, size = 2.5, stroke = 0.8) +
  geom_errorbarh(
    aes(xmin = estimate - 1.96 * std.error,
        xmax = estimate + 1.96 * std.error),
    height = 0.15
  ) +
  ggh4x::facet_grid2(
    category ~ .,
    scales = "free",
    space = "free",
    strip = ggh4x::strip_themed(
      background_y = ggh4x::elem_list_rect(
        fill = cat_colors[levels(assoc_df$category)]
      )
    )
  ) +
  scale_color_gradient2(
    low = "blue4", mid = "grey85", high = "red4", midpoint = 0
  ) +
  scale_fill_gradient2(
    low = "blue4", mid = "grey85", high = "red4", midpoint = 0
  ) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = "Effect size", y = NULL,color="Effect Size") +
  theme_bw(base_size = 13) +
  theme(panel.grid.minor = element_blank(),
        strip.text.y = element_text(angle = 0, face = "bold.italic"),
        legend.position = "bottom")

p_stability <-
  combo_assoc_df |>
  ggplot(aes(x = sign_consistency,
             y = term,
             color = median_effect)) +
  geom_segment(
    aes(x = 0, xend = sign_consistency,
        yend = term),
    linewidth = 0.8
  ) +
  geom_point(size = 2.5) +
  ggh4x::facet_grid2(
    category ~ .,
    scales = "free",
    space = "free",
    strip = ggh4x::strip_themed(
      background_y = ggh4x::elem_list_rect(
        fill = cat_colors[levels(assoc_df$category)]
      )
    )
  ) +
  scale_color_gradient2(
        low = "blue4",
        mid = "grey85",
        high = "red4",
        midpoint = 0,
        limits = c(-0.15, 0.15),
        oob = scales::squish
    ) +
    scale_fill_gradient2(
        low = "blue4",
        mid = "grey85",
        high = "red4",
        midpoint = 0,
        limits = c(-0.15, 0.15),
        oob = scales::squish
    ) +
  labs(x = "Effect consistency", y = NULL, color = "Median Effect") +
  theme_bw(base_size = 13)+
  theme(panel.grid.minor = element_blank(),
        strip.text.y = element_text(angle = 0, face = "bold.italic"),
        legend.position = "bottom")

p_enet <-
  combo_assoc_df |>
  ggplot(aes(x = enet_estimate,
             y = term,
             color = enet_estimate)) +
  geom_segment(
    aes(x = 0, xend = enet_estimate,
        yend = term),
    linewidth = 0.8
  ) +
  geom_point(size = 2.5) +
  ggh4x::facet_grid2(
    category ~ .,
    scales = "free",
    space = "free",
    strip = ggh4x::strip_themed(
      background_y = ggh4x::elem_list_rect(
        fill = cat_colors[levels(assoc_df$category)]
      )
    )
  ) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_color_gradient2(
        low = "blue4",
        mid = "grey85",
        high = "red4",
        midpoint = 0,
        limits = c(-0.15, 0.15),
        oob = scales::squish
    ) +
    scale_fill_gradient2(
        low = "blue4",
        mid = "grey85",
        high = "red4",
        midpoint = 0,
        limits = c(-0.15, 0.15),
        oob = scales::squish
    ) +
  labs(x = "Elastic-net coefficient", y = NULL,,color="Effect Size") +
  theme_bw(base_size = 13)+
  theme(panel.grid.minor = element_blank(),
        strip.text.y = element_text(angle = 0, face = "bold.italic"),
        legend.position = "bottom")

p_assoc | p_stability | p_enet
```

### * Aggregate Plot 2

```{r}
combo_assoc_df <- fev1_fvc_expom |> 
    extract_results(result = "association") |> 
    pluck("assoc_exposures",
          "results_df") |> 
    dplyr::select(term,
                  estimate,
                  std.error,
                  p.value,
                  category) |> 
    left_join(sens_df |> 
                filter(sign_consistency>.5) |> 
                  dplyr::select(exposure,
                                median_effect,
                                sign_consistency),
              by=c("term"="exposure")) |> 
    left_join(broom::tidy(enet_res$glmnet.fit) |>
                  filter(lambda == enet_res$lambda.min,
                         term != "(Intercept)") |> 
                  dplyr::select(term,
                                enet_estimate=estimate)) |> 
    mutate(
        sig=p.value<0.05,
        term = case_when(
            sig == TRUE ~ paste0("* ",term),
            .default =term),
        term = forcats::fct_reorder(term, estimate),
        category = forcats::fct_reorder(
            category, estimate, .fun = median, .desc = TRUE)
    )

p_assoc <-
    combo_assoc_df |>
    ggplot(aes(x = estimate,
               y = term,
               color = estimate,
               fill = estimate)) +
    geom_point(shape = 21, size = 2.5, stroke = 0.8) +
    geom_errorbarh(
        aes(xmin = estimate - 1.96 * std.error,
            xmax = estimate + 1.96 * std.error),
        height = 0.15
    ) +
    ggh4x::facet_grid2(
        category ~ .,
        scales = "free",
        space = "free",
        strip = ggh4x::strip_themed(
            background_y = ggh4x::elem_list_rect(
                fill = cat_colors[levels(assoc_df$category)]
            )
        )
    ) +
    scale_color_gradient2(
        low = "blue4", mid = "grey85", high = "red4", midpoint = 0
    ) +
    scale_fill_gradient2(
        low = "blue4", mid = "grey85", high = "red4", midpoint = 0
    ) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    labs(x = "Effect size", y = NULL,color="Effect Size",fill="Effect Size") +
    theme_bw(base_size = 13) +
    theme(panel.grid.minor = element_blank(),
          legend.position = "bottom",
          strip.text.y = element_blank(),
          strip.background.y = element_blank())

p_stability <-
    combo_assoc_df |>
    ggplot(aes(x = sign_consistency,
               y = term,
               color = median_effect)) +
    geom_segment(
        aes(x = 0, xend = sign_consistency,
            yend = term),
        linewidth = 0.8
    ) +
    geom_point(size = 2.5) +
    ggh4x::facet_grid2(
        category ~ .,
        scales = "free",
        space = "free",
        strip = ggh4x::strip_themed(
            background_y = ggh4x::elem_list_rect(
                fill = cat_colors[levels(assoc_df$category)]
            )
        )
    ) +
    scale_color_gradient2(
        low = "blue4",
        mid = "grey85",
        high = "red4",
        midpoint = 0,
        limits = c(-0.15, 0.15),
        oob = scales::squish
    ) +
    scale_fill_gradient2(
        low = "blue4",
        mid = "grey85",
        high = "red4",
        midpoint = 0,
        limits = c(-0.15, 0.15),
        oob = scales::squish
    ) +scale_x_continuous(breaks = c(0,0.5,1))+
    labs(x = "Effect consistency", y = NULL, color = "Median Effect") +
    theme_bw(base_size = 13)+
    theme(panel.grid.minor = element_blank(),
          legend.position = "bottom",
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          strip.text.y = element_blank(),
          strip.background.y = element_blank(),)

p_enet <-
    combo_assoc_df |>
    ggplot(aes(x = enet_estimate,
               y = term,
               color = enet_estimate)) +
    geom_segment(
        aes(x = 0, xend = enet_estimate,
            yend = term),
        linewidth = 0.8
    ) +
    geom_point(size = 2.5) +
    ggh4x::facet_grid2(
        category ~ .,
        scales = "free",
        space = "free",
        strip = ggh4x::strip_themed(
            background_y = ggh4x::elem_list_rect(
                fill = cat_colors[levels(assoc_df$category)]
            )
        )
    ) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    scale_color_gradient2(
        low = "blue4",
        mid = "grey85",
        high = "red4",
        midpoint = 0,
        limits = c(-0.05, 0.05),
        oob = scales::squish
    ) +
    scale_fill_gradient2(
        low = "blue4",
        mid = "grey85",
        high = "red4",
        midpoint = 0,na.value = "white",
        limits = c(-0.05, 0.05),
        oob = scales::squish
    ) +
    labs(x = "Elastic-net coefficient", y = NULL,fill="Effect Size",color="Effect Size") +
    theme_bw(base_size = 13)+
    theme(panel.grid.minor = element_blank(),
          strip.text.y = element_text(angle = 0, face = "bold.italic"),
          legend.position = "bottom",
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank())

(p_assoc | p_stability | p_enet) +
    patchwork::plot_layout(widths = c(3, 3, 3))
```
### * Aggregate Plot 3

```{r}
combo_assoc_df <- fev1_fvc_expom |> 
    extract_results(result = "association") |> 
    pluck("assoc_exposures",
          "results_df") |> 
    dplyr::select(term,
                  estimate,
                  std.error,
                  p.value,
                  category) |> 
    left_join(sens_df |> 
                  #filter(sign_consistency>.5) |> 
                  dplyr::select(exposure,
                                median_effect,
                                sign_consistency),
              by=c("term"="exposure")) |> 
    left_join(broom::tidy(enet_res$glmnet.fit) |>
                  filter(lambda == enet_res$lambda.min,
                         term != "(Intercept)") |> 
                  dplyr::select(term,
                                enet_estimate=estimate)) |> 
    mutate(
        sig=p.value<0.05,
        term = case_when(
            sig == TRUE ~ paste0("* ",term),
            .default =term),
        term = forcats::fct_reorder(term, estimate),
        category = forcats::fct_reorder(
            category, estimate, .fun = median, .desc = TRUE)
    )

p_assoc <-
    combo_assoc_df |>
    ggplot(aes(x = estimate,
               y = term,
               color = estimate,
               fill = estimate)) +
    geom_point(shape = 21, size = 2.5, stroke = 0.8) +
    geom_errorbarh(
        aes(xmin = estimate - 1.96 * std.error,
            xmax = estimate + 1.96 * std.error),
        height = 0.15
    ) +
    ggh4x::facet_grid2(
        category ~ .,
        scales = "free",
        space = "free",
        strip = ggh4x::strip_themed(
            background_y = ggh4x::elem_list_rect(
                fill = cat_colors[levels(assoc_df$category)]
            )
        )
    ) +
    scale_color_gradient2(
        low = "blue4", mid = "grey85", high = "red4", midpoint = 0
    ) +
    scale_fill_gradient2(
        low = "blue4", mid = "grey85", high = "red4", midpoint = 0
    ) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    labs(x = "Effect size", y = NULL,color="Effect Size",fill="Effect Size") +
    theme_bw(base_size = 13) +
    theme(panel.grid.minor = element_blank(),
          legend.position = "right",
          strip.text.y = element_blank(),
          strip.background.y = element_blank())

p_stability <-
    combo_assoc_df |>
    ggplot(aes(x = 1,
               y = term,
               color = sign_consistency)) +
    geom_point(aes(size=sign_consistency)) +
    ggh4x::facet_grid2(
        category ~ .,
        scales = "free",
        space = "free",
        strip = ggh4x::strip_themed(
            background_y = ggh4x::elem_list_rect(
                fill = cat_colors[levels(assoc_df$category)]
            )
        )
    ) +
    scale_color_gradientn(
        colors = c(
            "#3a1c71",
            "#d76d77",
            "#ffaf7b"
        ),
        na.value = "grey80")+
    scale_x_continuous(breaks = c(0,0.5,1))+
    labs(x = "Effect consistency", y = NULL, color = "Effect Consistency") +
    theme_void(base_size = 13)+guides(size="none")+
    theme(panel.grid.minor = element_blank(),
          legend.position = "right",
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          strip.text.y = element_blank(),
          strip.background.y = element_blank(),)

p_enet <-
    combo_assoc_df |>
    ggplot(aes(x = 1,
               y = term,
               color = enet_estimate)) +
    geom_point(aes(size = enet_estimate)) +
    ggh4x::facet_grid2(
        category ~ .,
        scales = "free",
        space = "free",
        strip = ggh4x::strip_themed(
            background_y = ggh4x::elem_list_rect(
                fill = cat_colors[levels(assoc_df$category)]
            )
        )
    ) +
    #geom_vline(xintercept = 0, linetype = "dashed") +
    scale_color_gradient2(
        low = "#5D26C1",
        mid = "grey85",
        high = "#4BC0C8",
        midpoint = 0,
        limits = c(-0.05, 0.05),
        oob = scales::squish
    ) +
    labs(x = "Elastic-net coefficient", y = NULL,color="Enet Effect Size") +
    theme_void(base_size = 13)+guides(size="none")+
    theme(panel.grid.minor = element_blank(),
          strip.text.y = element_text(angle = 0, face = "bold.italic"),
          legend.position = "right",
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank())

(p_assoc | p_stability | p_enet) +
    patchwork::plot_layout(widths = c(3, 1, 1),guides = "collect")
```

## Interaction Analysis

```{r}
# pairs <- combn(exp_vars, 2, simplify = FALSE)
# 
# meta <- tidyexposomics::pivot_sample()(fev1_fvc_expom)
# 
# library(furrr)
# 
# plan(multisession)
# 
# interaction_results <- future_map_dfr(
#   pairs,
#   ~ test_pair(meta, .x[1], .x[2]),
#   .progress = TRUE
# ) |> 
#   filter(mode != "Other")
```


### Effect Direction Barplot

```{r}
# interaction_results |>  
#     tabyl(effect_direction) |> 
#   ggplot(aes(
#     x=n,
#     y=reorder(effect_direction   ,n),
#     fill=effect_direction
#   ))+
#   geom_bar(stat ="identity",
#            color="black",
#            linewidth = 0.005)+
#   ggsci::scale_fill_cosmic()+
#   theme_classic()+
#   labs(
#     x="Freq.",
#     y="",
#     fill="Freq."
#   )
```

### Frequent Negative interactions

```{r}
# interaction_results |>  
#     filter(grepl("Negative", effect_direction)) |> 
#     select(var1, var2) |> 
#     pivot_longer(cols = everything(), values_to = "variable") |> 
#     dplyr::count(variable, sort = TRUE) |> 
#   slice_head(n=15) |> 
#   ggplot(aes(
#     x=n,
#     y=reorder(variable,n),
#     fill=n
#   ))+
#   geom_bar(stat ="identity",
#            color="black",
#            linewidth = 0.005)+
#   theme_classic()+
#   scale_fill_gradient(
#     low = "#c31432",
#     high = "#240b36"
#   )+
#   labs(
#     x="Freq.",
#     y="",
#     fill="Freq."
#   )
```

### Fill by Interaction Type

```{r}
# interaction_results|>  
#     filter(grepl("Negative", effect_direction)) |> 
#     pivot_longer(cols = c(var1, var2), values_to = "variable") |> 
#     dplyr::count(variable, effect_direction) |> 
#   group_by(variable) |> 
#   mutate(total=sum(n)) |> 
#   ungroup() |> 
#   filter(variable %in% (
#     interaction_results|>  
#     filter(grepl("Negative", effect_direction)) |> 
#     pivot_longer(cols = c(var1, var2), values_to = "variable") |> 
#     dplyr::count(variable) |> 
#       arrange(desc(n)) |> 
#       slice_head(n=15) |>
#       pull(variable)
#   )) |> 
#     ggplot(aes(y = reorder(variable,total), 
#                x = n, 
#                fill = effect_direction)) +
#   geom_bar(stat ="identity",
#            color="black",
#            linewidth = 0.005)+
#   ggsci::scale_fill_cosmic()+
#   theme_classic()+
#   labs(
#     x="Freq.",
#     y="",
#     fill="Freq."
#   )
```

### Number of Interaction Types

```{r}
# interaction_results |>  
#     filter(grepl("Negative", effect_direction)) |> 
#     select(var1, var2) |> 
#     pivot_longer(cols = everything(), 
#                  values_to = "variable") |>
#   inner_join(aw_cb,by="variable") |> 
#   tabyl(category) |> 
#   arrange(desc(n)) |> 
#   ggplot(aes(
#     x=n,
#     y=reorder(category   ,n),
#     fill=category
#   ))+
#   geom_bar(stat ="identity",
#            color="black",
#            linewidth = 0.005)+
#   scale_fill_manual( values=cat_colors)+
#   theme_classic()+
#   labs(
#     x="Freq.",
#     y="",
#     fill="Freq."
#   )
```

### Network Plot

```{r}
# # Build edges
# edges <- interaction_results |>
#     filter(grepl("Negative", effect_direction)) |>
#     select(from = var1, to = var2, int_est)
# 
# # Convert to tidygraph 
# g_tbl <- as_tbl_graph(
#   igraph::graph_from_data_frame(
#     edges, 
#     directed = FALSE)) |>
#     activate(nodes) |>
#     left_join(aw_cb |>
#                 select(variable, category),
#               by = c("name" = "variable")) |>
#     mutate(
#         degree = centrality_degree(),      
#         top10  = name %in% names(
#           sort(degree, decreasing = TRUE))[1:10]
#     )
# 
# # Plot
# ggraph(g_tbl) +
#     geom_edge_link(aes(
#       edge_width = abs(int_est)), 
#       alpha = 0.4, 
#       color = "grey80") +
#     geom_node_point(aes(
#       size = degree, 
#       color = category),
#       alpha = 0.9) +
#   geom_node_point(aes(
#       size = degree, 
#       color = category),
#                   color = "black",
#                   stroke = 0.005,
#                   shape = 21,
#                   fill = NA) +
#     geom_node_label(
#       aes(label = ifelse(top10, name, NA)),
#       size = 3, repel = TRUE) +
#   scale_color_manual( values=scales::alpha(get_palette("aaas",15), 0.7))+
#     scale_size_continuous(range = c(2, 10)) +
#     theme_void()
```


## WQS Regression

```{r}
# group_list <- aw_cb |> 
#   filter(category %in% c(
#     "Indoor Air",
#   "Parabens",
#   "Phenols",
#   "Antimicrobials",
#   
#   "Dust Allergens",
#   "IgE Respiratory",
#   "IgE Superantigen",
#   "IgE Food",
#   
#   "Serum Essential Metals",
#   "Serum Non-Essential Metals",
#   "Urine Essential Metals",
#   "Urine Non-Essential Metals"
#   )) |> 
#   group_by(category) |>
#   summarise(vars = list(variable)) |> 
#   deframe()
# 
# 
# bws_res <- map(names(group_list), ~{
#   run_bws_group(
#     group_name = .x,
#     dat=fev1_fvc_expom |> 
#       tidyexposomics::pivot_sample()(),
#     outcome = "pftfev1fvc_actual",
#     covariates =c(
#       "gli_age",
#       "gli_sex",
#       "gli_height",
#       "fis"
#     ),
#     family = "gaussian")
# })
```

## Summary

* Three clusters, identified using DIANA clustering, appear to be driven by IgE subtypes and urinary metals.

* Metadata variables are highly correlated within variable subtypes. 

* Age is highly correlated with spirometry metrics.

* FEV1/FVC and FEV1 % Predicted are correlated with similar metrics - parental smoking, increased particulate matter, and IgE (ragweed & peanut) are associated with improved lung function and eosinophils are associated with reduced lung function. The directionality of the smoking status/particulate matter maybe explained by the fact that children of smokers are bothered by their asthma more and are using more medication.

* Increased Eosinophil levels are associated with monocytye, basophil, and platelet levels, exhaled nitric oxide, and total IgE levels.

## Save Data

```{r}
# Save MultiAssayExperiment Objects
saveRDS(fev1_fvc_expom,file = "../results/exwas/fev1_fvc_expom.rds")

# Save Bootstrapping Results
saveRDS(exwas_stability_res,file =
          "../results/exwas/exwas_stability_res.rds")
```

## Session Info

```{r}
sessionInfo()
```

